### 2017-09-14 Texevier-master/Readme.Rmd
final draft with all the relevant code and output data. Note than running the R code (rolling period code.R) takes some time and so the esults where stored as text files names "h step ahead results" for h=1,5,20

---
output:
  md_document:
    variant: markdown_github
---

# Introduction

This package helps a user set up an Elsevier template for writing academic journal reports in Rmarkdown. The setup requires very few lines of code and will get you up and running in _a few seconds_ (provided you have all the relevent dependencies installed - see below), with a fully working and good-to-go template.

Using Texevier you can simply start writing your paper / thesis / phd in R, with a template guiding you through the most important aspects you will encounter in your writing. Your template will look as follows:

![Screenshot](inst/ScreenShot/Example.PNG)

## Dependencies Required

Make sure you have the following installed on your computer:

* [R](http://cran.r-project.org/bin/windows/base/)
* [Rstudio](http://www.rstudio.com/)

## Optional (Should be installed automatically)

* [MikTex](http://miktex.org/download)

## Code

After installing all three at the top, simply run the following code:

    if (!require("devtools")) install.packages("devtools")
    library(devtools)
    devtools::install_github("Nicktz/Texevier")
    library(Texevier)
    # Input your directory as dir:
    dir <- "YOUR DIRECTORY HERE"
    create_template(directory = dir, template_name = "Template", launch_template = TRUE, ShowPDFatLaunch = TRUE)

This will automatically install all the needed files and components, and allow you to start working with an Elsevier template in less than a minute. A pdf will pop up, if you keep ShowPDFatLaunch = TRUE, which you can scroll through to see what your template looks like. REMEMBER to close the pdf before attempting to edit the template. Conversely, set ShowPDFatLaunch = FALSE. By default your template R files will be launched. To only create the folder, set launch_template = FALSE.

A template.Rmd file will also be opened automatically from the directory provided, which contains all the needed notes on how to write an academic paper in R to produce the pdf example shown. I recommend going through the code in depth for a good reference before you start working. Also, start your paper journey by creating a R.proj within your working folder.

## Motivation
Doing research in R allows you to verify your work easily. This template is designed to save the researcher time in terms of setting up a proper template for doing formal research. This is something which could feel daunting and can be time-consuming for even intermediate LaTeX  and R users. This package is intended to get you started in less than a minute.

Feel free to edit the templates and use this as a starting point for your own research.

### 2017-09-14 template/Draft paper without UCT guidelines.Rmd
final draft with all the relevant code and output data. Note than running the R code (rolling period code.R) takes some time and so the esults where stored as text files names "h step ahead results" for h=1,5,20

---
title: "Draft Paper w/o UCT template guideline"
author: "Merveilleuse"
date: "August 29, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## INTRODUCTION

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2017-09-14 template/Draft_Paper.Rmd
final draft with all the relevant code and output data. Note than running the R code (rolling period code.R) takes some time and so the esults where stored as text files names "h step ahead results" for h=1,5,20

---
# IMPORTANT: Change settings here, but DO NOT change the spacing. 
# Remove comments and add values where applicable. 
# The descriptions below should be self-explanatory
title: "Comparing Facebook Prophet Forecasts to those of an ARIMA Model"
subtitle: "A Diebold-Mariano Evaluation of the FTSE/JSE Top40 Index"
Author: "Marvelous Mubenesha"  # First Author
Student: "MBNMAR005"
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 5 # Adjust default figure sizes. This can also be done in the chunks of the text
    fig_height: 4
abstract: |
  Facebook's data science team open-sourced Prophet, a package that allows analysts to forecast a wide range of business time series at-scale. Prophet employs intensive Bayesian modelling in two tiers. Purely through MAP parameter estimation, and in a quasi-Bayesian form through an automated forecast evaluation that enables analyst to improve the model is it falls short when compared to other models. This paper assess the predictive accuracy of Prophet forecasts to those of an ARIMA model that was identified using the Box-Jenkins methodology. This is achieved through a Diebold-Mariano evaluation of rolling period forecasts of returns on the FTSE/JSE Top40 Index across daily, weekly and monthly time horizons. The results of the study suggest that there is insufficient evidence to conclude that the forecasting models have unequal predictive ability within the scope of the JSE/Top40 Index. 
---
<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
# Note: Include = FALSE implies the code is executed, but not printed in your pdf.
# warning and message = FALSE implies ugly messages and warnings are removed from your pdf. These should be picked up when you execute the command chunks (code sections below) in your rmd, not printed in your paper!

```

<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Introduction\label{Introduction}

In business, including that involveing the stock market, analyyst ay require a large number of time series forecasts to infomr decision making.The challenge with existing reliable approaches to forecasting time series data are that the models require a considerable background in statistics to build and adapt to prevailing market conditions when needed. Several authors have recommended automatic forecasting, which are algorithms that can automatically forecast a large number of univariate time series. However, the forecasts generated by these methods have been found to be brittle as well as inflexible in allowing for prior information to be incorporated into the forecasting model [@taylor2017forecasting]. In February of 2017, Facebook's Core Datascience team open sourced ,Prophet, their time series forecasting tool whose main selling point is that it enables "forecasting-at-scale". A term coined by the data science team which is defined as, "an approach that allows a large number of analysts to forecast a large number and variety of business time series" [@taylor2017forecasting]. It has the added advantage of being less 'expensive'
than other alternatives. In addition, it is specifically prefarable over ARIMA models because of its non-linearity, flexibility and ability to accomodate varying time intervals [@taylor2017forecasting].

The objective of this paper is to compare the forecasts generated by Prophet - which uses automated, intensive Bayesian Modelling to 'forecast at scale'-, to those obtained by an ARIMA model -constructed through the Box-Jenkins methodology-,
through a Diebold-Mariano evaluation of the JSE Top40 Index. The study is particularly useful because analysts from a wide background can be able to easily forecast stock price returns, automatically and by using their prior knowledge, without a wholesome understanding of the statistical intricacies involved.

The paper is organized as follows. A review of literature relating to stock price forecasting with ARIMA models an other alternative approaches is introduced to give the reader a background into the subject matter. Thereafter, theory relating to the mathematical formulation and estimation of both the ARIMA and Prophet models is included in sections 2.2 and 2.3 respectively, to build the readers understanding of what the models look like and how they operate. Information relating to the data that was used in this study is contained in section 3, after which, the methodology used to answer the research question is described. Section 5, contains the results obtained when the methodology was executed. whilst section 6 includes a discussion of the results and how they can be interpreted to answer the research question. Lastly, the conclusion is contain in section 7 and is aimed at highlighting the main findings of the study.

# Background 
In this section we will briefly review literature that relates to the comparison of different approaches to forecasting stock returns. This will include both traditional approaches as well new methods such as automatic forecasting, machine learning techniques and quasi-Bayesian forms which incorporate analysts' prior information. The review of literature is concluded by emphasing how the findings of this paper can compliment existing research. The theoretical and mathematical formulation of an ARIMA forecasting model is  then briefly introduced. Thereafter, the theoretical framework and mathematical formulation of Facebooks' Prophet forecasting tool is discussed. Specific attention is drawn to Prophets distinguising features to give the reader an understanding of the value that this forecasting technique can potentially add to stock price prediction. Lastly, the Diebold-Mariano Evaluation technique is introduced as a method of formally testing superior predictive accuracy between two forecasting models.

##Literature 
###Forecasting stock prices using the ARIMA model

ARIMA models have been widely used as a standard model to forecast financial data and have been shown to yield acceptable forecasting errors [@adebiyi2014comparison]. @adebiyi2014comparison used the Box-Jenkins methodology to build short-term stock price prediction models for stocks on the New York Stock Exchange(NYSE) and the Nigerian Stock Exchange(NSE). Their results showed that the prediction errors of the model were within acceptable bounds.

ARIMA models rely on the assumption that residuals are heteroskedastic(they have a constant variance) and normally distributed, however, some financial data displays heteroskedasticity making GARCH models more applicable. This theory is backed by a study conducted by @Muten2014 who compared the ability of ARIMA and GARCH models to forecast stock prices on the Zimbabwean Stock Exchange (ZSE). They found that the GARCH model outperforms the ARIMA model which suggests that the residuals are heteroskedastic. This is a case in point for emerging market stocks. The researchers noted that poor liquidity in the market could be a cause of the results they observed [@Muten2014].

###Forecasting stock prices using alternative methods: The case of Artificial Neural Networks and Hybrid Models

Though the ARIMA model is tractable given that it is simple, interpretable and yields forecasts that are significantly accurate when compared to other methods, it has its limitations. The most popular being its inability to capture non-linear patterns in data, even after its evolution from the standard form to more adaptable formulations [@moreno2011artificial]. Over the past two to three decades with the evolution of computational power and statistical advancement, other stock price prediction methods have been proposed, the most prominent being a machine learning approach called Artificial Neural Networks(ANNs)[@lin2009short]. Also popular amongst the new stock price forecasting approaches are hybrids of existing methods that incorporate the benefits of different approaches.

Artificial Neural Networks (ANNs) are a multi-layered perceptron as illustrated in  figure 1.  
![](ANN.png)  
*Figure 1 : Illustration of an Artificial neural network* [@tanikic2012artificial]  
An ANN consists of a sorted triple $(K, A, \omega)$. $K$ is a set of a multi-layered structure, V a paired set, that represents the value of the estimate in the $i^{th}$ node of the $k^{th}$ layer of the network. In Figure 1, the enlarged $a_{i,k} \in A$. Lastly, $\omega$ represents a function that defines the weights ($w_{i,r}$) of connections between neurons [@kriesel2007brief]. Artificial Neural networks apply iterated optimization of model parameters across the network in the form of weights conditional on observed values to learn [@segaran2007programming]. Several studies have compared ARIMA model forecasts to ANN and their conclusions are contradictory.

Researchers have found that the performance of either method depends on the nature of the data and forecasting problem [@kihoro2004seasonal]. Stock price data is proposed to be nonlinear which suggests that nonlinear approaches have the potential to produce better forecasts than linear models. Furthermore, ANNs make no assumptions about the distribution of the errors as compared to the linear ARIMA model [@adebiyi2014comparison]. @adebiyi2014comparison compared NYSE stock index forecasts of an ARIMA model to those of an ANN and found that the forecasting accuracy of the ANN model was superior to that of the ARIMA model. It is evident that ANN are preferred to ARIMA models as a model free, nonlinear alternative. However, the model construction of an ANN requires trial and error to initialize parameter estimates and these parameters are not easily interpretable by analysts [@moreno2011artificial]. Researchers have found different ways to guide analysts in this respect. In 1996, @wang1996stock proposed an ARIMA-based ANN to forecast the medium-term price of the Taiwan Stock Exchange Weighted Stock Index (TSEWSI). They used the Box-Jenkins methodology to difference the series and then trained the data on a neural network with initialisations that were guided by their observations of the ARIMA model. This approach yielded forecasts with an acceptable prediction accuracy based on residual analysis of the out-of-sample data. Wang and Leu therefore concluded that the ARIMA-based Neural Network outperformed a Neural Network trained using raw stock price data [@wang1996stock]. @zhang2009stock used a combination of the backpropagation algorithm from Neural Networks with Improved Bacterial Chemotaxis Optimization (IBCO) to build a model that forecasts the S&P 500 stock index by minimizing the mean square error. Model forecasts were evaluated through simulation experiments and the results led to the conclusion that the hybrid model produced superior forecasts [@zhang2009stock]. This study further highlighted the potential that nonlinear approaches have in forecasting stock prices. An issue that arises with such methods is the complexity of the proposed models which limits the flexibility of unseasoned analysts to adjust model parameters as a way of improving forecasting accuracy. As we will see, Prophet forecasting elegantly deals with this issue in the form of a nonlinear, Generalized Additive Model (GAM).

###Quasi-Bayesian methodss incorporating Analysts' knowlegde

The stock price forecasting approaches considered so far are purely apply technical data analysis methods to generate stock price forecasting models. However, studies that have compared pure technical approaches to those that incorporate the analysts knowledge in a quasi-Bayesian form have been shown to be superior at predicting market prices [@givoly1984quality \& @guerard1989combining]. The first instance of combining time-series model forecasts and an analysts forecasts to obtain superior forecasts of a stocks annual earnings can be dated to 1989. @guerard1989combining used an additive model to combine consensus security analyst forecasts from the S&P Annual Earnings forecaster and annual earnings forecasts generated by an ARIMA model with a constant. The combined model that was estimated using ordinary least squares reduced the mean square error of the time series and analysts forecasts from 1.28 and 1.27, respectively to 1.04. These results suggest that analysts can substantially reduce forecasting errors by combining both approaches [@guerard1989combining]. @zahedi2015application applied the same overarching principle when they used principle component analysis to determine an appropriate input variable, which they then used to train an ANN in an attempt to predict stock prices on the Tehran Stock Exchange. Their model yielded acceptable forecasting errors and performed better than the pure ANN which further reiterates the potential of incorporating an analysts knowledge to improve stock price forecasts, in both linear and nonlinear models.

###Diebold-Mariano Evaluation
As is evident from the previous sections, the modelling approach and assumptions of Prophet forecasting-at-scale differ to those of an ARIMA model. Since the two methods differ, the Diebold-Mariano evaluation (DM Test) is preferred as a means of testing forecasting accuracy. This is because the DM Test is a model free test of forecasting accuracy. It is applicable to a wide range of situations from multi-period forecasts, non-Gaussian forecast residuals, non-quadratic loss functions and even serially correlated data[@Mariano2000].

## ARIMA Model Formulation

The ARIMA model is a generalisation of Autoregressive Moving Average (ARMA) models which combine autoregression and moving average features [@box1970time]. In the context of returns, the ARIMA formulation for the return at time index $t$, $r_t$, is given by,
$$  r_t = \phi_1 r_{t-1} + \phi_2 r_{t-2} + ...+ \phi_p r_{t-p} + \epsilon_t +
        \psi_1 \epsilon_{t-1} + \psi_2 \epsilon_{t-2} + ... +\psi_q \epsilon_{t-q} $$  
where:  
$\bullet \phi_i$ and $\phi_j$ are the parameters to be estimated;  
$\bullet r_{t-k}$ is the $k^{th}$ lagged return;   
$\bullet \epsilon_t$ is the error term at time $t$ which is assumed to be white noise and  
$\bullet\epsilon_{t-k}$ is the $k^{th}$ lagged error.   
$\bullet$p is the order of the autoregressive component (dependenceon history of $r_t$)
$\bullet$q is the order of the moving average component (dependenceon history of $\epsilon_t$)

The ARIMA model will be used as a standard forecasting method which Prophet will be compared against. Model building and parameter estimation will follow the approach outlined by @box1970time and is further discussed in the methodology

## Prophet Model Formulation and Estimation

The challenge with many forecasting methods such as the Box-Jenkins methodology and machine-learning approaches are that they usually require a sufficient statistical background. Furthermore, these approaches tend to be inflexible with respect to incorporating prior information. Though this can be partially acheived by initializing parameter estimates for methods such as Artificial Neural Networks ,an analyst with no statistical training can have difficulty interpreting and thus modifying parameter estimates[@taylor2017forecasting].  
Prophet is an **automatic forecasting** tool that uses a Bayesian Generalized Additive model to generate time series forecasts. This is in comparison to conventional methods such as the linear stochastic dependence that exists in ARIMA models. Prophet combines a configurable model that includes performance analysis evaluation with the interaction of an analyst in the loop [@taylor2017forecasting]. The model is configurable which allows an analyst to incorporate their knowledge of the behaviour of the series into the model building process through easily interpretable initial parameters that can be modified and interactive feedback when forecasts under-perform [@taylor2017forecasting].

Through this mechanism, a large numer of forecasts can be reliably generated **automatically** with the flexibility of enabling an analyst to modify the model in-the-loop during model specification. The automated forecasting procedure with an analyst-in-the-loop is illustrated in Figure 2.

  The first step of the forecasting procedure begins with the analyst simply specifying a general model. Prophet then goes into the automated process of estimating model parameters and doing forecast evaluation. If forecasts are considered problematic based on evaluations of the Simulated Historical Forecast (SHF) errors, then the package will surface these problems. The process then transitions to the top left of the cycle in Figure 2 labelled analyst-in-the-loop. Here, the analyst can then use visually presented aspects of the results that raised the issues to modify model parameters before a better model is then estimated. The loop is repeated until forecast evaluation surfaces no problems.

![Figure 2](analyst in the loop.png){ width=100% }  
*Figure 2: Illustration of Prophets' automated forecasting procedure with an analyst in the loop* [@taylor2017forecasting]

The interpretability of Prophets' model parameters stems from the mathematical formulation of the forecasting model. The formulation posits that the series is generated by an additive, parametric function of time $r(t)$, defined by,
$$ r(t) = g(t) + s(t) +  h(t) + \epsilon_t.$$ The additive components include growth, seasonality and a component that adjusts for holidays as well as once-off events e.g. anticipated market shocks in the instance of financial time series.

The growth component, $g(t)$, is modelled using a generalized form of the logistic population growth model and is of the form;  
$$g(t) = \frac{C(t)}{1+exp(-(k+\boldsymbol{a}(t)^T \delta(t-(b+\boldsymbol{a}(t)^T \gamma)))}$$  
Where:  
$\bullet C(t)$ represents the carrying capacity of the growth component and can be modelled using a polynomial function of time, the simplest being a constant or linear model.  
$\bullet (k+\boldsymbol{a}(t)^T\delta)$ is the growth rate factor with the $\boldsymbol{a}(t)^T\delta$ term enabling the forecaster to choose where the growth rate changes( i.e change points) and   
$\bullet (b+\boldsymbol{a}(t)^T \gamma)$ is the adjusted offset parameter.

The growth component of the model is generalized with two useful features that allow the carrying capacity and growth rate of the model to vary with time. The model enables an analyst to manually define when and how the growth rate changes at different change points [@taylor2017forecasting]. This feature is particularly useful for forecasting stock price data since analysts are able to incorporate market movements that lead to the growth rate either decreasing, increasing or remaining constant.  
The seasonality component, $s(t)$ with periodicity $P$ and is given by,
$$s(t) = \sum_{n = -N}^{N}c_n \exp(j\frac{2\pi nt}{P})$$

Yearly and weekly seasonality is modelled using the standard Fourier series. Analysts can thus use prior knowledge to account for the effects of periodicities they have observed seasonally which is similar to the periodicity of Seasonal ARIMA models.

Lastly, $h(t)$ is the holiday and events component, a feature that ARIMA models are not adapted to. The holiday component enables analysts to incorporate shocks and once-off events that do not follow a periodic pattern but may have an effect on the price of a stock. Holidays and events are modelled using,

$$ h(t) = \sum_{j = 1}^{L}\kappa_j \boldsymbol{1}(t \in D_j) \label{eqn5}
$$
Where, $\kappa$ is normally distributed scaling factor and  $D_j$ is a set of past and future dates where holidays or events occur.

Prophet translates model parameter estimation into a curve-fitting exercise by employing Maximum Apriori Posterior (MAP) estimation, a quasi-Bayesian approach, to estimate the optimal parameter values for the forecasting model. MAP estimation finds the maximum posterior estimates for the parameters by using the likelihood function and an apriori distribution to find a posterior distribution which is optimized to return point estimates of the parameters [@taylor2017forecasting]. Thereafter, the package automatically evaluates the forecasts by comparing the Simulated Historical Forecast (SHF) errors of baseline forecasting methods (including ARIMA forecasts) to those of Prophet. SHF errors are forecast errors generated from random points in the history of the time series model. In Prophet, SHF errors are compared visually to allow the analysts to easily adjust the interpretable parameters before automatically estimating the model and evaluating forecasts.

#Data

The data set used throughout this paper comprises of the daily closing prices of the equally-weighted FTSE/JSE Top40 price index for the five year period from 31/07/2012 to 30/07/2017. This equates to a total of 1766 trading days excluding weekends and holidays. This period was chosen for its recency and to avoid exposing the results to the noise associated with the high volatility experienced in the stock market during the 2008 financial crisis. The aim is to compare the predictive accuracy of the two models under standard market conditions before other confounding factors can be closely analysed.  Figure 33 shows the actual realisations of the price index which was sourced from Datastream.

![Figure 2](plot of Price index.png){ width=100% }  
*Figure 3: Realisation of the FTSE/JSE Top40 Price Index over the period of study*

The price index was transformation to log returns as a means of avoid spurious regression issues which can be caused by the existence of unit roots in stock prices.

In this paper, the **log return**(which will be used interchangeably with **return**), over the period $[t,t+1]$ is denoted by $r_t$, and was calculated by evaluating;  
$$
r_t = log(\frac{P_{t+1}}{P_t})  \label{eq1}
$$  
Where $P_{t+1}$ and $P_{t+1}$ represent the price of the index at time $t$ and $t+1$ respectively.

#Methodology

Preliminary data processing was done to clean and transform the data before formally testing for outliers. The data was then split into in-sample and out-of-sample periods. The in-sample period comprises of 1400 trading days from 30 July 2012 to the 1 May 2016. The out_of_sample period comprises of the remaining 366 trading days from the dataset.

The in-sample data was used to identify the ARIMA forecasting model using the method outlined by @box1970time for building and estimating an ARIMA model.  
Thereafter, one-step rolling forecasts with re-estimation as outlined by @RJHyndman2014 were evaluated for an automated Prophet model and the ARIMA model that was identified using the Box-Jenkins methodology. This procedure was then repeated for a 5-step and 20-step forecast horizon in order to determine whether the performance of the models vary when estimating weekly or monthly stock returns.

The forecasts generated above where then used to calculate **h-step rolling forecast errors** for both the ARIMA and Prophet models. For each time point $t$ in the out-of-sample period, the rolling forecast error is given by,

$$ \hat{e_t} = r_t - \hat{r_t},$$  where $r_t$ is the actual return and $\hat{r_t}$ is the rolling forecast at time $t$.  
Summary statistics of the errors where used to compare the performance of the forecasting models before a formal statistical test was done using a Diebold-Mariano evaluation.

#Results
##Box-Jenkins Methodology: Arima Model Identification and Selection

To identify what ARIMA model to use, the methodology as outlined in @box1970time was applied to the data. Firstly, stationarity conditions were verified formally by performing an augmented Dickey-Fuller test on the in-sample returns data. The results of the test suggest the p_value is less that 1%, which implies that the in-sample returns series is stationary. The results in the autocorrelation(ACF) and partial autocorrelation(PACF) function illustrated in figure 4  suggest that the second lag of both functions might be significantly different from zero.  
![](acf & pacf plot.png){ width=100%}  
*Figure 4 : ACF and PACF plots for in-sample data*

Therefore, the candidate *ARIMA(p,d,q)* models in Table 1 were fit to the in-sample data and the AIC used to choose the model that best fits the data. In Table 1, the candidate model with the lowest AIC is the $ARIMA(2,0,2)$. This model was thus selected as the ARIMA forecasting model.  
![](AIC.png){ width=100% }  
*Table 1: AIC for Candidate ARIMA models*

## Forecast Error Evaluation
The results of model estimation and forecast evaluation using the rolling window for h-step forecasts yielded forecasts errors with roughly equal distributions. This is visible from the boxplot in Figure 5. The presence of outliers seems consistent across the different models and forecastign horizons. Furthermore, the variation in the accuracy across different forecasting horizons for each model are not observable in Figure 5. The mean error and spread of both models across different time horizons appears to be constant. 
![](Forecast error boxplot.png){ width=105% }  
*Figure 5: Boxplot illustrating distribution of rolling forecast errors for different time horizons, h*

### Forecast Error Statistics

Table 2 summarizes the forecast error statistics for both the ARIMA and Prophet model across the three different time horizons.  
Across all forecast horizons, the ARIMA model appears to be overestimating returns. This is evident from the negative mean error(ME) across $h = (1,5,20)$. On the contrary, the Prophet model underestimates the forecasts for one-step and five-step ahead forecasts. It then overstimates returns when forecasting twenty-steps ahead. The ME results shown in Table 2 fail to clearly illustrate which model performs better. Furthermore, the ME values are relatively small when compared against other error statistics.

![ErrorStats](error stats.png){ width=100% }  
*Table  2 : Forecast error diagnostic statistics over different forecasting horizons*

The Root Mean Square Error (RMSE) across different forecast horizons and across both models are considerably close to each other in value relative to the other error statistics. The same observation can be made for the Mean Absolute Errors (MAE). These results suggest that the models possibly have the same forecasting accuracy.

### Diebold-Mariano Evaluation

The observations made from Table 2 in section 5.2 suggest that the ARIMA and Prophet models have equal predictive accuracy. These assertions can be formally quantified through a Diebold-Mariano(DM) test which evaluates whether the models have equal predictive accuracy against the null hypothesis that one model has superior accuracy over the other. Table 3 reports the results obtained when the rolling forecast errors were formally compared against each other over different time horizons.  
![](dm results.png){ width=105%}  
*Table 3 Results of the Diebold-Mariano Hypothesis Test*  
The P-value across all the time horizons for either linear and quadratic loss functions suggest that we would fail to reject the null hypothesis which implies that the predictive accuracy of the models could be equal and the 95% level. This observation validates the results obtained in section 5.2.

#Discussion of Results
The results of the study show that there is no significant evidence to conclude that Prophet generates superior forecasts for returns on the FTSE/JSE Top40 Index when compared against those of an ARIMA model when the Diebold-Mariano Test is used as a method of evaluation. These results are consistent for daily, weekly and monthly forecasting horizons.  
Nevertheless, the study has flagged the possibility that, on average, the ARIMA model tends to overestimate forecasted returns on the index for the forecast horizons under consideration. This is contrary to Prophet, that, on average, underestimates forecasted returns over daily and weekly time horizons. This information could be useful for an analyst engaged in short-term strategies since the choice of the forecasting model can reflect a particular sentiment or risk appetite. For example, if the analyst feels like prevailing returns are underpriced by the market, they might choose to use an ARIMA model, whereas an opposite sentiment could warrant the use pf Prophet.  
Furthermore, it is worth noting that the merits of Prophet may have not been fully exposed in this study due to the choice of the forecasting horizon, i.e. excluding periods of high volatility and uncertainity during and after the 2008 Financial Crisis. Nonetheless, it is reassuring that the model has performed as well as the ARIMA model. This observation could create room for the application of Prophet in circumstances where flexibility in the model formulation is desireable.

#Conclusion

The challenge of building models that generate superior stock return forecasts has led to widespread research in the field of quantitative analysis for decades, but the ARIMA modelremains the standard model against which to compare new innovations. This is because of its relative simplicity and ability to evolve into more complex applications e.g in the hybrid models considered in the literature review  
Prophet has reduced the problem of estimating a time series model to a cure fittig exercise by defining an additive, parametric function of time. The functions' parameters are then estimated using Bayesian estimation methods after which the tool automatically executes forecast evaluation to optimize its predictive ability.  
The results of this study suggest that, using a Diebold-Mariano evaluation, ARIMA and Prophet do not have a significant difference in predictive accuracy, even across time horizons of up to a month. The implications of this are that analysts can then chosse which model to use based on their needs without significantly compromising forecast  accuracy.

# References

### 2017-09-14 template/Draft_Paper.tex
final draft with all the relevant code and output data. Note than running the R code (rolling period code.R) takes some time and so the esults where stored as text files names "h step ahead results" for h=1,5,20

\documentclass[12pt,a4paper]{article}

\usepackage{url}
\usepackage{appendix}
\usepackage[british]{babel}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{graphicx}
\numberwithin{equation}{section}
\numberwithin{figure}{section}
\numberwithin{table}{section}

\usepackage[round]{natbib}
\bibliographystyle{natbib}
\def\bibsection{\section*{References}}

% Wrap around which gives all figures included the [H] command, or places it "here". This can be tedious to code in Rmarkdown.
\usepackage{float}
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
    \expandafter\origfigure\expandafter[H]
} {
    \endorigfigure
}

\let\origtable\table
\let\endorigtable\endtable
\renewenvironment{table}[1][2] {
    \expandafter\origtable\expandafter[H]
} {
    \endorigtable
}

\def\tightlist{}

\bibpunct[:]{(}{)}{;}{a}{,}{,}

\renewcommand{\baselinestretch}{1.5}

\begin{document}

\begin{titlepage}

\begin{center}
{\Huge \bf Comparing Facebook Prophet Forecasts to those of an ARIMA Model}\\
\today\\
Marvelous Mubenesha ( MBNMAR005)\\
{\tt MBNMAR005@uct.co.za}
\end{center}

\begin{abstract}
Facebook's data science team open-sourced Prophet, a package that allows
analysts to forecast a wide range of business time series at-scale.
Prophet employs intensive Bayesian modelling in two tiers. Purely
through MAP parameter estimation, and in a quasi-Bayesian form through
an automated forecast evaluation that enables analyst to improve the
model is it falls short when compared to other models. This paper assess
the predictive accuracy of Prophet forecasts to those of an ARIMA model
that was identified using the Box-Jenkins methodology. This is achieved
through a Diebold-Mariano evaluation of rolling period forecasts of
returns on the FTSE/JSE Top40 Index across daily, weekly and monthly
time horizons. The results of the study suggest that there is
insufficient evidence to conclude that the forecasting models have
unequal predictive ability within the scope of the JSE/Top40 Index.
\noindent
Keywords: prophet forecasting, arima forecasting, Box-Jenkins methodology,Diebold-Mariano test,multi-step rolling window forecasts, FTSE/JSE Top40 index.
\end{abstract}
\end{titlepage}

\pagenumbering{arabic}

\section{\texorpdfstring{Introduction\label{Introduction}}{Introduction}}\label{introduction}

In business, including that involveing the stock market, analyyst ay
require a large number of time series forecasts to infomr decision
making.The challenge with existing reliable approaches to forecasting
time series data are that the models require a considerable background
in statistics to build and adapt to prevailing market conditions when
needed. Several authors have recommended automatic forecasting, which
are algorithms that can automatically forecast a large number of
univariate time series. However, the forecasts generated by these
methods have been found to be brittle as well as inflexible in allowing
for prior information to be incorporated into the forecasting model
(Taylor and Letham 2017). In February of 2017, Facebook's Core
Datascience team open sourced ,Prophet, their time series forecasting
tool whose main selling point is that it enables
``forecasting-at-scale''. A term coined by the data science team which
is defined as, ``an approach that allows a large number of analysts to
forecast a large number and variety of business time series'' (Taylor
and Letham 2017). It has the added advantage of being less `expensive'
than other alternatives. In addition, it is specifically prefarable over
ARIMA models because of its non-linearity, flexibility and ability to
accomodate varying time intervals (Taylor and Letham 2017).

The objective of this paper is to compare the forecasts generated by
Prophet - which uses automated, intensive Bayesian Modelling to
`forecast at scale'-, to those obtained by an ARIMA model -constructed
through the Box-Jenkins methodology-, through a Diebold-Mariano
evaluation of the JSE Top40 Index. The study is particularly useful
because analysts from a wide background can be able to easily forecast
stock price returns, automatically and by using their prior knowledge,
without a wholesome understanding of the statistical intricacies
involved.

The paper is organized as follows. A review of literature relating to
stock price forecasting with ARIMA models an other alternative
approaches is introduced to give the reader a background into the
subject matter. Thereafter, theory relating to the mathematical
formulation and estimation of both the ARIMA and Prophet models is
included in sections 2.2 and 2.3 respectively, to build the readers
understanding of what the models look like and how they operate.
Information relating to the data that was used in this study is
contained in section 3, after which, the methodology used to answer the
research question is described. Section 5, contains the results obtained
when the methodology was executed. whilst section 6 includes a
discussion of the results and how they can be interpreted to answer the
research question. Lastly, the conclusion is contain in section 7 and is
aimed at highlighting the main findings of the study.

\section{Background}\label{background}

In this section we will briefly review literature that relates to the
comparison of different approaches to forecasting stock returns. This
will include both traditional approaches as well new methods such as
automatic forecasting, machine learning techniques and quasi-Bayesian
forms which incorporate analysts' prior information. The review of
literature is concluded by emphasing how the findings of this paper can
compliment existing research. The theoretical and mathematical
formulation of an ARIMA forecasting model is then briefly introduced.
Thereafter, the theoretical framework and mathematical formulation of
Facebooks' Prophet forecasting tool is discussed. Specific attention is
drawn to Prophets distinguising features to give the reader an
understanding of the value that this forecasting technique can
potentially add to stock price prediction. Lastly, the Diebold-Mariano
Evaluation technique is introduced as a method of formally testing
superior predictive accuracy between two forecasting models.

\subsection{Literature}\label{literature}

\subsubsection{Forecasting stock prices using the ARIMA
model}\label{forecasting-stock-prices-using-the-arima-model}

ARIMA models have been widely used as a standard model to forecast
financial data and have been shown to yield acceptable forecasting
errors (Adebiyi, Adewumi, and Ayo 2014). Adebiyi, Adewumi, and Ayo
(2014) used the Box-Jenkins methodology to build short-term stock price
prediction models for stocks on the New York Stock Exchange(NYSE) and
the Nigerian Stock Exchange(NSE). Their results showed that the
prediction errors of the model were within acceptable bounds.

ARIMA models rely on the assumption that residuals are
heteroskedastic(they have a constant variance) and normally distributed,
however, some financial data displays heteroskedasticity making GARCH
models more applicable. This theory is backed by a study conducted by
Mutendadzamera and Mutasa (2014) who compared the ability of ARIMA and
GARCH models to forecast stock prices on the Zimbabwean Stock Exchange
(ZSE). They found that the GARCH model outperforms the ARIMA model which
suggests that the residuals are heteroskedastic. This is a case in point
for emerging market stocks. The researchers noted that poor liquidity in
the market could be a cause of the results they observed (Mutendadzamera
and Mutasa 2014).

\subsubsection{Forecasting stock prices using alternative methods: The
case of Artificial Neural Networks and Hybrid
Models}\label{forecasting-stock-prices-using-alternative-methods-the-case-of-artificial-neural-networks-and-hybrid-models}

Though the ARIMA model is tractable given that it is simple,
interpretable and yields forecasts that are significantly accurate when
compared to other methods, it has its limitations. The most popular
being its inability to capture non-linear patterns in data, even after
its evolution from the standard form to more adaptable formulations
(Moreno, Pol, and Gracia 2011). Over the past two to three decades with
the evolution of computational power and statistical advancement, other
stock price prediction methods have been proposed, the most prominent
being a machine learning approach called Artificial Neural
Networks(ANNs)(Lin, Yang, and Song 2009). Also popular amongst the new
stock price forecasting approaches are hybrids of existing methods that
incorporate the benefits of different approaches.

Artificial Neural Networks (ANNs) are a multi-layered perceptron as
illustrated in figure 1.\\
\includegraphics{ANN.png}\\
\emph{Figure 1 : Illustration of an Artificial neural network} (Tanikic
and Despotovic 2012)\\
An ANN consists of a sorted triple \((K, A, \omega)\). \(K\) is a set of
a multi-layered structure, V a paired set, that represents the value of
the estimate in the \(i^{th}\) node of the \(k^{th}\) layer of the
network. In Figure 1, the enlarged \(a_{i,k} \in A\). Lastly, \(\omega\)
represents a function that defines the weights (\(w_{i,r}\)) of
connections between neurons (Kriesel 2007). Artificial Neural networks
apply iterated optimization of model parameters across the network in
the form of weights conditional on observed values to learn (Segaran
2007). Several studies have compared ARIMA model forecasts to ANN and
their conclusions are contradictory.

Researchers have found that the performance of either method depends on
the nature of the data and forecasting problem (Kihoro, Otieno, and
Wafula 2004). Stock price data is proposed to be nonlinear which
suggests that nonlinear approaches have the potential to produce better
forecasts than linear models. Furthermore, ANNs make no assumptions
about the distribution of the errors as compared to the linear ARIMA
model (Adebiyi, Adewumi, and Ayo 2014). Adebiyi, Adewumi, and Ayo (2014)
compared NYSE stock index forecasts of an ARIMA model to those of an ANN
and found that the forecasting accuracy of the ANN model was superior to
that of the ARIMA model. It is evident that ANN are preferred to ARIMA
models as a model free, nonlinear alternative. However, the model
construction of an ANN requires trial and error to initialize parameter
estimates and these parameters are not easily interpretable by analysts
(Moreno, Pol, and Gracia 2011). Researchers have found different ways to
guide analysts in this respect. In 1996, Wang and Leu (1996) proposed an
ARIMA-based ANN to forecast the medium-term price of the Taiwan Stock
Exchange Weighted Stock Index (TSEWSI). They used the Box-Jenkins
methodology to difference the series and then trained the data on a
neural network with initialisations that were guided by their
observations of the ARIMA model. This approach yielded forecasts with an
acceptable prediction accuracy based on residual analysis of the
out-of-sample data. Wang and Leu therefore concluded that the
ARIMA-based Neural Network outperformed a Neural Network trained using
raw stock price data (Wang and Leu 1996). Zhang and Wu (2009) used a
combination of the backpropagation algorithm from Neural Networks with
Improved Bacterial Chemotaxis Optimization (IBCO) to build a model that
forecasts the S\&P 500 stock index by minimizing the mean square error.
Model forecasts were evaluated through simulation experiments and the
results led to the conclusion that the hybrid model produced superior
forecasts (Zhang and Wu 2009). This study further highlighted the
potential that nonlinear approaches have in forecasting stock prices. An
issue that arises with such methods is the complexity of the proposed
models which limits the flexibility of unseasoned analysts to adjust
model parameters as a way of improving forecasting accuracy. As we will
see, Prophet forecasting elegantly deals with this issue in the form of
a nonlinear, Generalized Additive Model (GAM).

\subsubsection{Quasi-Bayesian methodss incorporating Analysts'
knowlegde}\label{quasi-bayesian-methodss-incorporating-analysts-knowlegde}

The stock price forecasting approaches considered so far are purely
apply technical data analysis methods to generate stock price
forecasting models. However, studies that have compared pure technical
approaches to those that incorporate the analysts knowledge in a
quasi-Bayesian form have been shown to be superior at predicting market
prices (Givoly and Lakonishok 1984 \& Guerard (1989)). The first
instance of combining time-series model forecasts and an analysts
forecasts to obtain superior forecasts of a stocks annual earnings can
be dated to 1989. Guerard (1989) used an additive model to combine
consensus security analyst forecasts from the S\&P Annual Earnings
forecaster and annual earnings forecasts generated by an ARIMA model
with a constant. The combined model that was estimated using ordinary
least squares reduced the mean square error of the time series and
analysts forecasts from 1.28 and 1.27, respectively to 1.04. These
results suggest that analysts can substantially reduce forecasting
errors by combining both approaches (Guerard 1989). Zahedi and Rounaghi
(2015) applied the same overarching principle when they used principle
component analysis to determine an appropriate input variable, which
they then used to train an ANN in an attempt to predict stock prices on
the Tehran Stock Exchange. Their model yielded acceptable forecasting
errors and performed better than the pure ANN which further reiterates
the potential of incorporating an analysts knowledge to improve stock
price forecasts, in both linear and nonlinear models.

\subsubsection{Diebold-Mariano
Evaluation}\label{diebold-mariano-evaluation}

As is evident from the previous sections, the modelling approach and
assumptions of Prophet forecasting-at-scale differ to those of an ARIMA
model. Since the two methods differ, the Diebold-Mariano evaluation (DM
Test) is preferred as a means of testing forecasting accuracy. This is
because the DM Test is a model free test of forecasting accuracy. It is
applicable to a wide range of situations from multi-period forecasts,
non-Gaussian forecast residuals, non-quadratic loss functions and even
serially correlated data(Mariano 2000).

\subsection{ARIMA Model Formulation}\label{arima-model-formulation}

The ARIMA model is a generalisation of Autoregressive Moving Average
(ARMA) models which combine autoregression and moving average features
(Box and Jenkins 1970). In the context of returns, the ARIMA formulation
for the return at time index \(t\), \(r_t\), is given by,
\[  r_t = \phi_1 r_{t-1} + \phi_2 r_{t-2} + ...+ \phi_p r_{t-p} + \epsilon_t +
        \psi_1 \epsilon_{t-1} + \psi_2 \epsilon_{t-2} + ... +\psi_q \epsilon_{t-q} \]\\
where:\\
\(\bullet \phi_i\) and \(\phi_j\) are the parameters to be estimated;\\
\(\bullet r_{t-k}\) is the \(k^{th}\) lagged return;\\
\(\bullet \epsilon_t\) is the error term at time \(t\) which is assumed
to be white noise and\\
\(\bullet\epsilon_{t-k}\) is the \(k^{th}\) lagged error.\\
\(\bullet\)p is the order of the autoregressive component (dependenceon
history of \(r_t\)) \(\bullet\)q is the order of the moving average
component (dependenceon history of \(\epsilon_t\))

The ARIMA model will be used as a standard forecasting method which
Prophet will be compared against. Model building and parameter
estimation will follow the approach outlined by Box and Jenkins (1970)
and is further discussed in the methodology

\subsection{Prophet Model Formulation and
Estimation}\label{prophet-model-formulation-and-estimation}

The challenge with many forecasting methods such as the Box-Jenkins
methodology and machine-learning approaches are that they usually
require a sufficient statistical background. Furthermore, these
approaches tend to be inflexible with respect to incorporating prior
information. Though this can be partially acheived by initializing
parameter estimates for methods such as Artificial Neural Networks ,an
analyst with no statistical training can have difficulty interpreting
and thus modifying parameter estimates(Taylor and Letham 2017).\\
Prophet is an \textbf{automatic forecasting} tool that uses a Bayesian
Generalized Additive model to generate time series forecasts. This is in
comparison to conventional methods such as the linear stochastic
dependence that exists in ARIMA models. Prophet combines a configurable
model that includes performance analysis evaluation with the interaction
of an analyst in the loop (Taylor and Letham 2017). The model is
configurable which allows an analyst to incorporate their knowledge of
the behaviour of the series into the model building process through
easily interpretable initial parameters that can be modified and
interactive feedback when forecasts under-perform (Taylor and Letham
2017).

Through this mechanism, a large numer of forecasts can be reliably
generated \textbf{automatically} with the flexibility of enabling an
analyst to modify the model in-the-loop during model specification. The
automated forecasting procedure with an analyst-in-the-loop is
illustrated in Figure 2.

The first step of the forecasting procedure begins with the analyst
simply specifying a general model. Prophet then goes into the automated
process of estimating model parameters and doing forecast evaluation. If
forecasts are considered problematic based on evaluations of the
Simulated Historical Forecast (SHF) errors, then the package will
surface these problems. The process then transitions to the top left of
the cycle in Figure 2 labelled analyst-in-the-loop. Here, the analyst
can then use visually presented aspects of the results that raised the
issues to modify model parameters before a better model is then
estimated. The loop is repeated until forecast evaluation surfaces no
problems.

\includegraphics[width=1.00000\textwidth]{analyst in the loop.png}\\
\emph{Figure 2: Illustration of Prophets' automated forecasting
procedure with an analyst in the loop} (Taylor and Letham 2017)

The interpretability of Prophets' model parameters stems from the
mathematical formulation of the forecasting model. The formulation
posits that the series is generated by an additive, parametric function
of time \(r(t)\), defined by,
\[ r(t) = g(t) + s(t) +  h(t) + \epsilon_t.\] The additive components
include growth, seasonality and a component that adjusts for holidays as
well as once-off events e.g.~anticipated market shocks in the instance
of financial time series.

The growth component, \(g(t)\), is modelled using a generalized form of
the logistic population growth model and is of the form;\\
\[g(t) = \frac{C(t)}{1+exp(-(k+\boldsymbol{a}(t)^T \delta(t-(b+\boldsymbol{a}(t)^T \gamma)))}\]\\
Where:\\
\(\bullet C(t)\) represents the carrying capacity of the growth
component and can be modelled using a polynomial function of time, the
simplest being a constant or linear model.\\
\(\bullet (k+\boldsymbol{a}(t)^T\delta)\) is the growth rate factor with
the \(\boldsymbol{a}(t)^T\delta\) term enabling the forecaster to choose
where the growth rate changes( i.e change points) and\\
\(\bullet (b+\boldsymbol{a}(t)^T \gamma)\) is the adjusted offset
parameter.

The growth component of the model is generalized with two useful
features that allow the carrying capacity and growth rate of the model
to vary with time. The model enables an analyst to manually define when
and how the growth rate changes at different change points (Taylor and
Letham 2017). This feature is particularly useful for forecasting stock
price data since analysts are able to incorporate market movements that
lead to the growth rate either decreasing, increasing or remaining
constant.\\
The seasonality component, \(s(t)\) with periodicity \(P\) and is given
by, \[s(t) = \sum_{n = -N}^{N}c_n \exp(j\frac{2\pi nt}{P})\]

Yearly and weekly seasonality is modelled using the standard Fourier
series. Analysts can thus use prior knowledge to account for the effects
of periodicities they have observed seasonally which is similar to the
periodicity of Seasonal ARIMA models.

Lastly, \(h(t)\) is the holiday and events component, a feature that
ARIMA models are not adapted to. The holiday component enables analysts
to incorporate shocks and once-off events that do not follow a periodic
pattern but may have an effect on the price of a stock. Holidays and
events are modelled using,

\[ h(t) = \sum_{j = 1}^{L}\kappa_j \boldsymbol{1}(t \in D_j) \label{eqn5}
\] Where, \(\kappa\) is normally distributed scaling factor and \(D_j\)
is a set of past and future dates where holidays or events occur.

Prophet translates model parameter estimation into a curve-fitting
exercise by employing Maximum Apriori Posterior (MAP) estimation, a
quasi-Bayesian approach, to estimate the optimal parameter values for
the forecasting model. MAP estimation finds the maximum posterior
estimates for the parameters by using the likelihood function and an
apriori distribution to find a posterior distribution which is optimized
to return point estimates of the parameters (Taylor and Letham 2017).
Thereafter, the package automatically evaluates the forecasts by
comparing the Simulated Historical Forecast (SHF) errors of baseline
forecasting methods (including ARIMA forecasts) to those of Prophet. SHF
errors are forecast errors generated from random points in the history
of the time series model. In Prophet, SHF errors are compared visually
to allow the analysts to easily adjust the interpretable parameters
before automatically estimating the model and evaluating forecasts.

\section{Data}\label{data}

The data set used throughout this paper comprises of the daily closing
prices of the equally-weighted FTSE/JSE Top40 price index for the five
year period from 31/07/2012 to 30/07/2017. This equates to a total of
1766 trading days excluding weekends and holidays. This period was
chosen for its recency and to avoid exposing the results to the noise
associated with the high volatility experienced in the stock market
during the 2008 financial crisis. The aim is to compare the predictive
accuracy of the two models under standard market conditions before other
confounding factors can be closely analysed. Figure 33 shows the actual
realisations of the price index which was sourced from Datastream.

\includegraphics[width=1.00000\textwidth]{plot of Price index.png}\\
\emph{Figure 3: Realisation of the FTSE/JSE Top40 Price Index over the
period of study}

The price index was transformation to log returns as a means of avoid
spurious regression issues which can be caused by the existence of unit
roots in stock prices.

In this paper, the \textbf{log return}(which will be used
interchangeably with \textbf{return}), over the period \([t,t+1]\) is
denoted by \(r_t\), and was calculated by evaluating;\\
\[
r_t = log(\frac{P_{t+1}}{P_t})  \label{eq1}
\]\\
Where \(P_{t+1}\) and \(P_{t+1}\) represent the price of the index at
time \(t\) and \(t+1\) respectively.

\section{Methodology}\label{methodology}

Preliminary data processing was done to clean and transform the data
before formally testing for outliers. The data was then split into
in-sample and out-of-sample periods. The in-sample period comprises of
1400 trading days from 30 July 2012 to the 1 May 2016. The
out\_of\_sample period comprises of the remaining 366 trading days from
the dataset.

The in-sample data was used to identify the ARIMA forecasting model
using the method outlined by Box and Jenkins (1970) for building and
estimating an ARIMA model.\\
Thereafter, one-step rolling forecasts with re-estimation as outlined by
Hyndman (2014) were evaluated for an automated Prophet model and the
ARIMA model that was identified using the Box-Jenkins methodology. This
procedure was then repeated for a 5-step and 20-step forecast horizon in
order to determine whether the performance of the models vary when
estimating weekly or monthly stock returns.

The forecasts generated above where then used to calculate
\textbf{h-step rolling forecast errors} for both the ARIMA and Prophet
models. For each time point \(t\) in the out-of-sample period, the
rolling forecast error is given by,

\[ \hat{e_t} = r_t - \hat{r_t},\] where \(r_t\) is the actual return and
\(\hat{r_t}\) is the rolling forecast at time \(t\).\\
Summary statistics of the errors where used to compare the performance
of the forecasting models before a formal statistical test was done
using a Diebold-Mariano evaluation.

\section{Results}\label{results}

\subsection{Box-Jenkins Methodology: Arima Model Identification and
Selection}\label{box-jenkins-methodology-arima-model-identification-and-selection}

To identify what ARIMA model to use, the methodology as outlined in Box
and Jenkins (1970) was applied to the data. Firstly, stationarity
conditions were verified formally by performing an augmented
Dickey-Fuller test on the in-sample returns data. The results of the
test suggest the p\_value is less that 1\%, which implies that the
in-sample returns series is stationary. The results in the
autocorrelation(ACF) and partial autocorrelation(PACF) function
illustrated in figure 4 suggest that the second lag of both functions
might be significantly different from zero.\\
\includegraphics[width=1.00000\textwidth]{acf \& pacf plot.png}\\
\emph{Figure 4 : ACF and PACF plots for in-sample data}

Therefore, the candidate \emph{ARIMA(p,d,q)} models in Table 1 were fit
to the in-sample data and the AIC used to choose the model that best
fits the data. In Table 1, the candidate model with the lowest AIC is
the \(ARIMA(2,0,2)\). This model was thus selected as the ARIMA
forecasting model.\\
\includegraphics[width=1.00000\textwidth]{AIC.png}\\
\emph{Table 1: AIC for Candidate ARIMA models}

\subsection{Forecast Error Evaluation}\label{forecast-error-evaluation}

The results of model estimation and forecast evaluation using the
rolling window for h-step forecasts yielded forecasts errors with
roughly equal distributions. This is visible from the boxplot in Figure
5. The presence of outliers seems consistent across the different models
and forecastign horizons. Furthermore, the variation in the accuracy
across different forecasting horizons for each model are not observable
in Figure 5. The mean error and spread of both models across different
time horizons appears to be constant.
\includegraphics[width=1.05000\textwidth]{Forecast error boxplot.png}\\
\emph{Figure 5: Boxplot illustrating distribution of rolling forecast
errors for different time horizons, h}

\subsubsection{Forecast Error
Statistics}\label{forecast-error-statistics}

Table 2 summarizes the forecast error statistics for both the ARIMA and
Prophet model across the three different time horizons.\\
Across all forecast horizons, the ARIMA model appears to be
overestimating returns. This is evident from the negative mean error(ME)
across \(h = (1,5,20)\). On the contrary, the Prophet model
underestimates the forecasts for one-step and five-step ahead forecasts.
It then overstimates returns when forecasting twenty-steps ahead. The ME
results shown in Table 2 fail to clearly illustrate which model performs
better. Furthermore, the ME values are relatively small when compared
against other error statistics.

\includegraphics[width=1.00000\textwidth]{error stats.png}\\
\emph{Table 2 : Forecast error diagnostic statistics over different
forecasting horizons}

The Root Mean Square Error (RMSE) across different forecast horizons and
across both models are considerably close to each other in value
relative to the other error statistics. The same observation can be made
for the Mean Absolute Errors (MAE). These results suggest that the
models possibly have the same forecasting accuracy.

\subsubsection{Diebold-Mariano
Evaluation}\label{diebold-mariano-evaluation-1}

The observations made from Table 2 in section 5.2 suggest that the ARIMA
and Prophet models have equal predictive accuracy. These assertions can
be formally quantified through a Diebold-Mariano(DM) test which
evaluates whether the models have equal predictive accuracy against the
null hypothesis that one model has superior accuracy over the other.
Table 3 reports the results obtained when the rolling forecast errors
were formally compared against each other over different time
horizons.\\
\includegraphics[width=1.05000\textwidth]{dm results.png}\\
\emph{Table 3 Results of the Diebold-Mariano Hypothesis Test}\\
The P-value across all the time horizons for either linear and quadratic
loss functions suggest that we would fail to reject the null hypothesis
which implies that the predictive accuracy of the models could be equal
and the 95\% level. This observation validates the results obtained in
section 5.2.

\section{Discussion of Results}\label{discussion-of-results}

The results of the study show that there is no significant evidence to
conclude that Prophet generates superior forecasts for returns on the
FTSE/JSE Top40 Index when compared against those of an ARIMA model when
the Diebold-Mariano Test is used as a method of evaluation. These
results are consistent for daily, weekly and monthly forecasting
horizons.\\
Nevertheless, the study has flagged the possibility that, on average,
the ARIMA model tends to overestimate forecasted returns on the index
for the forecast horizons under consideration. This is contrary to
Prophet, that, on average, underestimates forecasted returns over daily
and weekly time horizons. This information could be useful for an
analyst engaged in short-term strategies since the choice of the
forecasting model can reflect a particular sentiment or risk appetite.
For example, if the analyst feels like prevailing returns are
underpriced by the market, they might choose to use an ARIMA model,
whereas an opposite sentiment could warrant the use pf Prophet.\\
Furthermore, it is worth noting that the merits of Prophet may have not
been fully exposed in this study due to the choice of the forecasting
horizon, i.e.~excluding periods of high volatility and uncertainity
during and after the 2008 Financial Crisis. Nonetheless, it is
reassuring that the model has performed as well as the ARIMA model. This
observation could create room for the application of Prophet in
circumstances where flexibility in the model formulation is desireable.

\section{Conclusion}\label{conclusion}

The challenge of building models that generate superior stock return
forecasts has led to widespread research in the field of quantitative
analysis for decades, but the ARIMA modelremains the standard model
against which to compare new innovations. This is because of its
relative simplicity and ability to evolve into more complex applications
e.g in the hybrid models considered in the literature review\\
Prophet has reduced the problem of estimating a time series model to a
cure fittig exercise by defining an additive, parametric function of
time. The functions' parameters are then estimated using Bayesian
estimation methods after which the tool automatically executes forecast
evaluation to optimize its predictive ability.\\
The results of this study suggest that, using a Diebold-Mariano
evaluation, ARIMA and Prophet do not have a significant difference in
predictive accuracy, even across time horizons of up to a month. The
implications of this are that analysts can then chosse which model to
use based on their needs without significantly compromising forecast
accuracy.

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-adebiyi2014comparison}{}
Adebiyi, Ayodele Ariyo, Aderemi Oluyinka Adewumi, and Charles Korede
Ayo. 2014. ``Comparison of Arima and Artificial Neural Networks Models
for Stock Price Prediction.'' \emph{Journal of Applied Mathematics}
2014. Hindawi Publishing Corporation.

\hypertarget{ref-box1970time}{}
Box, George Edward P, and Gwilym M Jenkins. 1970. ``Time Series
Analysis: Forecasting and Control, 1976.'' \emph{ISBN: 0-8162-1104-3}.

\hypertarget{ref-givoly1984quality}{}
Givoly, Dan, and Josef Lakonishok. 1984. ``The Quality of Analysts'
Forecasts of Earnings.'' \emph{Financial Analysts Journal}. JSTOR,
40--47.

\hypertarget{ref-guerard1989combining}{}
Guerard, John B. 1989. ``Combining Time-Series Model Forecasts and
Analysts' Forecasts for Superior Forecasts of Annual Earnings.''
\emph{Financial Analysts Journal}. JSTOR, 69--71.

\hypertarget{ref-RJHyndman2014}{}
Hyndman, Robert J. 2014. ``Variations on Rolling Forecasts.''
\url{https://robjhyndman.com/hyndsight/rolling-forecasts/}.

\hypertarget{ref-kihoro2004seasonal}{}
Kihoro, J, R Otieno, and C Wafula. 2004. ``Seasonal Time Series
Forecasting: A Comparative Study of Arima and Ann Models.'' \emph{AJST}
5 (2).

\hypertarget{ref-kriesel2007brief}{}
Kriesel, David. 2007. ``A Brief Introduction on Neural Networks.''
Citeseer.

\hypertarget{ref-lin2009short}{}
Lin, Xiaowei, Zehong Yang, and Yixu Song. 2009. ``Short-Term Stock Price
Prediction Based on Echo State Networks.'' \emph{Expert Systems with
Applications} 36 (3). Elsevier: 7313--7.

\hypertarget{ref-Mariano2000}{}
Mariano, Robert S. 2000. ``Testing Forecast Accuracy.'' University of
Pennsylvania; Course Notes.

\hypertarget{ref-moreno2011artificial}{}
Moreno, Juan José Montaño, Alfonso Palmer Pol, and Pilar Muñoz Gracia.
2011. ``Artificial Neural Networks Applied to Forecasting Time Series.''
\emph{Psicothema} 23 (2): 322--29.

\hypertarget{ref-Muten2014}{}
Mutendadzamera, S, and Farikayi K Mutasa. 2014. ``Forecasting Stock
Prices on the Zimbabwe Stock Exchange (Zse) Using Arima and Arch/Garch
Models.'' \emph{International Journal of Management Sciences} 3 (6):
419--32.

\hypertarget{ref-segaran2007programming}{}
Segaran, Toby. 2007. \emph{Programming Collective Intelligence: Building
Smart Web 2.0 Applications}. `` O'Reilly Media, Inc.''

\hypertarget{ref-tanikic2012artificial}{}
Tanikic, Dejan, and Vladimir Despotovic. 2012. ``Artificial Intelligence
Techniques for Modelling of Temperature in the Metal Cutting Process.''
In \emph{Metallurgy-Advances in Materials and Processes}. InTech.

\hypertarget{ref-taylor2017forecasting}{}
Taylor, Sean J, and Benjamin Letham. 2017. ``Forecasting at Scale.''

\hypertarget{ref-wang1996stock}{}
Wang, Jung-Hua, and Jia-Yann Leu. 1996. ``Stock Market Trend Prediction
Using Arima-Based Neural Networks.'' In \emph{Neural Networks, 1996.,
Ieee International Conference on}, 4:2160--5. IEEE.

\hypertarget{ref-zahedi2015application}{}
Zahedi, Javad, and Mohammad Mahdi Rounaghi. 2015. ``Application of
Artificial Neural Network Models and Principal Component Analysis Method
in Predicting Stock Prices on Tehran Stock Exchange.'' \emph{Physica A:
Statistical Mechanics and Its Applications} 438. Elsevier: 178--87.

\hypertarget{ref-zhang2009stock}{}
Zhang, Yudong, and Lenan Wu. 2009. ``Stock Market Prediction of S\&P 500
via Combination of Improved Bco Approach and Bp Neural Network.''
\emph{Expert Systems with Applications} 36 (5). Elsevier: 8849--54.

\newpage
\renewcommand{\baselinestretch}{1}
\nocite{*}
\bibliography{}

\end{document}

### 2017-09-14 template/Template.tex
final draft with all the relevant code and output data. Note than running the R code (rolling period code.R) takes some time and so the esults where stored as text files names "h step ahead results" for h=1,5,20

\documentclass[12pt,a4paper]{article}

\usepackage{url}
\usepackage{appendix}
\usepackage[british]{babel}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{graphicx}
\numberwithin{equation}{section}
\numberwithin{figure}{section}
\numberwithin{table}{section}

\usepackage[round]{natbib}
\bibliographystyle{natbib}
\def\bibsection{\section*{References}}

% Wrap around which gives all figures included the [H] command, or places it "here". This can be tedious to code in Rmarkdown.
\usepackage{float}
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
    \expandafter\origfigure\expandafter[H]
} {
    \endorigfigure
}

\let\origtable\table
\let\endorigtable\endtable
\renewenvironment{table}[1][2] {
    \expandafter\origtable\expandafter[H]
} {
    \endorigtable
}

\bibpunct[:]{(}{)}{;}{a}{,}{,}

\renewcommand{\baselinestretch}{1.5}

\begin{document}

\begin{titlepage}

\begin{center}
{\Huge \bf Helping You Write Academic Papers in R using Texevier}\\
\today\\
A Student (student number )\\
{\tt @uct.co.za}
\end{center}

\begin{abstract}
Abstract to be written here. The abstract should not be too long and
should provide the reader with a good understanding what you are writing
about. Academic papers are not like novels where you keep the reader in
suspense. To be effective in getting others to read your paper, be as
open and concise about your findings here as possible. Ideally, upon
reading your abstract, the reader should feel he / she must read your
paper in entirety.
\noindent
Keywords: literature, methodology, research,   results, structure, template.
\end{abstract}
\end{titlepage}

\pagenumbering{arabic}

\section{\texorpdfstring{Introduction
\label{Introduction}}{Introduction }}\label{introduction}

References are to be made as follows: Fama and French
(\protect\hyperlink{ref-fama1997}{1997}, 33) and Grinold and Kahn
(\protect\hyperlink{ref-grinold2000}{2000}) Such authors could also be
referenced in brackets (Grinold and Kahn
\protect\hyperlink{ref-grinold2000}{2000}) and together (Fama and French
\protect\hyperlink{ref-fama1997}{1997} \& Grinold and Kahn
(\protect\hyperlink{ref-grinold2000}{2000})). Source the reference code
from scholar.google.com by clicking on ``cite'' below article name. Then
select BibTeX at the bottom of the Cite window, and proceed to copy and
paste this code into your ref.bib file, located in the directory's Tex
folder. Open this file in Rstudio for ease of management, else open it
in your preferred Tex environment. Add and manage your article details
here for simplicity - once saved, it will self-adjust in your paper.

\begin{quote}
I suggest renaming the top line after @article, as done in the template
ref.bib file, to something more intuitive for you to remember. Do not
change the rest of the code. Also, be mindful of the fact that bib
references from google scholar may at times be incorrect. Reference
Latex forums for correct bibtex notation.
\end{quote}

To reference a section, you have to set a label using
``\textbackslash{}label'' in R, and then reference it in-text as e.g.:
section \ref{Data}.

Writing in Rmarkdown is surprizingly easy - see
\href{https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf}{this
website} cheatsheet for a summary on writing Rmd writing tips.

\section{\texorpdfstring{Data \label{Data}}{Data }}\label{data}

Discussion of data should be thorough with a table of statistics and
ideally a figure.

In your tempalte folder, you will find a Data and a Code folder. In
order to keep your data files neat, store all of them in your Data
folder. Also, I strongly suggest keeping this Rmd file for writing and
executing commands, not writing out long pieces of data-wrangling. In
the example below, I simply create a ggplot template for scatter plot
consistency. I suggest keeping all your data in a data folder.

\begin{figure}[H]

{\centering \includegraphics{Template_files/figure-latex/Figure1-1}

}

\caption{Caption Here \label{Figure1}}\label{fig:Figure1}
\end{figure}

To reference the plot above, add a ``\textbackslash{}label'' after the
caption in the chunk heading, as done above. Then reference the plot as
such: As can be seen, figure \ref{Figure1} is excellent. The nice thing
now is that it correctly numbers all your figures (and sections or
tables) and will update if it moves. The links are also dynamic.

I very strongly suggest using ggplot2 (ideally in combination with
dplyr) using the ggtheme package to change the themes of your figures.

Also note the information that I have placed above the chunks in the
code chunks for the figures. You can edit any of these easily - visit
the Rmarkdown webpage for more information.

Here follows another figure from built-in ggplot2 data:

\begin{figure}[H]

{\centering \includegraphics{Template_files/figure-latex/figure2-1}

}

\caption{Diamond Cut Plot \label{lit}}\label{fig:figure2}
\end{figure}

\section{Methodology}\label{methodology}

\subsection{Subsection}\label{subsection}

Ideally do not overuse subsections. It equates to bad writing.\footnote{This
  is an example of a footnote by the way. Something that should also not
  be overused.}

\subsection{Math section}\label{math-section}

Equations should be written as such:

\begin{align} 
\beta = \sum_{i = 1}^{\infty}\frac{\alpha^2}{\sigma_{t-1}^2} \label{eq1} \\ 
\int_{x = 1}^{\infty}x_{i} = 1 \notag
\end{align}

If you would like to see the equations as you type in Rmarkdown, use \$
symbols instead (see this for yourself by adjusted the equation):

\[
\beta = \sum_{i = 1}^{\infty}\frac{\alpha^2}{\sigma_{t-1}^2} \\ 
\int_{x = 1}^{\infty}x_{i} = 1
\]

\begin{align} 
    y_t &= c + B(L) y_{t-1} + e_t   \label{eq2}    \\ \notag 
    e_t &= H_t^{1/2}  z_t ; \quad z_t \sim  N(0,I_N) \quad \& \quad H_t = D_tR_tD_t \\ \notag
        D_t^2 &= {\sigma_{1,t}, \dots, \sigma_{N,t}}   \\ \notag
        \sigma_{i,t}^2 &= \gamma_i+\kappa_{i,t}  v_{i, t-1}^2 +\eta_i  \sigma_{i, t-1}^2, \quad \forall i \\ \notag
        R_{t, i, j} &= {diag(Q_{t, i, j}}^{-1}) . Q_{t, i, j} . diag(Q_{t, i, j}^{-1})  \\ \notag
        Q_{t, i, j} &= (1-\alpha-\beta)  \bar{Q} + \alpha  z_t  z_t'  + \beta  Q_{t, i, j} \notag
\end{align}

Noten that in \ref{eq2} I have aligned the equations by the equal signs.
I also want only one tag, and I create spaces using ``quads''.

See if you can figure out how to do complex math using the two examples
provided in \ref{eq1} and \ref{eq2}.

\section{Results}\label{results}

Tables can be included as follows. Use the \emph{xtable} (or kable)
package for tables. Table placement = H implies Latex tries to place the
table Here, and not on a new page (there are, however, very many ways to
skin this cat. Luckily there are many forums online!).

To reference calculations \textbf{in text}, \emph{do this:} From table
\ref{tab1} we see the average value of mpg is 20.98.

Including tables that span across pages, use e.g.:

\hfill

\section{Lists}\label{lists}

To add lists, simply using the following notation

\begin{itemize}
\item
  This is really simple

  \begin{itemize}
  \tightlist
  \item
    Just note the spaces here - writing in R you have to sometimes be
    pedantic about spaces\ldots{}
  \end{itemize}
\item
  Note that Rmarkdown notation removes the pain of defining
  \LaTeX environments!
\end{itemize}

\section{Conclusion}\label{conclusion}

I hope you find this template useful. Remember, stackoverflow is your
friend - use it to find answers to questions. Feel free to write me a
mail if you have any questions regarding the use of this package feel
free to email me. To cite this package, simply type
citation(``Texevier'') in Rstudio to get the citation for Katzke
(\protect\hyperlink{ref-Texevier}{2017}) (Note that uncited references
in your bibtex file will not be included in References).

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-fama1997}{}
Fama, Eugene F, and Kenneth R French. 1997. ``Industry Costs of
Equity.'' \emph{Journal of Financial Economics} 43 (2). Elsevier:
153--93.

\hypertarget{ref-grinold2000}{}
Grinold, Richard C, and Ronald N Kahn. 2000. ``Active Portfolio
Management.'' McGraw Hill New York, NY.

\hypertarget{ref-Texevier}{}
Katzke, N.F. 2017. \emph{Texevier: Package to Create Elsevier Templates
for Rmarkdown}. Stellenbosch, South Africa: Bureau for Economic
Research.

\newpage
\renewcommand{\baselinestretch}{1}
\nocite{*}
\bibliography{}

\end{document}

### 2017-09-14 Texevier-master/R/create_template.R

@@ -1,0 +1,115 @@
+#' @title create_template
+#' @description This function builds a template directory  with all the necessary files, as well as a rmarkdown script which is ready to edit.
+#' The result is an Elsevier ready document that conforms to most academic journal standards.
+#' In the directory folder provided, the user will have the tex files, as well as all relevant figure folders created in R. These should not be renamed or moved.
+#' Prior to submitting a paper, the user should zip the template_name.tex file, together with the figure files and Tex folder, and send it through to journal editors.
+#' Feel free to edit any of the settings in the template_name.tex file to conform to particular preferences (e.g. setting margins, etc.)
+#' @param directory A character string with the target folder directory This folder will be created if it does not already exist.
+#' @param template_name The folder, as well as the rmarkdown template, will be given this name.
+#' @param bib.location This can be a user specified location for a bib file. If omitted, the default bib file will be included. Note this command will create a renamed copy in the provided directory root.
+#' @param launch_template TRUE by default, this will launch the rmarkdown after creating the directory. The Working directory will also be changed.
+#' @return If launch_template is TRUE, the template Rmarkdown file will be opened in a new R session. This can then be used as a starting point for an academic research paper.
+#' @examples create_template(directory = "C:/Temp", template_name = "Project", launch_template = TRUE)
+#' @export
+
+create_template <- function (directory, template_name, bib.location, launch_template = TRUE, ShowPDFatLaunch = FALSE)
+{
+  sink(tempfile())
+  ifelse(!require(devtools, quietly = T), return(cat("Devtools library not found")),
+         FALSE)
+  tex.file.root <- system.file("Tex", package = "Texevier")
+  rmd.file.root <- system.file("Template", package = "Texevier")
+  data.file.root <- system.file("Data", package = "Texevier")
+  code.file.root <- system.file("Code", package = "Texevier")
+  if (missing(directory) | is.na(directory) == T) {
+    directory <- "C:/Template"
+  }
+  if (missing(template_name)) {
+    template_name <- "Template"
+  }
+  sink()
+  if (file.exists(directory)) {
+    answer <- readline(cat("--------------- \n PROMPT: \n \n Path: ",
+                           file.path(directory), " provided already exists. \n Proceed to create template in this folder? Type: Y or N....\n"))
+    while (!(answer %in% c("Y", "y", "N", "n"))) {
[…]

### 2017-09-14 rolling period code.R

@@ -1,0 +1,283 @@
+#2017 Honours Project: Draft Code           **
+#Author: Marvelous Mubenesha                **  
+#Student No. MBNMAR005                      **
+#*********************************************
+#*********************************************
+
+
+#--------- i) Import, clean and transform data-----------------------------------
+
+library(readxl)
+top40index <- read_excel("~/Honours_project/top40index.xlsx", 
+                         col_types = c("date", "numeric"))
+
+#Data cleaning : removing duplicates
+index <- top40index[1:1827,]  #warning message is because data is being duplicated repeatedly. Duplicates start after row 1827
+attach(index)
+#head(index) 
+
+#Calculating log returns from price index
+library(PerformanceAnalytics)
+log.returns <- CalculateReturns(as.ts(index[,2]), method = "log")[-1][61:1826]  #reduced data set because it was taking too long to run rolling window forecasts
+
+
+
+#-------- ii) Exploratiory Data Analysis --------------------------------------
+
+#Basic plots for price index and log returns ----------------------
+plot(index, type = "l",
+     main = "FTSE/JSE Top40 Index 07/2010 - 04/2017")
+
+par(mfrow = c(1,2))
+plot(log.returns, type = "l", main = "Log return on Price Index 07/2010 - 07/2017")
+plot(density(log.returns, adjust = 3), main = "Distribution of Log-Returns on Price Index 07/2010 - 04/2017", col = 1, lty = 1) 
+par(mfrow = c(1,1))
+
+#Outliers
+boxplot(returns, log.returns, names = c("returns","log returns"), main = "Boxplots of Returns and Log Returns on full sample")  #Comment : looks like there are some outliers eventhough outliers are 'relatively' clustered
+
+#Formal tests for Outliers in Time series --------------------------
+
+library(tsoutliers)  # conducts multiple tests for outliers #Reference Chen and Liu;s as authors : https://stats.stackexchange.com/questions/104882/detecting-outliers-in-time-series-ls-ao-tc-using-tsoutliers-package-in-r-how
+library(ggpubr)
+log.returns.outlier.detection <- tsoutliers::tso(as.ts(log.returns),types = c("AO","IO","LS","TC","SLS"), maxit.iloop = 10)
+log.return.outlier.detection  #Several outliers detected
[…]

### 2017-09-14 template/Draft_Paper.Rmd
corrected notation in draft paper(P_t)

@@
 $$  
-Where $P_{t+1}$ and $P_{t+1}$ represent the price of the index at time $t$ and $t+1$ respectively.
+Where $P_t$ and $P_{t+1}$ represent the price of the index at time $t$ and $t+1$ respectively.
 

### 2017-09-15 template/Draft_Paper.Rmd
Typo Edits

@@
 # The descriptions below should be self-explanatory
-title: "Comparing Facebook Prophet Forecasts to those of an ARIMA Model"
+title: "Comparing Facebook's Prophet Forecasting to an ARIMA Model: A Diebold-Mariano Evaluation of the Top40 Index"
 subtitle: "A Diebold-Mariano Evaluation of the FTSE/JSE Top40 Index"
@@
 Prophet employs intensive Bayesian modelling in two tiers.
-Purely through MAP parameter estimation, and in a quasi-Bayesian form through an automated forecast evaluation that enables analyst to improve the model is it falls short when compared to other models.
-This paper assess the predictive accuracy of Prophet forecasts to those of an ARIMA model that was identified using the Box-Jenkins methodology.
-This is achieved through a Diebold-Mariano evaluation of rolling period forecasts of returns on the FTSE/JSE Top40 Index across daily, weekly and monthly time horizons.
+Purely through MAP parameter estimation, and in a quasi-Bayesian form through an automated forecast evaluation that enables analyst to improve the model if it falls short when compared to other models.
+This paper assess the predictive accuracy of Prophets' forecasts to those of an ARIMA model that was identified using the Box-Jenkins methodology.
+This is achieved through a Diebold-Mariano evaluation of the rolling period forecast errors of returns on the FTSE/JSE Top40 Index across daily, weekly and monthly time horizons.
 The results of the study suggest that there is insufficient evidence to conclude that the forecasting models have unequal predictive ability within the scope of the JSE/Top40 Index.
@@
 
-In business, including that involveing the stock market, analyyst ay require a large number of time series forecasts to infomr decision making.The challenge with existing reliable approaches to forecasting time series data are that the models require a considerable background in statistics to build and adapt to prevailing market conditions when needed.
-Several authors have recommended automatic forecasting, which are algorithms that can automatically forecast a large number of univariate time series.
+In business, including that involving the stock market, analysts require a large number of time series forecasts to inform decision making.
+The challenge with existing reliable approaches to forecasting time series data are that the models require a considerable background in statistics to build and adapt to prevailing market conditions when needed.
+Several authors have recommended automatic forecasting, which refers to algorithms that can automatically forecast a large number of univariate time series.
 However, the forecasts generated by these methods have been found to be brittle as well as inflexible in allowing for prior information to be incorporated into the forecasting model [@taylor2017forecasting].
@@
 
-The objective of this paper is to compare the forecasts generated by Prophet - which uses automated, intensive Bayesian Modelling to 'forecast at scale'-, to those obtained by an ARIMA model -constructed through the Box-Jenkins methodology-,
+The objective of this paper is to compare the rolling period forecasts generated by Prophet (which uses automated, intensive Bayesian Modelling to 'forecast-at-scale'), to those obtained by an ARIMA model (constructed through the Box-Jenkins methodology),
 through a Diebold-Mariano evaluation of the JSE Top40 Index.
-The study is particularly useful because analysts from a wide background can be able to easily forecast stock price returns, automatically and by using their prior knowledge, without a wholesome understanding of the statistical intricacies involved.
+The study is particularly useful because analysts from a wide background can automatically forecast stock price returns, and incorporate their prior knowledge in the model , without a wholesome understanding of the statistical intricacies involved.
 
 The paper is organized as follows.
-A review of literature relating to stock price forecasting with ARIMA models an other alternative approaches is introduced to give the reader a background into the subject matter.
+A review of literature relating to stock price forecasting with ARIMA models and other alternative approaches is introduced to give the reader background information on research relating to the subject matter.
 Thereafter, theory relating to the mathematical formulation and estimation of both the ARIMA and Prophet models is included in sections 2.2 and 2.3 respectively, to build the readers understanding of what the models look like and how they operate.
-Information relating to the data that was used in this study is contained in section 3, after which, the methodology used to answer the research question is described.
-Section 5, contains the results obtained when the methodology was executed.
-whilst section 6 includes a discussion of the results and how they can be interpreted to answer the research question.
-Lastly, the conclusion is contain in section 7 and is aimed at highlighting the main findings of the study.
+Information relating to the data that was used in this study is contained in section 3, after which, the methodology used to answer the research question is described in section 4.
+Section 5, contains the results obtained when the methodology was executed, whilst section 6 includes a discussion of the results and how they can be interpreted to answer the research question.
+Lastly, the conclusion in section 7 highlights the main findings of the study and recommends further areas of research.
 
@@
 This will include both traditional approaches as well new methods such as automatic forecasting, machine learning techniques and quasi-Bayesian forms which incorporate analysts' prior information.
-The review of literature is concluded by emphasing how the findings of this paper can compliment existing research.
+The review of literature is concluded a brief introduction to the Diebold-Mariano evaluation and an outline of its suitability for the purpose of this study.
 The theoretical and mathematical formulation of an ARIMA forecasting model is  then briefly introduced.
@@
 An ANN consists of a sorted triple $(K, A, \omega)$.
-$K$ is a set of a multi-layered structure, V a paired set, that represents the value of the estimate in the $i^{th}$ node of the $k^{th}$ layer of the network.
-In Figure 1, the enlarged $a_{i,k} \in A$.
-Lastly, $\omega$ represents a function that defines the weights ($w_{i,r}$) of connections between neurons [@kriesel2007brief].
+$K$ is a set representing the multiple layers/levels of the network i.e.
+input, hidden and output levels.
+A is a set of pairs that represents the value of the index of the estimate in the $i^{th}$ node of the $k^{th}$ layer of the network.
+In the enlarged section of Figure 1, $a_{i,k} \in A$.
+Lastly, $\omega$ represents a function that defines the weights ($\omega_{i,r}$) of connections between neurons [@kriesel2007brief].
 Artificial Neural networks apply iterated optimization of model parameters across the network in the form of weights conditional on observed values to learn [@segaran2007programming].
@@
 
-The challenge of building models that generate superior stock return forecasts has led to widespread research in the field of quantitative analysis for decades, but the ARIMA modelremains the standard model against which to compare new innovations.
-This is because of its relative simplicity and ability to evolve into more complex applications e.g in the hybrid models considered in the literature review
-Prophet has reduced the problem of estimating a time series model to a cure fittig exercise by defining an additive, parametric function of time.
+The challenge of building models that generate superior stock return forecasts has led to widespread research in the field of quantitative analysis for decades, but the ARIMA model remains the standard against which to compare new innovations.
+This is because of its relative simplicity and ability to evolve into more complex applications e.g in the hybrid models considered in the literature review.
+Prophet has reduced the problem of estimating a time series model to a curve fittig exercise by defining an additive, parametric function of time.
 The functions' parameters are then estimated using Bayesian estimation methods after which the tool automatically executes forecast evaluation to optimize its predictive ability.
@@
 The implications of this are that analysts can then chosse which model to use based on their needs without significantly compromising forecast  accuracy.
+This study was confined to the analysis of predictive performance during general market conditions.
+Further research could explore the predictive ability of the models during different market conditions.For example, performance can be compared during bull and bear market conditions or during periods of high volatility.
+Furthermore, research can assess whether prophet would perform better on raw data that has not been transformed, since the model is built with growth and seasonality components and could thus predict better forecasts if these features are exploited.
 

### 2017-09-15 template/Draft_Paper.tex
corrected notation in draft paper(P_t)
Typo Edits

@@
 \begin{center}
-{\Huge \bf Comparing Facebook Prophet Forecasts to those of an ARIMA Model}\\
+{\Huge \bf Comparing Facebook's Prophet Forecasting to an ARIMA Model: A
+Diebold-Mariano Evaluation of the Top40 Index}\\
 \today\\
@@
 an automated forecast evaluation that enables analyst to improve the
-model is it falls short when compared to other models. This paper assess
-the predictive accuracy of Prophet forecasts to those of an ARIMA model
-that was identified using the Box-Jenkins methodology. This is achieved
-through a Diebold-Mariano evaluation of rolling period forecasts of
-returns on the FTSE/JSE Top40 Index across daily, weekly and monthly
-time horizons. The results of the study suggest that there is
-insufficient evidence to conclude that the forecasting models have
-unequal predictive ability within the scope of the JSE/Top40 Index.
+model if it falls short when compared to other models. This paper assess
+the predictive accuracy of Prophets' forecasts to those of an ARIMA
+model that was identified using the Box-Jenkins methodology. This is
+achieved through a Diebold-Mariano evaluation of the rolling period
+forecast errors of returns on the FTSE/JSE Top40 Index across daily,
+weekly and monthly time horizons. The results of the study suggest that
+there is insufficient evidence to conclude that the forecasting models
+have unequal predictive ability within the scope of the JSE/Top40 Index.
 \noindent
@@
 
-In business, including that involveing the stock market, analyyst ay
-require a large number of time series forecasts to infomr decision
-making.The challenge with existing reliable approaches to forecasting
-time series data are that the models require a considerable background
-in statistics to build and adapt to prevailing market conditions when
-needed. Several authors have recommended automatic forecasting, which
-are algorithms that can automatically forecast a large number of
-univariate time series. However, the forecasts generated by these
-methods have been found to be brittle as well as inflexible in allowing
-for prior information to be incorporated into the forecasting model
-(Taylor and Letham 2017). In February of 2017, Facebook's Core
-Datascience team open sourced ,Prophet, their time series forecasting
-tool whose main selling point is that it enables
-``forecasting-at-scale''. A term coined by the data science team which
-is defined as, ``an approach that allows a large number of analysts to
-forecast a large number and variety of business time series'' (Taylor
-and Letham 2017). It has the added advantage of being less `expensive'
-than other alternatives. In addition, it is specifically prefarable over
-ARIMA models because of its non-linearity, flexibility and ability to
-accomodate varying time intervals (Taylor and Letham 2017).
+In business, including that involving the stock market, analysts require
+a large number of time series forecasts to inform decision making. The
+challenge with existing reliable approaches to forecasting time series
+data are that the models require a considerable background in statistics
+to build and adapt to prevailing market conditions when needed. Several
+authors have recommended automatic forecasting, which refers to
+algorithms that can automatically forecast a large number of univariate
+time series. However, the forecasts generated by these methods have been
+found to be brittle as well as inflexible in allowing for prior
+information to be incorporated into the forecasting model (Taylor and
+Letham 2017). In February of 2017, Facebook's Core Datascience team open
+sourced ,Prophet, their time series forecasting tool whose main selling
+point is that it enables ``forecasting-at-scale''. A term coined by the
+data science team which is defined as, ``an approach that allows a large
+number of analysts to forecast a large number and variety of business
+time series'' (Taylor and Letham 2017). It has the added advantage of
+being less `expensive' than other alternatives. In addition, it is
+specifically prefarable over ARIMA models because of its non-linearity,
+flexibility and ability to accomodate varying time intervals (Taylor and
+Letham 2017).
 
-The objective of this paper is to compare the forecasts generated by
-Prophet - which uses automated, intensive Bayesian Modelling to
-`forecast at scale'-, to those obtained by an ARIMA model -constructed
-through the Box-Jenkins methodology-, through a Diebold-Mariano
-evaluation of the JSE Top40 Index. The study is particularly useful
-because analysts from a wide background can be able to easily forecast
-stock price returns, automatically and by using their prior knowledge,
-without a wholesome understanding of the statistical intricacies
-involved.
+The objective of this paper is to compare the rolling period forecasts
+generated by Prophet (which uses automated, intensive Bayesian Modelling
+to `forecast-at-scale'), to those obtained by an ARIMA model
+(constructed through the Box-Jenkins methodology), through a
+Diebold-Mariano evaluation of the JSE Top40 Index. The study is
+particularly useful because analysts from a wide background can
+automatically forecast stock price returns, and incorporate their prior
+knowledge in the model , without a wholesome understanding of the
+statistical intricacies involved.
 
 The paper is organized as follows. A review of literature relating to
-stock price forecasting with ARIMA models an other alternative
-approaches is introduced to give the reader a background into the
-subject matter. Thereafter, theory relating to the mathematical
-formulation and estimation of both the ARIMA and Prophet models is
-included in sections 2.2 and 2.3 respectively, to build the readers
-understanding of what the models look like and how they operate.
-Information relating to the data that was used in this study is
-contained in section 3, after which, the methodology used to answer the
-research question is described. Section 5, contains the results obtained
-when the methodology was executed. whilst section 6 includes a
-discussion of the results and how they can be interpreted to answer the
-research question. Lastly, the conclusion is contain in section 7 and is
-aimed at highlighting the main findings of the study.
+stock price forecasting with ARIMA models and other alternative
+approaches is introduced to give the reader background information on
+research relating to the subject matter. Thereafter, theory relating to
+the mathematical formulation and estimation of both the ARIMA and
+Prophet models is included in sections 2.2 and 2.3 respectively, to
+build the readers understanding of what the models look like and how
+they operate. Information relating to the data that was used in this
+study is contained in section 3, after which, the methodology used to
+answer the research question is described in section 4. Section 5,
+contains the results obtained when the methodology was executed, whilst
+section 6 includes a discussion of the results and how they can be
+interpreted to answer the research question. Lastly, the conclusion in
+section 7 highlights the main findings of the study and recommends
+further areas of research.
 
@@
 forms which incorporate analysts' prior information. The review of
-literature is concluded by emphasing how the findings of this paper can
-compliment existing research. The theoretical and mathematical
-formulation of an ARIMA forecasting model is then briefly introduced.
-Thereafter, the theoretical framework and mathematical formulation of
-Facebooks' Prophet forecasting tool is discussed. Specific attention is
-drawn to Prophets distinguising features to give the reader an
-understanding of the value that this forecasting technique can
-potentially add to stock price prediction. Lastly, the Diebold-Mariano
-Evaluation technique is introduced as a method of formally testing
-superior predictive accuracy between two forecasting models.
+literature is concluded a brief introduction to the Diebold-Mariano
+evaluation and an outline of its suitability for the purpose of this
+study. The theoretical and mathematical formulation of an ARIMA
+forecasting model is then briefly introduced. Thereafter, the
+theoretical framework and mathematical formulation of Facebooks' Prophet
+forecasting tool is discussed. Specific attention is drawn to Prophets
+distinguising features to give the reader an understanding of the value
+that this forecasting technique can potentially add to stock price
+prediction. Lastly, the Diebold-Mariano Evaluation technique is
+introduced as a method of formally testing superior predictive accuracy
+between two forecasting models.
 
@@
 and Despotovic 2012)\\
-An ANN consists of a sorted triple \((K, A, \omega)\). \(K\) is a set of
-a multi-layered structure, V a paired set, that represents the value of
-the estimate in the \(i^{th}\) node of the \(k^{th}\) layer of the
-network. In Figure 1, the enlarged \(a_{i,k} \in A\). Lastly, \(\omega\)
-represents a function that defines the weights (\(w_{i,r}\)) of
-connections between neurons (Kriesel 2007). Artificial Neural networks
-apply iterated optimization of model parameters across the network in
-the form of weights conditional on observed values to learn (Segaran
-2007). Several studies have compared ARIMA model forecasts to ANN and
-their conclusions are contradictory.
+An ANN consists of a sorted triple \((K, A, \omega)\). \(K\) is a set
+representing the multiple layers/levels of the network i.e.~input,
+hidden and output levels. A is a set of pairs that represents the value
+of the index of the estimate in the \(i^{th}\) node of the \(k^{th}\)
+layer of the network. In the enlarged section of Figure 1,
+\(a_{i,k} \in A\). Lastly, \(\omega\) represents a function that defines
+the weights (\(\omega_{i,r}\)) of connections between neurons (Kriesel
+2007). Artificial Neural networks apply iterated optimization of model
+parameters across the network in the form of weights conditional on
+observed values to learn (Segaran 2007). Several studies have compared
+ARIMA model forecasts to ANN and their conclusions are contradictory.
 
@@
 \]\\
-Where \(P_{t+1}\) and \(P_{t+1}\) represent the price of the index at
-time \(t\) and \(t+1\) respectively.
+Where \(P_t\) and \(P_{t+1}\) represent the price of the index at time
+\(t\) and \(t+1\) respectively.
 
@@
 forecasts has led to widespread research in the field of quantitative
-analysis for decades, but the ARIMA modelremains the standard model
-against which to compare new innovations. This is because of its
-relative simplicity and ability to evolve into more complex applications
-e.g in the hybrid models considered in the literature review\\
+analysis for decades, but the ARIMA model remains the standard against
+which to compare new innovations. This is because of its relative
+simplicity and ability to evolve into more complex applications e.g in
+the hybrid models considered in the literature review.\\
 Prophet has reduced the problem of estimating a time series model to a
-cure fittig exercise by defining an additive, parametric function of
+curve fittig exercise by defining an additive, parametric function of
 time. The functions' parameters are then estimated using Bayesian
@@
 use based on their needs without significantly compromising forecast
-accuracy.
+accuracy.\\
+This study was confined to the analysis of predictive performance during
+general market conditions. Further research could explore the predictive
+ability of the models during different market conditions.For example,
+performance can be compared during bull and bear market conditions or
+during periods of high volatility. Furthermore, research can assess
+whether prophet would perform better on raw data that has not been
+transformed, since the model is built with growth and seasonality
+components and could thus predict better forecasts if these features are
+exploited.
 

### 2017-09-15 template/Draft_Paper.Rmd
Typo edits

@@
 This will include both traditional approaches as well new methods such as automatic forecasting, machine learning techniques and quasi-Bayesian forms which incorporate analysts' prior information.
-The review of literature is concluded a brief introduction to the Diebold-Mariano evaluation and an outline of its suitability for the purpose of this study.
+The review of literature is concluded with a brief introduction to the Diebold-Mariano evaluation and an outline of its suitability for the purpose of this study.
 The theoretical and mathematical formulation of an ARIMA forecasting model is  then briefly introduced.

### 2017-09-15 template/Draft_Paper.tex
Typo edits

@@
 forms which incorporate analysts' prior information. The review of
-literature is concluded a brief introduction to the Diebold-Mariano
+literature is concluded with a brief introduction to the Diebold-Mariano
 evaluation and an outline of its suitability for the purpose of this

### 2017-09-15 template/Draft_Paper.Rmd
Final typo edits

@@
 Furthermore, these approaches tend to be inflexible with respect to incorporating prior information.
-Though this can be partially acheived by initializing parameter estimates for methods such as Artificial Neural Networks ,an analyst with no statistical training can have difficulty interpreting and thus modifying parameter estimates[@taylor2017forecasting].
+Though this can be partially achieved by initializing parameter estimates for some methods (such as ANNs) ,an analyst with no statistical training can have difficulty interpreting and thus modifying parameter estimates[@taylor2017forecasting].
 Prophet is an **automatic forecasting** tool that uses a Bayesian Generalized Additive model to generate time series forecasts.
@@
 
-Through this mechanism, a large numer of forecasts can be reliably generated **automatically** with the flexibility of enabling an analyst to modify the model in-the-loop during model specification.
+Through this mechanism, a large number of forecasts can be reliably generated **automatically** with the flexibility of enabling an analyst to modify the model in-the-loop during model specification.
 The automated forecasting procedure with an analyst-in-the-loop is illustrated in Figure 2.
@@
 
-The growth component of the model is generalized with two useful features that allow the carrying capacity and growth rate of the model to vary with time.
-The model enables an analyst to manually define when and how the growth rate changes at different change points [@taylor2017forecasting].
-This feature is particularly useful for forecasting stock price data since analysts are able to incorporate market movements that lead to the growth rate either decreasing, increasing or remaining constant.
-The seasonality component, $s(t)$ with periodicity $P$ and is given by,
+The growth component of the model has two useful features that allow the carrying capacity and growth rate of the model to vary with time.
+This enables an analyst to manually define when and how the growth rate changes at different change points [@taylor2017forecasting].
+This feature is particularly useful for forecasting stock price data since analysts are able to incorporate market movements that lead to the growth rate either decreasing, increasing or becoming constant.
+The seasonality component, $s(t)$ with periodicity $P$ is given by,
 $$s(t) = \sum_{n = -N}^{N}c_n \exp(j\frac{2\pi nt}{P})$$
@@
 $$
-Where, $\kappa$ is normally distributed scaling factor and  $D_j$ is a set of past and future dates where holidays or events occur.
+Where, $\kappa$ is a normally distributed scaling factor and  $D_j$ is a set of past and future dates where holidays or events occur.
 
-Prophet translates model parameter estimation into a curve-fitting exercise by employing Maximum Apriori Posterior (MAP) estimation, a quasi-Bayesian approach, to estimate the optimal parameter values for the forecasting model.
-MAP estimation finds the maximum posterior estimates for the parameters by using the likelihood function and an apriori distribution to find a posterior distribution which is optimized to return point estimates of the parameters [@taylor2017forecasting].
+Prophet translates model parameter estimation into a curve-fitting exercise by employing Maximum Apriori Posterior (MAP) estimation, a Bayesian approach to estimating optimal parameter values for the forecasting model.
+MAP estimation finds the maximum posterior estimates for the parameters by using the likelihood function and an apriori distribution to evaluate a posterior distribution which is then optimized to return point estimates of the parameters [@taylor2017forecasting].
 Thereafter, the package automatically evaluates the forecasts by comparing the Simulated Historical Forecast (SHF) errors of baseline forecasting methods (including ARIMA forecasts) to those of Prophet.
@@
 The aim is to compare the predictive accuracy of the two models under standard market conditions before other confounding factors can be closely analysed.
-Figure 33 shows the actual realisations of the price index which was sourced from Datastream.
+Figure 3 shows the actual realisations of the price index data which was sourced from Datastream.
 
@@
 
-The price index was transformation to log returns as a means of avoid spurious regression issues which can be caused by the existence of unit roots in stock prices.
+The price index was transformed to log returns as a means of avoiding spurious regression issues which can be caused by the existence of unit roots in stock price data.
 
@@
 The in-sample period comprises of 1400 trading days from 30 July 2012 to the 1 May 2016.
-The out_of_sample period comprises of the remaining 366 trading days from the dataset.
+The out-of-sample period comprises of the remaining 366 trading days from the dataset.
 
@@
 To identify what ARIMA model to use, the methodology as outlined in @box1970time was applied to the data.
-Firstly, stationarity conditions were verified formally by performing an augmented Dickey-Fuller test on the in-sample returns data.
+Firstly, stationarity conditions were verified formally by performing an Augmented Dickey-Fuller test on the in-sample returns data.
 The results of the test suggest the p_value is less that 1%, which implies that the in-sample returns series is stationary.
-The results in the autocorrelation(ACF) and partial autocorrelation(PACF) function illustrated in figure 4  suggest that the second lag of both functions might be significantly different from zero.
+The results in the autocorrelation(ACF) and partial autocorrelation(PACF) function illustrated in Figure 4  suggest that the second lag of both functions might be significantly different from zero.
 ![](acf & pacf plot.png){ width=100%}  
@@
 
-Therefore, the candidate *ARIMA(p,d,q)* models in Table 1 were fit to the in-sample data and the AIC used to choose the model that best fits the data.
-In Table 1, the candidate model with the lowest AIC is the $ARIMA(2,0,2)$.
+Observations from the above plots informed the choice of candidate *ARIMA(p,d,q)* models to fit to the data.
+The candidate models that were fit to the in-sample data are shown in Table 1 with their respective AIC values.
+The candidate model with the lowest AIC is the $ARIMA(2,0,2)$.
 This model was thus selected as the ARIMA forecasting model.
@@
 This is visible from the boxplot in Figure 5.
-The presence of outliers seems consistent across the different models and forecastign horizons.
-Furthermore, the variation in the accuracy across different forecasting horizons for each model are not observable in Figure 5.
+The presence of outliers seems consistent across the different models **and** forecasting horizons.
+Furthermore, variations in accuracy across different forecasting horizons *for each* model is not observable in Figure 5.
 The mean error and spread of both models across different time horizons appears to be constant.
@@
 This is evident from the negative mean error(ME) across $h = (1,5,20)$.
-On the contrary, the Prophet model underestimates the forecasts for one-step and five-step ahead forecasts.
+On the contrary, the Prophet model underestimates the forecasts for one-step and five-step ahead forecasting horizons.
 It then overstimates returns when forecasting twenty-steps ahead.
@@
 
-The observations made from Table 2 in section 5.2 suggest that the ARIMA and Prophet models have equal predictive accuracy.
-These assertions can be formally quantified through a Diebold-Mariano(DM) test which evaluates whether the models have equal predictive accuracy against the null hypothesis that one model has superior accuracy over the other.
+The observations made from Table 2 in section 5.2.1 suggest that the ARIMA and Prophet models have equal predictive accuracy.
+These assertions can be formally quantified through a Diebold-Mariano(DM) test which evaluates whether the models have equal predictive accuracy against the alternative hypothesis that one model has superior accuracy over the other.
 Table 3 reports the results obtained when the rolling forecast errors were formally compared against each other over different time horizons.
@@
 *Table 3 Results of the Diebold-Mariano Hypothesis Test*  
-The P-value across all the time horizons for either linear and quadratic loss functions suggest that we would fail to reject the null hypothesis which implies that the predictive accuracy of the models could be equal and the 95% level.
-This observation validates the results obtained in section 5.2.
+The p-value across all the time horizons for either linear and quadratic loss functions suggest that we would fail to reject the null hypothesis.
+This implies that the predictive accuracy of the models could be equal.
+These results formally validate the observations made obtained in section 5.2.1.
 
@@
 This information could be useful for an analyst engaged in short-term strategies since the choice of the forecasting model can reflect a particular sentiment or risk appetite.
-For example, if the analyst feels like prevailing returns are underpriced by the market, they might choose to use an ARIMA model, whereas an opposite sentiment could warrant the use pf Prophet.
-Furthermore, it is worth noting that the merits of Prophet may have not been fully exposed in this study due to the choice of the forecasting horizon, i.e.
-excluding periods of high volatility and uncertainity during and after the 2008 Financial Crisis.
+For example, if the analyst feels like prevailing returns are underpriced by the market, they might choose to use an ARIMA model, whereas an opposite sentiment could warrant the use of Prophet.
+Furthermore, it is worth noting that the merits of Prophet may have not been fully exposed in this study due to the use of transformed data and the choice of a forecasting horizon that excluded periods of high volatility.
 Nonetheless, it is reassuring that the model has performed as well as the ARIMA model.
@@
 The functions' parameters are then estimated using Bayesian estimation methods after which the tool automatically executes forecast evaluation to optimize its predictive ability.
-The results of this study suggest that, using a Diebold-Mariano evaluation, ARIMA and Prophet do not have a significant difference in predictive accuracy, even across time horizons of up to a month.
-The implications of this are that analysts can then chosse which model to use based on their needs without significantly compromising forecast  accuracy.
+The results of this study suggest that, using a Diebold-Mariano evaluation, the ARIMA and Prophet forecasting models do not have a significant difference in predictive accuracy across time horizons of up to a month.
+The implications of this are that analysts can then choose which model to use based on their needs without significantly compromising forecast  accuracy.
 This study was confined to the analysis of predictive performance during general market conditions.
-Further research could explore the predictive ability of the models during different market conditions.For example, performance can be compared during bull and bear market conditions or during periods of high volatility.
-Furthermore, research can assess whether prophet would perform better on raw data that has not been transformed, since the model is built with growth and seasonality components and could thus predict better forecasts if these features are exploited.
+Further research could explore the predictive ability of the models during different market conditions.
+For example, performance can be compared during bull and bear market conditions or during periods of high volatility.
+Additionally, further research can assess whether prophet would perform better on raw data that has not been transformed, since the model is built with growth and seasonality components.
+This use of which could lead to Prophet models that generate superior forecasts when compared against alternatives.
 

### 2017-09-15 template/Draft_Paper.tex
Final typo edits

@@
 approaches tend to be inflexible with respect to incorporating prior
-information. Though this can be partially acheived by initializing
-parameter estimates for methods such as Artificial Neural Networks ,an
-analyst with no statistical training can have difficulty interpreting
-and thus modifying parameter estimates(Taylor and Letham 2017).\\
+information. Though this can be partially achieved by initializing
+parameter estimates for some methods (such as ANNs) ,an analyst with no
+statistical training can have difficulty interpreting and thus modifying
+parameter estimates(Taylor and Letham 2017).\\
 Prophet is an \textbf{automatic forecasting} tool that uses a Bayesian
@@
 
-Through this mechanism, a large numer of forecasts can be reliably
+Through this mechanism, a large number of forecasts can be reliably
 generated \textbf{automatically} with the flexibility of enabling an
@@
 
-The growth component of the model is generalized with two useful
-features that allow the carrying capacity and growth rate of the model
-to vary with time. The model enables an analyst to manually define when
-and how the growth rate changes at different change points (Taylor and
-Letham 2017). This feature is particularly useful for forecasting stock
-price data since analysts are able to incorporate market movements that
-lead to the growth rate either decreasing, increasing or remaining
-constant.\\
-The seasonality component, \(s(t)\) with periodicity \(P\) and is given
-by, \[s(t) = \sum_{n = -N}^{N}c_n \exp(j\frac{2\pi nt}{P})\]
+The growth component of the model has two useful features that allow the
+carrying capacity and growth rate of the model to vary with time. This
+enables an analyst to manually define when and how the growth rate
+changes at different change points (Taylor and Letham 2017). This
+feature is particularly useful for forecasting stock price data since
+analysts are able to incorporate market movements that lead to the
+growth rate either decreasing, increasing or becoming constant.\\
+The seasonality component, \(s(t)\) with periodicity \(P\) is given by,
+\[s(t) = \sum_{n = -N}^{N}c_n \exp(j\frac{2\pi nt}{P})\]
 
@@
 \[ h(t) = \sum_{j = 1}^{L}\kappa_j \boldsymbol{1}(t \in D_j) \label{eqn5}
-\] Where, \(\kappa\) is normally distributed scaling factor and \(D_j\)
-is a set of past and future dates where holidays or events occur.
+\] Where, \(\kappa\) is a normally distributed scaling factor and
+\(D_j\) is a set of past and future dates where holidays or events
+occur.
 
@@
 exercise by employing Maximum Apriori Posterior (MAP) estimation, a
-quasi-Bayesian approach, to estimate the optimal parameter values for
-the forecasting model. MAP estimation finds the maximum posterior
-estimates for the parameters by using the likelihood function and an
-apriori distribution to find a posterior distribution which is optimized
-to return point estimates of the parameters (Taylor and Letham 2017).
-Thereafter, the package automatically evaluates the forecasts by
+Bayesian approach to estimating optimal parameter values for the
+forecasting model. MAP estimation finds the maximum posterior estimates
+for the parameters by using the likelihood function and an apriori
+distribution to evaluate a posterior distribution which is then
+optimized to return point estimates of the parameters (Taylor and Letham
+2017). Thereafter, the package automatically evaluates the forecasts by
 comparing the Simulated Historical Forecast (SHF) errors of baseline
@@
 accuracy of the two models under standard market conditions before other
-confounding factors can be closely analysed. Figure 33 shows the actual
-realisations of the price index which was sourced from Datastream.
+confounding factors can be closely analysed. Figure 3 shows the actual
+realisations of the price index data which was sourced from Datastream.
 
@@
 
-The price index was transformation to log returns as a means of avoid
+The price index was transformed to log returns as a means of avoiding
 spurious regression issues which can be caused by the existence of unit
-roots in stock prices.
+roots in stock price data.
 
@@
 in-sample and out-of-sample periods. The in-sample period comprises of
-1400 trading days from 30 July 2012 to the 1 May 2016. The
-out\_of\_sample period comprises of the remaining 366 trading days from
-the dataset.
+1400 trading days from 30 July 2012 to the 1 May 2016. The out-of-sample
+period comprises of the remaining 366 trading days from the dataset.
 
@@
 and Jenkins (1970) was applied to the data. Firstly, stationarity
-conditions were verified formally by performing an augmented
+conditions were verified formally by performing an Augmented
 Dickey-Fuller test on the in-sample returns data. The results of the
@@
 autocorrelation(ACF) and partial autocorrelation(PACF) function
-illustrated in figure 4 suggest that the second lag of both functions
+illustrated in Figure 4 suggest that the second lag of both functions
 might be significantly different from zero.\\
@@
 
-Therefore, the candidate \emph{ARIMA(p,d,q)} models in Table 1 were fit
-to the in-sample data and the AIC used to choose the model that best
-fits the data. In Table 1, the candidate model with the lowest AIC is
-the \(ARIMA(2,0,2)\). This model was thus selected as the ARIMA
-forecasting model.\\
+Observations from the above plots informed the choice of candidate
+\emph{ARIMA(p,d,q)} models to fit to the data. The candidate models that
+were fit to the in-sample data are shown in Table 1 with their
+respective AIC values. The candidate model with the lowest AIC is the
+\(ARIMA(2,0,2)\). This model was thus selected as the ARIMA forecasting
+model.\\
 \includegraphics[width=1.00000\textwidth]{AIC.png}\\
@@
 5. The presence of outliers seems consistent across the different models
-and forecastign horizons. Furthermore, the variation in the accuracy
-across different forecasting horizons for each model are not observable
-in Figure 5. The mean error and spread of both models across different
-time horizons appears to be constant.
+\textbf{and} forecasting horizons. Furthermore, variations in accuracy
+across different forecasting horizons \emph{for each} model is not
+observable in Figure 5. The mean error and spread of both models across
+different time horizons appears to be constant.
 \includegraphics[width=1.05000\textwidth]{Forecast error boxplot.png}\\
@@
 across \(h = (1,5,20)\). On the contrary, the Prophet model
-underestimates the forecasts for one-step and five-step ahead forecasts.
-It then overstimates returns when forecasting twenty-steps ahead. The ME
-results shown in Table 2 fail to clearly illustrate which model performs
-better. Furthermore, the ME values are relatively small when compared
-against other error statistics.
+underestimates the forecasts for one-step and five-step ahead
+forecasting horizons. It then overstimates returns when forecasting
+twenty-steps ahead. The ME results shown in Table 2 fail to clearly
+illustrate which model performs better. Furthermore, the ME values are
+relatively small when compared against other error statistics.
 
@@
 
-The observations made from Table 2 in section 5.2 suggest that the ARIMA
-and Prophet models have equal predictive accuracy. These assertions can
-be formally quantified through a Diebold-Mariano(DM) test which
-evaluates whether the models have equal predictive accuracy against the
-null hypothesis that one model has superior accuracy over the other.
-Table 3 reports the results obtained when the rolling forecast errors
-were formally compared against each other over different time
-horizons.\\
+The observations made from Table 2 in section 5.2.1 suggest that the
+ARIMA and Prophet models have equal predictive accuracy. These
+assertions can be formally quantified through a Diebold-Mariano(DM) test
+which evaluates whether the models have equal predictive accuracy
+against the alternative hypothesis that one model has superior accuracy
+over the other. Table 3 reports the results obtained when the rolling
+forecast errors were formally compared against each other over different
+time horizons.\\
 \includegraphics[width=1.05000\textwidth]{dm results.png}\\
 \emph{Table 3 Results of the Diebold-Mariano Hypothesis Test}\\
-The P-value across all the time horizons for either linear and quadratic
-loss functions suggest that we would fail to reject the null hypothesis
-which implies that the predictive accuracy of the models could be equal
-and the 95\% level. This observation validates the results obtained in
-section 5.2.
+The p-value across all the time horizons for either linear and quadratic
+loss functions suggest that we would fail to reject the null hypothesis.
+This implies that the predictive accuracy of the models could be equal.
+These results formally validate the observations made obtained in
+section 5.2.1.
 
@@
 underpriced by the market, they might choose to use an ARIMA model,
-whereas an opposite sentiment could warrant the use pf Prophet.\\
+whereas an opposite sentiment could warrant the use of Prophet.\\
 Furthermore, it is worth noting that the merits of Prophet may have not
-been fully exposed in this study due to the choice of the forecasting
-horizon, i.e.~excluding periods of high volatility and uncertainity
-during and after the 2008 Financial Crisis. Nonetheless, it is
-reassuring that the model has performed as well as the ARIMA model. This
-observation could create room for the application of Prophet in
-circumstances where flexibility in the model formulation is desireable.
+been fully exposed in this study due to the use of transformed data and
+the choice of a forecasting horizon that excluded periods of high
+volatility. Nonetheless, it is reassuring that the model has performed
+as well as the ARIMA model. This observation could create room for the
+application of Prophet in circumstances where flexibility in the model
+formulation is desireable.
 
@@
 The results of this study suggest that, using a Diebold-Mariano
-evaluation, ARIMA and Prophet do not have a significant difference in
-predictive accuracy, even across time horizons of up to a month. The
-implications of this are that analysts can then chosse which model to
-use based on their needs without significantly compromising forecast
-accuracy.\\
+evaluation, the ARIMA and Prophet forecasting models do not have a
+significant difference in predictive accuracy across time horizons of up
+to a month. The implications of this are that analysts can then choose
+which model to use based on their needs without significantly
+compromising forecast accuracy.\\
 This study was confined to the analysis of predictive performance during
 general market conditions. Further research could explore the predictive
-ability of the models during different market conditions.For example,
+ability of the models during different market conditions. For example,
 performance can be compared during bull and bear market conditions or
-during periods of high volatility. Furthermore, research can assess
-whether prophet would perform better on raw data that has not been
-transformed, since the model is built with growth and seasonality
-components and could thus predict better forecasts if these features are
-exploited.
+during periods of high volatility. Additionally, further research can
+assess whether prophet would perform better on raw data that has not
+been transformed, since the model is built with growth and seasonality
+components. This use of which could lead to Prophet models that generate
+superior forecasts when compared against alternatives.
 

### 2017-09-17 template/Draft_Paper.Rmd
last ush didnt connect to internet and i only saw that now

@@
 input, hidden and output levels.
-A is a set of pairs that represents the value of the index of the estimate in the $i^{th}$ node of the $k^{th}$ layer of the network.
+A is a set of pairs that represents index for the estimate in the $i^{th}$ node of the $k^{th}$ layer of the network.
 In the enlarged section of Figure 1, $a_{i,k} \in A$.
-Lastly, $\omega$ represents a function that defines the weights ($\omega_{i,r}$) of connections between neurons [@kriesel2007brief].
+Lastly, $\omega$ represents a function that defines the weights ($\omega_{i,r}$) of connections between neurons/nodes [@kriesel2007brief].
 Artificial Neural networks apply iterated optimization of model parameters across the network in the form of weights conditional on observed values to learn [@segaran2007programming].

### 2017-10-14 template/Draft_Paper.Rmd
Draft paper edits: Data plot fixed, table 1 and 2 changed to markdown tables, Some data plots were printed directly from markdown (rest of the code will be added to markdown doc later)

@@
 
-###Quasi-Bayesian methodss incorporating Analysts' knowlegde
+###Quasi-Bayesian methods incorporating Analysts' knowledge
 
-The stock price forecasting approaches considered so far are purely apply technical data analysis methods to generate stock price forecasting models.
+The stock price forecasting approaches considered so far purely apply technical data analysis methods to generate stock price forecasting models.
 However, studies that have compared pure technical approaches to those that incorporate the analysts knowledge in a quasi-Bayesian form have been shown to be superior at predicting market prices [@givoly1984quality \& @guerard1989combining].
@@
 ###Diebold-Mariano Evaluation
-As is evident from the previous sections, the modelling approach and assumptions of Prophet forecasting-at-scale differ to those of an ARIMA model.
+As is evident from the previous sections, the modelling approach and assumptions of Prophet differ to those of an ARIMA model.
 Since the two methods differ, the Diebold-Mariano evaluation (DM Test) is preferred as a means of testing forecasting accuracy.
@@
 where:  
-$\bullet \phi_i$ and $\phi_j$ are the parameters to be estimated;  
+$\bullet \phi_i$ and $\psi_j$ are the parameters to be estimated;  
 $\bullet r_{t-k}$ is the $k^{th}$ lagged return;   

### 2017-10-14 template/Draft_Paper.tex
Draft paper edits: Data plot fixed, table 1 and 2 changed to markdown tables, Some data plots were printed directly from markdown (rest of the code will be added to markdown doc later)

@@
 representing the multiple layers/levels of the network i.e.~input,
-hidden and output levels. A is a set of pairs that represents the value
-of the index of the estimate in the \(i^{th}\) node of the \(k^{th}\)
-layer of the network. In the enlarged section of Figure 1,
-\(a_{i,k} \in A\). Lastly, \(\omega\) represents a function that defines
-the weights (\(\omega_{i,r}\)) of connections between neurons (Kriesel
-2007). Artificial Neural networks apply iterated optimization of model
+hidden and output levels. A is a set of pairs that represents index for
+the estimate in the \(i^{th}\) node of the \(k^{th}\) layer of the
+network. In the enlarged section of Figure 1, \(a_{i,k} \in A\). Lastly,
+\(\omega\) represents a function that defines the weights
+(\(\omega_{i,r}\)) of connections between neurons/nodes (Kriesel 2007).
+Artificial Neural networks apply iterated optimization of model
 parameters across the network in the form of weights conditional on
@@
 
-\subsubsection{Quasi-Bayesian methodss incorporating Analysts'
-knowlegde}\label{quasi-bayesian-methodss-incorporating-analysts-knowlegde}
+\subsubsection{Quasi-Bayesian methods incorporating Analysts'
+knowledge}\label{quasi-bayesian-methods-incorporating-analysts-knowledge}
 
-The stock price forecasting approaches considered so far are purely
-apply technical data analysis methods to generate stock price
-forecasting models. However, studies that have compared pure technical
-approaches to those that incorporate the analysts knowledge in a
-quasi-Bayesian form have been shown to be superior at predicting market
-prices (Givoly and Lakonishok 1984 \& Guerard (1989)). The first
-instance of combining time-series model forecasts and an analysts
-forecasts to obtain superior forecasts of a stocks annual earnings can
-be dated to 1989. Guerard (1989) used an additive model to combine
-consensus security analyst forecasts from the S\&P Annual Earnings
-forecaster and annual earnings forecasts generated by an ARIMA model
-with a constant. The combined model that was estimated using ordinary
-least squares reduced the mean square error of the time series and
-analysts forecasts from 1.28 and 1.27, respectively to 1.04. These
-results suggest that analysts can substantially reduce forecasting
-errors by combining both approaches (Guerard 1989). Zahedi and Rounaghi
-(2015) applied the same overarching principle when they used principle
-component analysis to determine an appropriate input variable, which
-they then used to train an ANN in an attempt to predict stock prices on
-the Tehran Stock Exchange. Their model yielded acceptable forecasting
-errors and performed better than the pure ANN which further reiterates
-the potential of incorporating an analysts knowledge to improve stock
-price forecasts, in both linear and nonlinear models.
+The stock price forecasting approaches considered so far purely apply
+technical data analysis methods to generate stock price forecasting
+models. However, studies that have compared pure technical approaches to
+those that incorporate the analysts knowledge in a quasi-Bayesian form
+have been shown to be superior at predicting market prices (Givoly and
+Lakonishok 1984 \& Guerard (1989)). The first instance of combining
+time-series model forecasts and an analysts forecasts to obtain superior
+forecasts of a stocks annual earnings can be dated to 1989. Guerard
+(1989) used an additive model to combine consensus security analyst
+forecasts from the S\&P Annual Earnings forecaster and annual earnings
+forecasts generated by an ARIMA model with a constant. The combined
+model that was estimated using ordinary least squares reduced the mean
+square error of the time series and analysts forecasts from 1.28 and
+1.27, respectively to 1.04. These results suggest that analysts can
+substantially reduce forecasting errors by combining both approaches
+(Guerard 1989). Zahedi and Rounaghi (2015) applied the same overarching
+principle when they used principle component analysis to determine an
+appropriate input variable, which they then used to train an ANN in an
+attempt to predict stock prices on the Tehran Stock Exchange. Their
+model yielded acceptable forecasting errors and performed better than
+the pure ANN which further reiterates the potential of incorporating an
+analysts knowledge to improve stock price forecasts, in both linear and
+nonlinear models.
 
@@
 As is evident from the previous sections, the modelling approach and
-assumptions of Prophet forecasting-at-scale differ to those of an ARIMA
-model. Since the two methods differ, the Diebold-Mariano evaluation (DM
-Test) is preferred as a means of testing forecasting accuracy. This is
-because the DM Test is a model free test of forecasting accuracy. It is
-applicable to a wide range of situations from multi-period forecasts,
-non-Gaussian forecast residuals, non-quadratic loss functions and even
-serially correlated data(Mariano 2000).
+assumptions of Prophet differ to those of an ARIMA model. Since the two
+methods differ, the Diebold-Mariano evaluation (DM Test) is preferred as
+a means of testing forecasting accuracy. This is because the DM Test is
+a model free test of forecasting accuracy. It is applicable to a wide
+range of situations from multi-period forecasts, non-Gaussian forecast
+residuals, non-quadratic loss functions and even serially correlated
+data(Mariano 2000).
 
@@
 where:\\
-\(\bullet \phi_i\) and \(\phi_j\) are the parameters to be estimated;\\
+\(\bullet \phi_i\) and \(\psi_j\) are the parameters to be estimated;\\
 \(\bullet r_{t-k}\) is the \(k^{th}\) lagged return;\\
@@
 The data set used throughout this paper comprises of the daily closing
-prices of the equally-weighted FTSE/JSE Top40 price index for the five
-year period from 31/07/2012 to 30/07/2017. This equates to a total of
-1766 trading days excluding weekends and holidays. This period was
-chosen for its recency and to avoid exposing the results to the noise
-associated with the high volatility experienced in the stock market
-during the 2008 financial crisis. The aim is to compare the predictive
-accuracy of the two models under standard market conditions before other
-confounding factors can be closely analysed. Figure 3 shows the actual
-realisations of the price index data which was sourced from Datastream.
+prices of the FTSE/JSE Top40 price index for the five year period from
+31/07/2012 to 31/07/2017. This equates to a total of 1305 trading days
+excluding weekends and holidays. This period was chosen for its recency
+and to avoid exposing the results to the noise associated with the high
+volatility experienced in the stock market during the 2008 financial
+crisis. The aim is to compare the predictive accuracy of the two models
+under standard market conditions before other confounding factors can be
+closely analysed. Figure 3 shows the actual realisations of the price
+index data which was sourced from Datastream.
 
-\includegraphics[width=1.00000\textwidth]{plot of Price index.png}\\
+\includegraphics{Draft_Paper_files/figure-latex/unnamed-chunk-1-1.pdf}
 \emph{Figure 3: Realisation of the FTSE/JSE Top40 Price Index over the
@@
 
-In this paper, the \textbf{log return}(which will be used
+In this paper, the \textbf{log return} (which will be used
 interchangeably with \textbf{return}), over the period \([t,t+1]\) is
@@
 \(ARIMA(2,0,2)\). This model was thus selected as the ARIMA forecasting
-model.\\
-\includegraphics[width=1.00000\textwidth]{AIC.png}\\
-\emph{Table 1: AIC for Candidate ARIMA models}
+model.
+
+\begin{table}[H]
+\centering
+\begin{tabular}{ll}
+  \hline
+ARIMA(p,d,q) model & AIC \\ 
+  \hline
+(0,0,0) & -8804 \\ 
+  (2,0,0) & -8808 \\ 
+  (0,0,2) & -8808 \\ 
+  (2,0,2) & -8817 \\ 
+   \hline
+\end{tabular}
+\caption{AIC for Candidate ARIMA models} 
+\end{table}
 
@@
 relatively small when compared against other error statistics.
+
+\begin{table}[H]
+\centering
+\begin{tabular}{lllll}
+  \hline
+Forecast Horizon, h & Forecast Model &  ME  & RMSE & MAE \\ 
+  \hline
+1 & ARIMA & -0.0004110 & 0.0103 & 0.007718 \\ 
+    & Prophet & 0.0000176 & 0.0104 & 0.007749 \\ 
+  5 & ARIMA & -0.0003635 & 0.0103 & 0.007692 \\ 
+    & Prophet & 0.004445 & 0.0103 & 0.007709 \\ 
+  20 & ARIMA & -0.0001769 & 0.0102 & 0.007649 \\ 
+    & Prophet & -0.0000406 & 0.0102 & 0.007660 \\ 
+   \hline
+\end{tabular}
+\caption{Forecast error diagnostic statistics over different forecasting horizons} 
+\end{table}
 

### 2017-10-14 rolling period code.R

@@ -6,35 +6,17 @@
 
 
 #--------- i) Import, clean and transform data-----------------------------------
-
+library(dplyr)
 library(readxl)
-top40index <- read_excel("~/Honours_project/top40index.xlsx", 
+top40index <- read_excel("~/Honours_project/Draft-Paper/top40index.xlsx", 
                          col_types = c("date", "numeric"))
-
-#Data cleaning : removing duplicates
-index <- top40index[1:1827,]  #warning message is because data is being duplicated repeatedly. Duplicates start after row 1827
-attach(index)
-#head(index) 
-
-#Calculating log returns from price index
-library(PerformanceAnalytics)
-log.returns <- CalculateReturns(as.ts(index[,2]), method = "log")[-1][61:1826]  #reduced data set because it was taking too long to run rolling window forecasts
+index <- top40index%>%distinct(date, .keep_all = TRUE) #Removing duplicate dates
 
 
-
-#-------- ii) Exploratiory Data Analysis --------------------------------------
-
-#Basic plots for price index and log returns ----------------------
-plot(index, type = "l",
-     main = "FTSE/JSE Top40 Index 07/2010 - 04/2017")
-
-par(mfrow = c(1,2))
-plot(log.returns, type = "l", main = "Log return on Price Index 07/2010 - 07/2017")
-plot(density(log.returns, adjust = 3), main = "Distribution of Log-Returns on Price Index 07/2010 - 04/2017", col = 1, lty = 1) 
-par(mfrow = c(1,1))
-
-#Outliers
-boxplot(returns, log.returns, names = c("returns","log returns"), main = "Boxplots of Returns and Log Returns on full sample")  #Comment : looks like there are some outliers eventhough outliers are 'relatively' clustered
+# Outliers
+boxplot(c(index[,2]), names = c("price index"), main = "Boxplot of Price Index")  #Comment : looks like there are some outliers eventhough outliers are 'relatively' clustered
+log.returns <- CalculateReturns(as.ts(top40index[,2]), method = "log")[-1] 
+boxplot(log.returns, names = c("log returns"), main = "Boxplot of Log Returns on the Price Index")
 
 #Formal tests for Outliers in Time series --------------------------
 
@@ -42,15 +24,16 @@ library(tsoutliers)  # conducts multiple tests for outliers #Reference Chen and 
 library(ggpubr)
[…]

### 2017-10-14 template/Draft_Paper.Rmd
Tables in markdown, some, but not all plots in markdown, fixed issue with dates on figure 1, in-sample data with dplyr

@@
 
-The data set used throughout this paper comprises of the daily closing prices of the equally-weighted FTSE/JSE Top40 price index for the five year period from 31/07/2012 to 30/07/2017.
-This equates to a total of 1766 trading days excluding weekends and holidays.
+The data set used throughout this paper comprises of the daily closing prices of the FTSE/JSE Top40 price index for the five year period from 31/07/2012 to 31/07/2017.
+This equates to a total of 1305 trading days excluding weekends and holidays.
 This period was chosen for its recency and to avoid exposing the results to the noise associated with the high volatility experienced in the stock market during the 2008 financial crisis.
@@
 
-![Figure 2](plot of Price index.png){ width=100% }  
+```{r, echo=FALSE}
+#Import, transform and plot data
+library(dplyr)
+library(readxl)
+top40index <- read_excel("~/Honours_project/Draft-Paper/top40index.xlsx", 
+                         col_types = c("date", "numeric"))
+index <- top40index%>%distinct(date, .keep_all = TRUE) #Removing duplicate dates
+
+plot(index, type = "l",
+     main = "FTSE/JSE Top40 Index 07/2012 - 07/2017")
+```
 *Figure 3: Realisation of the FTSE/JSE Top40 Price Index over the period of study*
@@
 
-In this paper, the **log return**(which will be used interchangeably with **return**), over the period $[t,t+1]$ is denoted by $r_t$, and was calculated by evaluating;  
+In this paper, the **log return** (which will be used interchangeably with **return**), over the period $[t,t+1]$ is denoted by $r_t$, and was calculated by evaluating;  
 $$
@@
 The out-of-sample period comprises of the remaining 366 trading days from the dataset.
+
+```{r, eval=FALSE}
+# Outliers
+boxplot(c(index[,2]), names = c("price index"), main = "Boxplot of Price Index")  #Comment : looks like there are some outliers eventhough outliers are 'relatively' clustered
+log.returns <- CalculateReturns(as.ts(top40index[,2]), method = "log")[-1] 
+boxplot(log.returns, names = c("log returns"), main = "Boxplot of Log Returns on the Price Index")
+
+#Formal tests for Outliers in Time series --------------------------
+
+library(tsoutliers)  # conducts multiple tests for outliers #Reference Chen and Liu;s as authors : https://stats.stackexchange.com/questions/104882/detecting-outliers-in-time-series-ls-ao-tc-using-tsoutliers-package-in-r-how
+library(ggpubr)
+log.returns.outlier.detection <- tsoutliers::tso(as.ts(log.returns),types = c("AO","IO","LS","TC","SLS"), maxit.iloop = 10)
+log.return.outlier.detection  #Several outliers detected
+#Further investigation shows that these outliers occur when the return is large in abs value, because this is due to natural market activity and the outliers do not reflect in the boxplot of the price index, the outliers have been kept in the data.
+See : log.returns[c(299,302,348,1144,1156,1540)]
+
+```
 
@@
 This model was thus selected as the ARIMA forecasting model.
-![](AIC.png){ width=100% }  
-*Table 1: AIC for Candidate ARIMA models*
 
+```{r ShortTable, results = 'asis'}
+
+library(xtable)
+
+candidate <- c("(0,0,0)","(2,0,0)","(0,0,2)","(2,0,2)")
+aic.values <- c(-8804,-8808,-8808,-8817)
+
+aic.mat <- matrix(data = c(candidate,aic.values),ncol = 2)
+dimnames(aic.mat) <- list(c(" "," "," "," "),   #row names
+                          c("ARIMA(p,d,q) model","AIC"))  #column names
+data = aic.mat %>% tbl_df()
+
+table <- xtable(data, caption = "AIC for Candidate ARIMA models")
+  print.xtable(table, 
+             # tabular.environment = "longtable",
+             floating = TRUE,
+             table.placement = 'H', 
+             # scalebox = 0.3, 
+             comment = FALSE,
+             include.rownames = getOption("xtable.include.rownames", FALSE), #exclude row number
+             caption.placement = 'bottom'
+             )
+
+```
 ## Forecast Error Evaluation
@@
 
-![ErrorStats](error stats.png){ width=100% }  
-*Table  2 : Forecast error diagnostic statistics over different forecasting horizons*
+```{r ForecastErrorStatisticsTable, results = 'asis'}
+
+library(xtable)
+
+c1 <- c("1"," ","5"," ","20"," ")
+c2 <- rep(c("ARIMA","Prophet"),3)
+c3 <- c('-0.0004110','0.0000176','-0.0003635','0.004445','-0.0001769','-0.0000406')
+c4 <- c(0.0103,0.0104,0.0103,0.0103,0.0102,0.0102)
+c5 <- c(0.007718,0.007749,0.007692,0.007709,0.007649,"0.007660")
+
+error.stats.mat <- matrix(data = c(c1,c2,c3,c4,c5),ncol = 5)
+dimnames(error.stats.mat) <- list(c(" "," "," "," "," "," "),   #row names
+                          c("Forecast Horizon, h","Forecast Model"," ME ","RMSE","MAE"))  #column names
+err.stat.data = error.stats.mat %>% tbl_df()
+
+error.table <- xtable(err.stat.data, caption = "Forecast error diagnostic statistics over different forecasting horizons")
+  print.xtable(error.table, 
+             # tabular.environment = "longtable",
+             floating = TRUE,
+             table.placement = 'H', 
+             # scalebox = 0.3, 
+             comment = FALSE,
+             include.rownames = getOption("xtable.include.rownames", FALSE), #exclude row number
+             caption.placement = 'bottom'
+             )
+
+```
 

### 2017-10-14 template/Draft_Paper.Rmd
All tables changed to r markdon tables

@@
 
-```{r ShortTable, results = 'asis'}
+```{r AIC Table for candidate ARIMA Models, results = 'asis'}
 
@@
 
-```{r ForecastErrorStatisticsTable, results = 'asis'}
+```{r Forecast Error Statistics Table, results = 'asis'}
 
@@
 Table 3 reports the results obtained when the rolling forecast errors were formally compared against each other over different time horizons.
-![](dm results.png){ width=105%}  
-*Table 3 Results of the Diebold-Mariano Hypothesis Test*  
+
+```{r Table For Diebold Mariano Evaluation results, results = 'asis'}
+
+library(xtable)
+
+col1 <- c("1"," ","5"," ","20"," ")
+col2 <- rep(c("Linear","Quadratic"),3)
+col3 <- c(0.553,0.128,0.691,0.334,'0.660',0.296)
+
+dm.mat <- matrix(data = c(col1,col2,col3),ncol = 3)
+dimnames(dm.mat) <- list(c(" "," "," "," "," "," "),   #row names
+                          c("Forecast Horizon, h","Loss Function","P Value"))  #column names
+dm.data = dm.mat %>% tbl_df()
+
+dm.table <- xtable(dm.data, caption = "Results of the Diebold-Mariano Hypothesis Test for different loss functions")
+  print.xtable(dm.table, 
+             # tabular.environment = "longtable",
+             floating = TRUE,
+             table.placement = 'H', 
+             # scalebox = 0.3, 
+             comment = FALSE,
+             include.rownames = getOption("xtable.include.rownames", FALSE), #exclude row number
+             caption.placement = 'bottom'
+             )
+
+```
+
 The p-value across all the time horizons for either linear and quadratic loss functions suggest that we would fail to reject the null hypothesis.

### 2017-10-14 template/Draft_Paper.tex
Tables in markdown, some, but not all plots in markdown, fixed issue with dates on figure 1, in-sample data with dplyr
All tables changed to r markdon tables

@@
 
-\includegraphics[width=1.00000\textwidth]{error stats.png}\\
-\emph{Table 2 : Forecast error diagnostic statistics over different
-forecasting horizons}
-
 The Root Mean Square Error (RMSE) across different forecast horizons and
@@
 forecast errors were formally compared against each other over different
-time horizons.\\
-\includegraphics[width=1.05000\textwidth]{dm results.png}\\
-\emph{Table 3 Results of the Diebold-Mariano Hypothesis Test}\\
+time horizons.
+
+\begin{table}[H]
+\centering
+\begin{tabular}{lll}
+  \hline
+Forecast Horizon, h & Loss Function & P Value \\ 
+  \hline
+1 & Linear & 0.553 \\ 
+    & Quadratic & 0.128 \\ 
+  5 & Linear & 0.691 \\ 
+    & Quadratic & 0.334 \\ 
+  20 & Linear & 0.660 \\ 
+    & Quadratic & 0.296 \\ 
+   \hline
+\end{tabular}
+\caption{Results of the Diebold-Mariano Hypothesis Test for different loss functions} 
+\end{table}
+
 The p-value across all the time horizons for either linear and quadratic

### 2017-10-17 template/Draft_Paper.Rmd
Added the DM Test mathematical formulation and moved section 2.1.4 to 2.4

@@
 The most popular being its inability to capture non-linear patterns in data, even after its evolution from the standard form to more adaptable formulations [@moreno2011artificial].
-Over the past two to three decades with the evolution of computational power and statistical advancement, other stock price prediction methods have been proposed, the most prominent being a machine learning approach called Artificial Neural Networks(ANNs)[@lin2009short].
+Over the past two to three decades with the evolution of computational power and statistical advancement, other stock price prediction methods have been proposed, the most prominent being a machine learning approach called Artificial Neural Networks(ANNs) [@lin2009short].
 Also popular amongst the new stock price forecasting approaches are hybrids of existing methods that incorporate the benefits of different approaches.
@@
 
-###Diebold-Mariano Evaluation
-As is evident from the previous sections, the modelling approach and assumptions of Prophet differ to those of an ARIMA model.
-Since the two methods differ, the Diebold-Mariano evaluation (DM Test) is preferred as a means of testing forecasting accuracy.
-This is because the DM Test is a model free test of forecasting accuracy.
-It is applicable to a wide range of situations from multi-period forecasts, non-Gaussian forecast residuals, non-quadratic loss functions and even serially correlated data[@Mariano2000].
-
 ## ARIMA Model Formulation
@@
 Furthermore, these approaches tend to be inflexible with respect to incorporating prior information.
-Though this can be partially achieved by initializing parameter estimates for some methods (such as ANNs) ,an analyst with no statistical training can have difficulty interpreting and thus modifying parameter estimates[@taylor2017forecasting].
+Though this can be partially achieved by initializing parameter estimates for some methods (such as ANNs) ,an analyst with no statistical training can have difficulty interpreting and thus modifying parameter estimates [@taylor2017forecasting].
 Prophet is an **automatic forecasting** tool that uses a Bayesian Generalized Additive model to generate time series forecasts.
@@
 In Prophet, SHF errors are compared visually to allow the analysts to easily adjust the interpretable parameters before automatically estimating the model and evaluating forecasts.
+
+##Diebold-Mariano Evaluation
+As is evident from sections 2.2 and 2.3, the modelling approach and assumptions of Prophet differ to those of an ARIMA model.
+Since the two methods differ, the Diebold-Mariano evaluation is preferred as a means of comparing the forecast accuracy of the two different methods.
+The Diebold-Mariano evaluation is a **model free** hypothesis test for the equal predictive accuracy of two forecasting methods.
+It is applicable to a wide range of situations from forecast residuals that are non-Gaussian, asymmetric loss functions, serially correlated errors and multi-period forecasts [@Mariano2000].
+The Diebold-Mariano test (DM test) operates by evaluating the forecast errors for both model 1 and model 2 (i.e $e_{t_1}$ and $e_{t_2}$) at a loss function $g(\bullet)$.
+The loss function can be choosen flexibly as long as it is zero when the error is zero, is stricly positive for all values on the real line and, it increases with an increase in the absolute value of the error size.
+The loss differential between two forecasts at time $t$ is then defined by, $$ d_t =  g(e_{t1}) - g(e_{t2})$$ for $t = (1, 2, ..., M)$ where $M$ is the total number of forecasts that have been generated from each model.
+We then test whether, $$ \mu = E[d_t] = 0 $$ by conducting the following hypothesis test;
+Null: $$ E[d_t] = 0  \,\,\,\, \forall t$$   
+Alternative : $$ E[d_t] \neq 0 $$  
+Test statistic: $$DM_{stat} = \frac{\sqrt M (\bar{d} - \mu)}{\sqrt{\hat{\gamma_d(0)}}} $$
+
+$\bar{d}$ is the sample mean of the $M$ loss differentials and $\frac{\hat{\gamma_d(0)}}{M}$ is the asymptotic variance of  ($\bar{d} - \mu$) with $\hat{\gamma_d(k)}$ representing the autocovariance of the loss differential series at lag $k=0$.
+Asymptotically, the $DM_{stat}$ has a standard normal distribution if the loss differential series is short memory and covariance stationary [@Mariano2000].
+There are two caveats to using the DM Test for the purposes of this paper.
+Firstly,both a linear and quadratic loss function was  <<<<<<**correct the tense!!!!!**>>>>> used to conduct the DM Test since literature does not clearly stipulate which method is more reliable **CITE???**.
+Using either a linear or quadratic loss function implicitly assumes that the forecast errors are symmetric.
+The assumption  which is a reasonable assumption since **<<<<<<JUSTIFY THE ASSUMPTION BY REFERENCING THE BOXPLOT WHERE ERRORS LOOKED SYMMETRIC >>>>>>>**.
+secondly, literature does not clearly stipulate what sample size is sufficient for the asymptotic distribution assumption to hold.
+In response to this, 20% of the available data was used as the out-of-sample period.
+Nonethless, the DM test results must used conservatively .
 

### 2017-10-17 template/Draft_Paper.tex
Added the DM Test mathematical formulation and moved section 2.1.4 to 2.4

@@
 being a machine learning approach called Artificial Neural
-Networks(ANNs)(Lin, Yang, and Song 2009). Also popular amongst the new
+Networks(ANNs) (Lin, Yang, and Song 2009). Also popular amongst the new
 stock price forecasting approaches are hybrids of existing methods that
@@
 
-\subsubsection{Diebold-Mariano
-Evaluation}\label{diebold-mariano-evaluation}
-
-As is evident from the previous sections, the modelling approach and
-assumptions of Prophet differ to those of an ARIMA model. Since the two
-methods differ, the Diebold-Mariano evaluation (DM Test) is preferred as
-a means of testing forecasting accuracy. This is because the DM Test is
-a model free test of forecasting accuracy. It is applicable to a wide
-range of situations from multi-period forecasts, non-Gaussian forecast
-residuals, non-quadratic loss functions and even serially correlated
-data(Mariano 2000).
-
 \subsection{ARIMA Model Formulation}\label{arima-model-formulation}
@@
 statistical training can have difficulty interpreting and thus modifying
-parameter estimates(Taylor and Letham 2017).\\
+parameter estimates (Taylor and Letham 2017).\\
 Prophet is an \textbf{automatic forecasting} tool that uses a Bayesian
@@
 before automatically estimating the model and evaluating forecasts.
+
+\subsection{Diebold-Mariano
+Evaluation}\label{diebold-mariano-evaluation}
+
+As is evident from sections 2.2 and 2.3, the modelling approach and
+assumptions of Prophet differ to those of an ARIMA model. Since the two
+methods differ, the Diebold-Mariano evaluation is preferred as a means
+of comparing the forecast accuracy of the two different methods. The
+Diebold-Mariano evaluation is a \textbf{model free} hypothesis test for
+the equal predictive accuracy of two forecasting methods. It is
+applicable to a wide range of situations from forecast residuals that
+are non-Gaussian, asymmetric loss functions, serially correlated errors
+and multi-period forecasts (Mariano 2000). The Diebold-Mariano test (DM
+test) operates by evaluating the forecast errors for both model 1 and
+model 2 (i.e \(e_{t_1}\) and \(e_{t_2}\)) at a loss function
+\(g(\bullet)\). The loss function can be choosen flexibly as long as it
+is zero when the error is zero, is stricly positive for all values on
+the real line and it increases with an increase in the absolute value of
+the error size. The loss differential between two forecasts at time
+\(t\) is then defined by, \[ d_t =  g(e_{t1}) - g(e_{t2})\] for
+\(t = (1, 2, ..., M)\) where \(M\) is the total number of forecasts that
+have been generated from each model. We then test whethe,r
+\[ \mu = E[d_t] = 0 \] under the following hypothesis test;\\
+Null: \[ E[d_t] = 0  \,\,\,\, \forall t\]\\
+Alternative : \[ E[d_t] \neq 0 \]\\
+Test statistic:
+\[DM_{stat} = \frac{\sqrt M (\bar{d} - \mu)}{\sqrt{\hat{\gamma_d(0)}}} \]
+
+\(\bar{d}\) is the sample mean of the \(M\) loss differentials and
+\(\frac{\hat{\gamma_d(0)}}{M}\) is the asymptotic variance of
+(\(\bar{d} - \mu\)) with \(\hat{\gamma_d(k)}\) representing the
+autocovariance of the loss differential series at lag \(k=0\). The
+asymptotic distribution of \(DM_{stat}\) is then standard normal
+asumming the loss differential series is short memory and covariance
+stationary (Mariano 2000).\\
+There are two caveats to using the DM Test for the puposes of this
+paper. Firstly,both a linear and quadratic loss function was
+\textless{}\textless{}\textless{}\textless{}\textless{}\textless{}\textbf{correct
+the
+tense!!!!!}\textgreater{}\textgreater{}\textgreater{}\textgreater{}\textgreater{}
+used to conduct the DM Test since literature does not clearly stipulate
+which method is more reliable \textbf{CITE???}. Using either a linear or
+quadratic loss function implicitly assumes that the forecast errors are
+symmetric. The assumption which is a reasonable assumption since
+\textbf{\textless{}\textless{}\textless{}\textless{}\textless{}\textgreater{}\textgreater{}\textgreater{}\textgreater{}\textgreater{}\textgreater{}}.
+secondly, literature does not clearly stipulate what sample size is
+sufficient for the asymptotic distribution assumption to hold. In
+response to this, 20\% of the available data was used as the
+out-of-sample period. Nonethless, the DM test results must used
+conservatively .
 

### 2017-10-18 forecast plot.R
forecast plot without confidence interval

@@ -1,0 +1,30 @@
+#plotting forecasts
+
+library(ggplot2)
+library(scales)
+
+
+results.data.1 <- read.table("1 step ahead results.txt", sep = " ")
+head(results.data.1)
+
+arima_gg1 <- ggplot(data=results.data.1, aes(x = date, y = actual, group = 1)) # y = (Pt+1 /Pt)-1
+arima_gg1 <- arima_gg1+geom_line(col = 'red')
+arima_gg1 <- arima_gg1+geom_line(aes(y=fore_model_1),col='blue')
+arima_gg1 <- arima_gg1+geom_line(aes(y=fore_model_2),col='green')
+arima_gg1
+
+
+
+#Sample from this link  : https://www.r-bloggers.com/plotting-forecast-objects-in-ggplot-part-2-visualize-observations-fits-and-forecasts/ 
+
+p1a<-ggplot(data=pd,aes(x=date,y=observed)) 
+p1a<-p1a+geom_line(col='red')
+p1a<-p1a+geom_line(aes(y=fitted),col='blue')
+#This ads the grey confidence interval
+p1a<-p1a+geom_line(aes(y=forecast))+geom_ribbon(aes(ymin=lo95,ymax=hi95),alpha=.25)
+p1a<-p1a+scale_x_date(name='',breaks='1 year',minor_breaks='1 month',labels=date_format("%b-%y"),expand=c(0,0))
+p1a<-p1a+scale_y_continuous(name='Units of Y')
+p1a<-p1a+opts(axis.text.x=theme_text(size=10),title='Arima Fit to Simulated Data\n (black=forecast, blue=fitted, red=data, shadow=95% conf. interval)')
+#p1a
+arima_gg1
+

### 2017-10-18 forecast plot.R
forecast plot without confidence interval

@@ -7,24 +7,8 @@ library(scales)
 results.data.1 <- read.table("1 step ahead results.txt", sep = " ")
 head(results.data.1)
 
-arima_gg1 <- ggplot(data=results.data.1, aes(x = date, y = actual, group = 1)) # y = (Pt+1 /Pt)-1
-arima_gg1 <- arima_gg1+geom_line(col = 'red')
+arima_gg1 <- ggplot(data=results.data.1, aes(x = as.Date(date), y = actual, group = 1)) 
+arima_gg1 <- arima_gg1+geom_line(col = 'brown')
 arima_gg1 <- arima_gg1+geom_line(aes(y=fore_model_1),col='blue')
 arima_gg1 <- arima_gg1+geom_line(aes(y=fore_model_2),col='green')
 arima_gg1
-
-
-
-#Sample from this link  : https://www.r-bloggers.com/plotting-forecast-objects-in-ggplot-part-2-visualize-observations-fits-and-forecasts/ 
-
-p1a<-ggplot(data=pd,aes(x=date,y=observed)) 
-p1a<-p1a+geom_line(col='red')
-p1a<-p1a+geom_line(aes(y=fitted),col='blue')
-#This ads the grey confidence interval
-p1a<-p1a+geom_line(aes(y=forecast))+geom_ribbon(aes(ymin=lo95,ymax=hi95),alpha=.25)
-p1a<-p1a+scale_x_date(name='',breaks='1 year',minor_breaks='1 month',labels=date_format("%b-%y"),expand=c(0,0))
-p1a<-p1a+scale_y_continuous(name='Units of Y')
-p1a<-p1a+opts(axis.text.x=theme_text(size=10),title='Arima Fit to Simulated Data\n (black=forecast, blue=fitted, red=data, shadow=95% conf. interval)')
-#p1a
-arima_gg1
-

### 2017-10-19 template/Draft_Paper.Rmd
Update Draft_Paper.Rmd

@@
 
-ARIMA models rely on the assumption that residuals are heteroskedastic(they have a constant variance) and normally distributed, however, some financial data displays heteroskedasticity making GARCH models more applicable.
+ARIMA models rely on the assumption that residuals are homoskedastic(they have a constant variance) and normally distributed, however, some financial data displays heteroskedasticity making GARCH models more applicable.
 This theory is backed by a study conducted by @Muten2014 who compared the ability of ARIMA and GARCH models to forecast stock prices on the Zimbabwean Stock Exchange (ZSE).
@@
 There are two caveats to using the DM Test for the purposes of this paper.
-Firstly,both a linear and quadratic loss function was  <<<<<<**correct the tense!!!!!**>>>>> used to conduct the DM Test since literature does not clearly stipulate which method is more reliable **CITE???**.
-Using either a linear or quadratic loss function implicitly assumes that the forecast errors are symmetric.
-The assumption  which is a reasonable assumption since **<<<<<<JUSTIFY THE ASSUMPTION BY REFERENCING THE BOXPLOT WHERE ERRORS LOOKED SYMMETRIC >>>>>>>**.
-secondly, literature does not clearly stipulate what sample size is sufficient for the asymptotic distribution assumption to hold.
-In response to this, 20% of the available data was used as the out-of-sample period.
-Nonethless, the DM test results must used conservatively .
+Firstly, the choice of the loss function depends on whether the forecast errors can be assumed to be symmetric or asymmetric.
+When the forecast errors are symmetric, a linear or quadratic loss function can be used to evaluate the loss differential.
+For forecast errors that are asymmetric, a loss fuction that satisfies the conditions outlined in the first part of this section can be used.
+Secondly, literature does not clearly stipulate the sample size range that is sufficient for the asymptotic distribution assumption to  hold.
+This implies that the DM test results must be used conservatively.
+<<<ASK HANJO IF HE KNOWS OF ANY PAPERS, I'VE BEEN OUT OF LUCK WITH MY SEARCH>>>
 
@@
 The data was then split into in-sample and out-of-sample periods.
-The in-sample period comprises of 1400 trading days from 30 July 2012 to the 1 May 2016.
-The out-of-sample period comprises of the remaining 366 trading days from the dataset.
+As was mentioned in section 2.4, literature does not clearly stipulate what sample size is sufficient for the asymptotic distribution assumption to hold.
+In response to this, 20% of the available data was used as the out-of-sample period.
+The in-sample period comprises of 1400 trading days from 30 July 2012 to the 30 June 2016 whereas the out-of-sample period comprises of the remaining 365 trading days from the dataset.
 
@@
 $$ \hat{e_t} = r_t - \hat{r_t},$$  where $r_t$ is the actual return and $\hat{r_t}$ is the rolling forecast at time $t$.  
-Summary statistics of the errors where used to compare the performance of the forecasting models before a formal statistical test was done using a Diebold-Mariano evaluation.
+Summary statistics of the errors where used to compare the performance of the forecasting models before a formal statistical test was done using the Diebold-Mariano evaluation with both a linear and quadratic loss function.
+Using either a linear or quadratic loss function implicitly assumes that the forecast errors are symmetric [@Mariano2000].
+This assumption is reasonable given the apparent symmetry in the distribution of forecast errors as is evident from the boxplots in Figure 5.
 
@@
 The candidate models that were fit to the in-sample data are shown in Table 1 with their respective AIC values.
-The candidate model with the lowest AIC is the $ARIMA(2,0,2)$.
-This model was thus selected as the ARIMA forecasting model.
+The $ARIMA(2,0,2)$ model had the lowest AIC and was thus selected as the ARIMA forecasting model.
 
@@
 ## Forecast Error Evaluation
+
+>>>CONSIDER INSERTING THE FORECASTING PLOTS SOMEWHERE IN THIS SECTION HERE<<<<
 The results of model estimation and forecast evaluation using the rolling window for h-step forecasts yielded forecasts errors with roughly equal distributions.

### 2017-11-06 template/Draft_Paper.Rmd
Final draft with all edits

@@
 # The descriptions below should be self-explanatory
-title: "Comparing Facebook's Prophet Forecasting to an ARIMA Model: A Diebold-Mariano Evaluation of the Top40 Index"
+title: "Comparing rolling period forecasts of Facebooks' Prophet model to an ARIMA model: A Diebold-Mariano evaluation of the FTSE/JSE Top40 Index"
 subtitle: "A Diebold-Mariano Evaluation of the FTSE/JSE Top40 Index"
@@
 Prophet employs intensive Bayesian modelling in two tiers.
-Purely through MAP parameter estimation, and in a quasi-Bayesian form through an automated forecast evaluation that enables analyst to improve the model if it falls short when compared to other models.
-This paper assess the predictive accuracy of Prophets' forecasts to those of an ARIMA model that was identified using the Box-Jenkins methodology.
+Purely through MAP parameter estimation, and in a quasi-Bayesian form through an automated forecast evaluation that enables analyst to improve a forecasting model if it underperforms when compared to other traditional models.
+This paper assesses the predictive accuracy of Prophets' forecasts to those of an ARIMA model that was identified using the Box-Jenkins methodology.
 This is achieved through a Diebold-Mariano evaluation of the rolling period forecast errors of returns on the FTSE/JSE Top40 Index across daily, weekly and monthly time horizons.
-The results of the study suggest that there is insufficient evidence to conclude that the forecasting models have unequal predictive ability within the scope of the JSE/Top40 Index.
+The results of the study suggest that there is insufficient evidence to conclude that the forecasting models have unequal predictive ability within the scope of the FTSE/JSE Top40 Index.
 ---
@@
 However, the forecasts generated by these methods have been found to be brittle as well as inflexible in allowing for prior information to be incorporated into the forecasting model [@taylor2017forecasting].
-In February of 2017, Facebook's Core Datascience team open sourced ,Prophet, their time series forecasting tool whose main selling point is that it enables "forecasting-at-scale".
-A term coined by the data science team which is defined as, "an approach that allows a large number of analysts to forecast a large number and variety of business time series" [@taylor2017forecasting].
+In February of 2017, Facebook's Core Datascience team open sourced Prophet, a time series forecasting tool with unique features that enable "forecasting-at-scale".
+Forecasting-at-scale is a term coined by the data science team which they define as, "an approach that allows a large number of analysts to forecast a large number and variety of business time series" [@taylor2017forecasting].
 It has the added advantage of being less 'expensive'
 than other alternatives.
-In addition, it is specifically prefarable over ARIMA models because of its non-linearity, flexibility and ability to accomodate varying time intervals [@taylor2017forecasting].
+In addition, it is specifically preferable over ARIMA models because of its non-linearity, flexibility and ability to accomodate varying time intervals [@taylor2017forecasting].
 
-The objective of this paper is to compare the rolling period forecasts generated by Prophet (which uses automated, intensive Bayesian Modelling to 'forecast-at-scale'), to those obtained by an ARIMA model (constructed through the Box-Jenkins methodology),
-through a Diebold-Mariano evaluation of the JSE Top40 Index.
-The study is particularly useful because analysts from a wide background can automatically forecast stock price returns, and incorporate their prior knowledge in the model , without a wholesome understanding of the statistical intricacies involved.
+The objective of this paper is to compare the rolling period forecasts generated by Prophet (which uses automated, intensive Bayesian Modelling to 'forecast-at-scale'), to those obtained from an ARIMA model (constructed through the Box-Jenkins methodology),
+through a Diebold-Mariano evaluation of the FTSE/JSE Top40 Index.
+The study is particularly useful because analysts from a wide background can reliably forecast stock price returns automatically, and incorporate their prior knowledge in the forecasting model, without a wholesome understanding of the statistical intricacies involved.
+It is hoped that this could create opportunities for automated invesment mechanisms.
 
 The paper is organized as follows.
-A review of literature relating to stock price forecasting with ARIMA models and other alternative approaches is introduced to give the reader background information on research relating to the subject matter.
-Thereafter, theory relating to the mathematical formulation and estimation of both the ARIMA and Prophet models is included in sections 2.2 and 2.3 respectively, to build the readers understanding of what the models look like and how they operate.
+A review of literature relating to stock price forecasting with ARIMA models and other alternative approaches is introduced in section 2.1 to give the reader background information on existing research relating to the subject matter.
+Thereafter, theory relating to the mathematical formulation and estimation of an ARIMA and Prophet model is included in sections 2.2 and 2.3 respectively.
+Subsequently, in section 2.4, the theoretical motivation and mathematical formulation of the Diebold-Mariano test is introduced as a means of comparing the predictive accuracy of two prediction models.
 Information relating to the data that was used in this study is contained in section 3, after which, the methodology used to answer the research question is described in section 4.
-Section 5, contains the results obtained when the methodology was executed, whilst section 6 includes a discussion of the results and how they can be interpreted to answer the research question.
-Lastly, the conclusion in section 7 highlights the main findings of the study and recommends further areas of research.
+Section 5, contains the results obtained when the methodology was executed, whilst section 6 includes a discussion of the results.
+The discussion includes a response to the research question as well as some useful insights from the results.
+Furthermore, it highlights some recommendations for further areas of research based on the findings of the paper.
+Lastly, the conclusion in section 7 summarizes the results of the study relative to the main objective of the paper.
 
@@
 In this section we will briefly review literature that relates to the comparison of different approaches to forecasting stock returns.
-This will include both traditional approaches as well new methods such as automatic forecasting, machine learning techniques and quasi-Bayesian forms which incorporate analysts' prior information.
-The review of literature is concluded with a brief introduction to the Diebold-Mariano evaluation and an outline of its suitability for the purpose of this study.
+This will include both traditional approaches as well as relatively new methods such as automatic forecasting, machine learning techniques and quasi-Bayesian forms which incorporate analysts' prior information.
 The theoretical and mathematical formulation of an ARIMA forecasting model is  then briefly introduced.
@@
 Specific attention is drawn to Prophets distinguising features to give the reader an understanding of the value that this forecasting technique can potentially add to stock price prediction.
-Lastly, the Diebold-Mariano Evaluation technique is introduced as a method of formally testing superior predictive accuracy between two forecasting models.
+The background information is concluded with an introduction to the Diebold-Mariano Evaluation as a suitable framework to formally compare the predictive accuracy of two forecasting models.
 
@@
 
-ARIMA models have been widely used as a standard model to forecast financial data and have been shown to yield acceptable forecasting errors [@adebiyi2014comparison].
+ARIMA models have been widely used as a standard model to forecast financial data and have been shown to yield forecasting errors that lie within acceptable bounds [@adebiyi2014comparison].
 @adebiyi2014comparison used the Box-Jenkins methodology to build short-term stock price prediction models for stocks on the New York Stock Exchange(NYSE) and the Nigerian Stock Exchange(NSE).
-Their results showed that the prediction errors of the model were within acceptable bounds.
+Their results showed that the predictive power of the ARIMA model was satisfactory based on forecast error diagnostics.
 
-ARIMA models rely on the assumption that residuals are homoskedastic(they have a constant variance) and normally distributed, however, some financial data displays heteroskedasticity making GARCH models more applicable.
+ARIMA models rely on the assumption that residuals are homoskedastic(they have a constant variance) and normally distributed.
+However, the residual errors of some financial data display heteroskedasticity making GARCH models more applicable.
 This theory is backed by a study conducted by @Muten2014 who compared the ability of ARIMA and GARCH models to forecast stock prices on the Zimbabwean Stock Exchange (ZSE).
-They found that the GARCH model outperforms the ARIMA model which suggests that the residuals are heteroskedastic.
+They found that the GARCH model outperforms the ARIMA model which suggests that incorporating heteroskedasticity of the residual errors improves the forecasting ability of the model.
 This is a case in point for emerging market stocks.
@@
 
-Artificial Neural Networks (ANNs) are a multi-layered perceptron as illustrated in  figure 1.  
+An Artificial Neural Networks (ANNs) is a multi-layered perceptron with nodes that simulate the action of neurons as illustrated in  figure 1.  
 ![](ANN.png)  
-*Figure 1 : Illustration of an Artificial neural network* [@tanikic2012artificial]  
-An ANN consists of a sorted triple $(K, A, \omega)$.
-$K$ is a set representing the multiple layers/levels of the network i.e.
+**Figure 1 : Illustration of an Artificial neural network** [@tanikic2012artificial]  
+Formally, an ANN consists of a sorted triple $(K, A, \omega)$.
+$K$ is a set representing the multiple levels of the network i.e.
 input, hidden and output levels.
-A is a set of pairs that represents index for the estimate in the $i^{th}$ node of the $k^{th}$ layer of the network.
-In the enlarged section of Figure 1, $a_{i,k} \in A$.
-Lastly, $\omega$ represents a function that defines the weights ($\omega_{i,r}$) of connections between neurons/nodes [@kriesel2007brief].
+A is a set of pairs that represents the index for the estimated parameter at the $i^{th}$ node of the $k^{th}$ level in the network.
+This is illustrated by the curved down arrow in Figure 1 which points to an enlarged segment of the $i^{th}$ node in the $k^{th}$ level i.e.
+$a_{i,k} \in A$.
+Lastly, $\omega$ represents a function that defines the weights ($\omega_{i,r}$) of connections between nodes that interact across adjacent levels [@kriesel2007brief].
 Artificial Neural networks apply iterated optimization of model parameters across the network in the form of weights conditional on observed values to learn [@segaran2007programming].
@@
 They used the Box-Jenkins methodology to difference the series and then trained the data on a neural network with initialisations that were guided by their observations of the ARIMA model.
-This approach yielded forecasts with an acceptable prediction accuracy based on residual analysis of the out-of-sample data.
-Wang and Leu therefore concluded that the ARIMA-based Neural Network outperformed a Neural Network trained using raw stock price data [@wang1996stock].
+Their hybrid model yielded forecasts that were comparably superior to the individual models used in isolation based on an analysis of the out-of-sample residual error diagnostics.
+@wang1996stock therefore concluded that the ARIMA-based Neural Network outperformed a Neural Network trained using raw stock price data [@wang1996stock].
 @zhang2009stock used a combination of the backpropagation algorithm from Neural Networks with Improved Bacterial Chemotaxis Optimization (IBCO) to build a model that forecasts the S&P 500 stock index by minimizing the mean square error.
@@
 An issue that arises with such methods is the complexity of the proposed models which limits the flexibility of unseasoned analysts to adjust model parameters as a way of improving forecasting accuracy.
-As we will see, Prophet forecasting elegantly deals with this issue in the form of a nonlinear, Generalized Additive Model (GAM).
+As we will see, Prophet elegantly deals with this issue in the form of a nonlinear, Generalized Additive Model (GAM).
 
-###Quasi-Bayesian methods incorporating Analysts' knowledge
+###Quasi-Bayesian forecasting methods that incorporate an analysts' knowledge
 
 The stock price forecasting approaches considered so far purely apply technical data analysis methods to generate stock price forecasting models.
-However, studies that have compared pure technical approaches to those that incorporate the analysts knowledge in a quasi-Bayesian form have been shown to be superior at predicting market prices [@givoly1984quality \& @guerard1989combining].
+However, studies that have compared pure technical approaches to those that incorporate an analysts knowledge in a quasi-Bayesian form have been shown to be superior at predicting market prices [@givoly1984quality \& @guerard1989combining].
 The first instance of combining time-series model forecasts and an analysts forecasts to obtain superior forecasts of a stocks annual earnings can be dated to 1989.
-@guerard1989combining used an additive model to combine consensus security analyst forecasts from the S&P Annual Earnings forecaster and annual earnings forecasts generated by an ARIMA model with a constant.
+@guerard1989combining used an additive model to combine consensus security analyst forecasts from the S&P Annual Earnings forecaster and annual earnings forecasts generated by an ARIMA model with a nonzero mean.
 The combined model that was estimated using ordinary least squares reduced the mean square error of the time series and analysts forecasts from 1.28 and 1.27, respectively to 1.04.
-These results suggest that analysts can substantially reduce forecasting errors by combining both approaches [@guerard1989combining].
-@zahedi2015application applied the same overarching principle when they used principle component analysis to determine an appropriate input variable, which they then used to train an ANN in an attempt to predict stock prices on the Tehran Stock Exchange.
-Their model yielded acceptable forecasting errors and performed better than the pure ANN which further reiterates the potential of incorporating an analysts knowledge to improve stock price forecasts, in both linear and nonlinear models.
+These results suggest that analysts can reduce forecasting errors by combining technical methods with expertise developed through market knowledge [@guerard1989combining].
+@zahedi2015application applied the same overarching principle when they used principle component analysis to determine an appropriate input variable which they then used to train an ANN in an attempt to predict stock prices on the Tehran Stock Exchange.
+Their model yielded reduced forecasting errors and performed better than the pure ANN.
+These results further reiterate the potential of incorporating an analysts knowledge to improve stock price forecasts, in both linear and nonlinear models.
 
-## ARIMA Model Formulation
+## ARIMA model formulation
 
@@
 $$  r_t = \phi_1 r_{t-1} + \phi_2 r_{t-2} + ...+ \phi_p r_{t-p} + \epsilon_t +
-        \psi_1 \epsilon_{t-1} + \psi_2 \epsilon_{t-2} + ... +\psi_q \epsilon_{t-q} $$  
-where:  
-$\bullet \phi_i$ and $\psi_j$ are the parameters to be estimated;  
-$\bullet r_{t-k}$ is the $k^{th}$ lagged return;   
-$\bullet \epsilon_t$ is the error term at time $t$ which is assumed to be white noise and  
-$\bullet\epsilon_{t-k}$ is the $k^{th}$ lagged error.   
-$\bullet$p is the order of the autoregressive component (dependenceon history of $r_t$)
-$\bullet$q is the order of the moving average component (dependenceon history of $\epsilon_t$)
+        \psi_1 \epsilon_{t-1} + \psi_2 \epsilon_{t-2} + ... +\psi_q \epsilon_{t-q} $$where:  
+$\bullet$  $\phi_i$ and $\psi_j$ are the parameters to be estimated  
+$\bullet$  $r_{t-k}$ is the $k^{th}$ lagged return   
+$\bullet$  $\epsilon_t$ is the error term at time $t$ which is assumed to be white noise  
+$\bullet$  $\epsilon_{t-k}$ is the $k^{th}$ lagged error   
+$\bullet$  $p$ is the order of the autoregressive component (dependence on history of $r_t$) and  
+$\bullet$  $q$ is the order of the moving average component (dependence on history of $\epsilon_t$).
 
@@
 
-## Prophet Model Formulation and Estimation
+## Prophet model formulation and estimation
 
-The challenge with many forecasting methods such as the Box-Jenkins methodology and machine-learning approaches are that they usually require a sufficient statistical background.
-Furthermore, these approaches tend to be inflexible with respect to incorporating prior information.
-Though this can be partially achieved by initializing parameter estimates for some methods (such as ANNs) ,an analyst with no statistical training can have difficulty interpreting and thus modifying parameter estimates [@taylor2017forecasting].
+The challenge with many forecasting methods such as the Box-Jenkins methodology and machine learning approaches discussed in the literature review are that they usually require a sufficient statistical background to construct and modify if necessary.
+Furthermore, some of these approaches tend to be inflexible with respect to incorporating prior information.
+In some instances, such as with ANN, incorporating prior information can be partially achieved through initializing parameter estimates.
+However, an analyst with limited statistical training can have difficulty interpreting and thus modifying initial parameter estimates [@taylor2017forecasting].
 Prophet is an **automatic forecasting** tool that uses a Bayesian Generalized Additive model to generate time series forecasts.
 This is in comparison to conventional methods such as the linear stochastic dependence that exists in ARIMA models.
-Prophet combines a configurable model that includes performance analysis evaluation with the interaction of an analyst in the loop [@taylor2017forecasting].
+Prophet combines a configurable model that includes an automated evaluation of forecasts with the interaction of an analyst in the loop [@taylor2017forecasting].
 The model is configurable which allows an analyst to incorporate their knowledge of the behaviour of the series into the model building process through easily interpretable initial parameters that can be modified and interactive feedback when forecasts under-perform [@taylor2017forecasting].
@@
 
-  The first step of the forecasting procedure begins with the analyst simply specifying a general model.
-Prophet then goes into the automated process of estimating model parameters and doing forecast evaluation.
-If forecasts are considered problematic based on evaluations of the Simulated Historical Forecast (SHF) errors, then the package will surface these problems.
-The process then transitions to the top left of the cycle in Figure 2 labelled analyst-in-the-loop.
-Here, the analyst can then use visually presented aspects of the results that raised the issues to modify model parameters before a better model is then estimated.
-The loop is repeated until forecast evaluation surfaces no problems.
+  The first step of the forecasting procedure begins with the analyst simply specifying a general model (i.e.
+modelling).
+Prophet then goes into the automated process of estimating model parameters and performing forecast evaluation.
+Automated forecast evaluation is undertaken by constructing baseline forecasts using methods such as the sample mean, ARIMA, exponential smoothing and naive estimates (e.g.
+last value or seasonal last value).
+Thereafter, simulated historical forecasting (SHF) errors are evaluated by comparing forecasts of the baseline methods and prophet to the observed historical values.
+SHF errors are forecast errors generated by random points in the history of a time series model.
+Prophets SHF errors that are relatively large compared to the baseline methods are then surfaced with visual illustrations of the results.
+The process then transitions from surface problems to the top left of the cycle in Figure 2 labelled analyst-in-the-loop.
+Here, the analyst can then visually inspect forecasts to identify issues and modify features of the Prophet model before it is re-estimated in the modelling stage.
+The loop is repeated until forecast evaluation surfaces no problems [@taylor2017forecasting].
 
-![Figure 2](analyst in the loop.png){ width=100% }  
-*Figure 2: Illustration of Prophets' automated forecasting procedure with an analyst in the loop* [@taylor2017forecasting]
+![Figure 2](analyst in the loop.png){ width=80% }  
+**Figure 2: Illustration of Prophets' automated forecasting procedure with an analyst in the loop** [@taylor2017forecasting]
 
-The interpretability of Prophets' model parameters stems from the mathematical formulation of the forecasting model.
+The interpretability of Prophets' model parameters stems from the mathematical formulation of the generalised additive model.
 The formulation posits that the series is generated by an additive, parametric function of time $r(t)$, defined by,
@@
 
-The growth component, $g(t)$, is modelled using a generalized form of the logistic population growth model and is of the form;  
+Firstly, the growth component, $g(t)$, is modelled using a generalized form of the logistic population growth model and is of the form;  
 $$g(t) = \frac{C(t)}{1+exp(-(k+\boldsymbol{a}(t)^T \delta(t-(b+\boldsymbol{a}(t)^T \gamma)))}$$  
-Where:  
-$\bullet C(t)$ represents the carrying capacity of the growth component and can be modelled using a polynomial function of time, the simplest being a constant or linear model.  
-$\bullet (k+\boldsymbol{a}(t)^T\delta)$ is the growth rate factor with the $\boldsymbol{a}(t)^T\delta$ term enabling the forecaster to choose where the growth rate changes( i.e change points) and   
-$\bullet (b+\boldsymbol{a}(t)^T \gamma)$ is the adjusted offset parameter.
-
+where:  
+$\bullet$   $C(t)$ represents the carrying capacity of the growth component and can be modelled using a polynomial function of time, the simplest being a constant or linear model.  
+$\bullet$   $(k+\boldsymbol{a}(t)^T\delta)$ is the growth rate factor with the $\boldsymbol{a}(t)^T\delta$ term enabling the forecaster to choose where the growth rate changes( i.e change points) and
+$\bullet$   $(b+\boldsymbol{a}(t)^T \gamma)$ is the adjusted offset parameter.  
 The growth component of the model has two useful features that allow the carrying capacity and growth rate of the model to vary with time.
@@
 This feature is particularly useful for forecasting stock price data since analysts are able to incorporate market movements that lead to the growth rate either decreasing, increasing or becoming constant.
-The seasonality component, $s(t)$ with periodicity $P$ is given by,
+
+Secondly, the seasonality component, $s(t)$ with periodicity $P$ is given by,
 $$s(t) = \sum_{n = -N}^{N}c_n \exp(j\frac{2\pi nt}{P})$$
 
-Yearly and weekly seasonality is modelled using the standard Fourier series.
-Analysts can thus use prior knowledge to account for the effects of periodicities they have observed seasonally which is similar to the periodicity of Seasonal ARIMA models.
+Yearly and weekly seasonality is modelled using the standard Fourier series. Analysts can thus use prior knowledge to account for the effects of periodicities they have observed seasonally.
 
@@
 MAP estimation finds the maximum posterior estimates for the parameters by using the likelihood function and an apriori distribution to evaluate a posterior distribution which is then optimized to return point estimates of the parameters [@taylor2017forecasting].
-Thereafter, the package automatically evaluates the forecasts by comparing the Simulated Historical Forecast (SHF) errors of baseline forecasting methods (including ARIMA forecasts) to those of Prophet.
-SHF errors are forecast errors generated from random points in the history of the time series model.
-In Prophet, SHF errors are compared visually to allow the analysts to easily adjust the interpretable parameters before automatically estimating the model and evaluating forecasts.
 
-##Diebold-Mariano Evaluation
+##Diebold-Mariano evaluation
 As is evident from sections 2.2 and 2.3, the modelling approach and assumptions of Prophet differ to those of an ARIMA model.
-Since the two methods differ, the Diebold-Mariano evaluation is preferred as a means of comparing the forecast accuracy of the two different methods.
-The Diebold-Mariano evaluation is a **model free** hypothesis test for the equal predictive accuracy of two forecasting methods.
+Since the two methods differ, the Diebold-Mariano evaluation is preferred as a means of formally comparing the forecasting accuracy of the two models statistically.
+The Diebold-Mariano evaluation is a **model free** hypothesis test for the equal predictive accuracy of two forecasting methods, referred to as model 1 and model 2.
 It is applicable to a wide range of situations from forecast residuals that are non-Gaussian, asymmetric loss functions, serially correlated errors and multi-period forecasts [@Mariano2000].
 The Diebold-Mariano test (DM test) operates by evaluating the forecast errors for both model 1 and model 2 (i.e $e_{t_1}$ and $e_{t_2}$) at a loss function $g(\bullet)$.
-The loss function can be choosen flexibly as long as it is zero when the error is zero, is stricly positive for all values on the real line and, it increases with an increase in the absolute value of the error size.
+The loss function can be choosen flexibly as long as it is zero when the error is zero, is strictly positive for all values on the real line and, it increases with an increase in the absolute value of the error size.
 The loss differential between two forecasts at time $t$ is then defined by, $$ d_t =  g(e_{t1}) - g(e_{t2})$$ for $t = (1, 2, ..., M)$ where $M$ is the total number of forecasts that have been generated from each model.
 We then test whether, $$ \mu = E[d_t] = 0 $$ by conducting the following hypothesis test;
-Null: $$ E[d_t] = 0  \,\,\,\, \forall t$$   
-Alternative : $$ E[d_t] \neq 0 $$  
+Null hypothesis: $$ E[d_t] = 0  \,\,\,\, \forall t$$   
+Alternative hypothesis: $$ E[d_t] \neq 0 $$  
 Test statistic: $$DM_{stat} = \frac{\sqrt M (\bar{d} - \mu)}{\sqrt{\hat{\gamma_d(0)}}} $$
@@
 Secondly, literature does not clearly stipulate the sample size range that is sufficient for the asymptotic distribution assumption to  hold.
-This implies that the DM test results must be used conservatively.
-<<<ASK HANJO IF HE KNOWS OF ANY PAPERS, I'VE BEEN OUT OF LUCK WITH MY SEARCH>>>
+In light of this, relatively large sample sizes were used in this study and the DM test results are applied conservatively.
 
@@
 ```
-*Figure 3: Realisation of the FTSE/JSE Top40 Price Index over the period of study*
+**Figure 3: Realisation of the FTSE/JSE Top40 Price Index over the period of study**
 
@@
 In response to this, 20% of the available data was used as the out-of-sample period.
-The in-sample period comprises of 1400 trading days from 30 July 2012 to the 30 June 2016 whereas the out-of-sample period comprises of the remaining 365 trading days from the dataset.
+The in-sample period comprises of 1044 trading days between 30 July 2012 to the 31 July 2016 whereas the out-of-sample period comprises of the trading days between 1 August 2016 and 31 July 2017.
 
@@
 Thereafter, one-step rolling forecasts with re-estimation as outlined by @RJHyndman2014 were evaluated for an automated Prophet model and the ARIMA model that was identified using the Box-Jenkins methodology.
-This procedure was then repeated for a 5-step and 20-step forecast horizon in order to determine whether the performance of the models vary when estimating weekly or monthly stock returns.
+This procedure was then repeated for a 5-step and 20-step forecast horizon in order to determine whether the performance of the models vary when estimating daily,weekly or monthly stock returns.
 
-The forecasts generated above where then used to calculate **h-step rolling forecast errors** for both the ARIMA and Prophet models.
+The forecasts generated above were then used to calculate **h-step rolling forecast errors** for both the ARIMA and Prophet model.
 For each time point $t$ in the out-of-sample period, the rolling forecast error is given by,
@@
 $$ \hat{e_t} = r_t - \hat{r_t},$$  where $r_t$ is the actual return and $\hat{r_t}$ is the rolling forecast at time $t$.
-Summary statistics of the errors where used to compare the performance of the forecasting models before a formal statistical test was done using the Diebold-Mariano evaluation with both a linear and quadratic loss function.
+
+Summary statistics of the errors where used to compare the performance of the forecasting models before a formal statistical test was performed using the DM test with a linear and quadratic loss function.
 Using either a linear or quadratic loss function implicitly assumes that the forecast errors are symmetric [@Mariano2000].
-This assumption is reasonable given the apparent symmetry in the distribution of forecast errors as is evident from the boxplots in Figure 5.
+This assumption is reasonable given the apparent symmetry in the distribution of forecast errors as is evident from the boxplots in Figure 5, section 5.2.
 
 #Results
-##Box-Jenkins Methodology: Arima Model Identification and Selection
+##Box-Jenkins methodology: ARIMA model identification and selection
 
-To identify what ARIMA model to use, the methodology as outlined in @box1970time was applied to the data.
-Firstly, stationarity conditions were verified formally by performing an Augmented Dickey-Fuller test on the in-sample returns data.
-The results of the test suggest the p_value is less that 1%, which implies that the in-sample returns series is stationary.
-The results in the autocorrelation(ACF) and partial autocorrelation(PACF) function illustrated in Figure 4  suggest that the second lag of both functions might be significantly different from zero.
-![](acf & pacf plot.png){ width=100%}  
-*Figure 4 : ACF and PACF plots for in-sample data*
+The methodology as outlined in @box1970time was applied to the data through the following procedure.
+The stationarity of the time series was tested formally by performing an Augmented Dickey-Fuller test on the in-sample log returns data.
+The test yielded a p-value less than 1% which implies that we reject the null hypothesis which states that the in-sample returns series non-stationary.
 
-Observations from the above plots informed the choice of candidate *ARIMA(p,d,q)* models to fit to the data.
+The autocorrelation(ACF) and partial autocorrelation(PACF) functions in Figure 4 were used to identify candidate *ARIMA(p,d,q)* models to fit to the in-sample data.
+The results suggest that the second lag of both functions might be significantly different from zero.
+```{r in-sample and out of sample data import and log transformation , include=FALSE}
+#Partitioning in-sample and out-of-sample data
+library(PerformanceAnalytics)
+in.sample.data <- index %>% filter(between(date, as.POSIXct("2012-07-31"), as.POSIXct("2016-07-31")))
+
+oos.data <- index %>% filter(between(date, as.POSIXct("2016-07-29"), as.POSIXct("2017-07-31")))  #choose the 29th to be able to calc the first oos return
+
+#Transforming data to log returns
+in.samp.log.returns <- CalculateReturns(as.ts(in.sample.data[,2]), method = "log")[-1] 
+oos.log.returns <- CalculateReturns(as.ts(oos.data[,2]), method = "log")[-1]
+
+# Formally testing stationarity/Unit root test
+library(tseries)   #for Augmented Dickey-Fuller test
+adftest <- adf.test(in.samp.log.returns) # p < 0.01 therefore we reject the null that unit root exists
+
+```
+
+```{r acf_and_pacf_plot, fig.width=6.7, fig.height=3.5, fig.align="center"}
+#PACF and ACF
+par(mfrow = c(1,2))
+plot1 <- acf(in.samp.log.returns, main = "ACF of log returns", ylab="")  #ma(2)
+plot2 <- pacf(in.samp.log.returns, main = "PACF of log returns",ylab="")  #ar(2), ar(30)
+par(mfrow = c(1,1))
+```
+**Figure 4 : ACF and PACF plots for in-sample data**
+
 The candidate models that were fit to the in-sample data are shown in Table 1 with their respective AIC values.
@@
 
+**Table 1: AIC for candidate ARIMA models**
 ```{r AIC Table for candidate ARIMA Models, results = 'asis'}
@@
 
-table <- xtable(data, caption = "AIC for Candidate ARIMA models")
+table <- xtable(data)
   print.xtable(table, 
@@
              comment = FALSE,
-             include.rownames = getOption("xtable.include.rownames", FALSE), #exclude row number
-             caption.placement = 'bottom'
+             include.rownames = getOption("xtable.include.rownames", FALSE) #exclude row number
              )
@@
 ```
-## Forecast Error Evaluation
+## Forecast error evaluation
 
->>>CONSIDER INSERTING THE FORECASTING PLOTS SOMEWHERE IN THIS SECTION HERE<<<<
-The results of model estimation and forecast evaluation using the rolling window for h-step forecasts yielded forecasts errors with roughly equal distributions.
-This is visible from the boxplot in Figure 5.
-The presence of outliers seems consistent across the different models **and** forecasting horizons.
-Furthermore, variations in accuracy across different forecasting horizons *for each* model is not observable in Figure 5.
-The mean error and spread of both models across different time horizons appears to be constant.
-![](Forecast error boxplot.png){ width=105% }  
+The h-step ahead rolling period forecasts for the $ARIMA(2,0,2)$ and Prophet models yielded forecast errors whose distribution is shown in figure 5.
+The distribution across different time horizons for the ARIMA and Prophet models appear relatively similar.
+All six distributions are symmertrical about a mean of 0 and a spread of 0.02.
+The presence of outliers also appears to be consistent across the different models **and** forecasting horizons eventhough we can observe an increasing trend in the maximum values as the time horizon increases.
+
+\begin{figure}
+```{r box_plot, fig.width=6.5, fig.height=4.0, include=TRUE,results="hide",fig.align='center'}
+fore_res_1 <- read.table("CopyOf1 step ahead results.txt") 
+fore_res_5 <- read.table("CopyOf5 step ahead results.txt")
+fore_res_20 <- read.table("CopyOf20 step ahead results.txt")
+
+library(plotrix)
+par(mfrow=c(1,1))
+
+names = c("Arima(h = 1)","Prophet(h = 1)",
+          "Arima(h = 5)","Prophet(h = 5)",
+          "Arima(h = 20)","Prophet(h = 20)")
+
+boxplot(fore_res_1$error1, fore_res_1$error2,
+        fore_res_5$error1,fore_res_5$error2,
+        fore_res_20$error1,fore_res_20$error2,
+        main = "Rolling Period Forecast Errors",xlab = "",xaxt = "n")
+staxlab(1,1:6,names, srt = 45,cex = 1)
+
+```
+\end{figure}
 *Figure 5: Boxplot illustrating distribution of rolling forecast errors for different time horizons, h*
@@
 Table 2 summarizes the forecast error statistics for both the ARIMA and Prophet model across the three different time horizons.  
-Across all forecast horizons, the ARIMA model appears to be overestimating returns.
+For all forecast horizons, the ARIMA model appears to be overestimating returns.
 This is evident from the negative mean error(ME) across $h = (1,5,20)$.
@@
 It then overstimates returns when forecasting twenty-steps ahead.
-The ME results shown in Table 2 fail to clearly illustrate which model performs better.
-Furthermore, the ME values are relatively small when compared against other error statistics.
+Furthermore, the relative margin of the distance between the ME as the time horizons increasese is relatively larger for the Prophet model than it is for the ARIMA model.
+This could mean that the function of the ME against the forecast horizon is more volatile for the Prophet model than it is for the ARIMA model.
+This observation suggests that the Prophet model might have unstable forecasts across longer time horizons and could be inappropriate for use outside short term horizons in the domain of stock price data.
 
+**Table 2: Forecast error diagnostic statistics over different forecasting horizons**
 ```{r Forecast Error Statistics Table, results = 'asis'}
@@
 
-error.table <- xtable(err.stat.data, caption = "Forecast error diagnostic statistics over different forecasting horizons")
+error.table <- xtable(err.stat.data)
   print.xtable(error.table, 
@@
              comment = FALSE,
-             include.rownames = getOption("xtable.include.rownames", FALSE), #exclude row number
-             caption.placement = 'bottom'
+             include.rownames = getOption("xtable.include.rownames", FALSE) #exclude row number
+
              )
@@
 The Root Mean Square Error (RMSE) across different forecast horizons and across both models are considerably close to each other in value relative to the other error statistics.
-The same observation can be made for the Mean Absolute Errors (MAE).
 These results suggest that the models possibly have the same forecasting accuracy.
+Furthermore, the ME and MAE results shown in Table 2 are comparably small which further suggests that the models possibly have the same forecasting accuracy.
+However, comparing the predictive accuracy of the two models using the error diagnostics alone is insufficient.
 
@@
 These assertions can be formally quantified through a Diebold-Mariano(DM) test which evaluates whether the models have equal predictive accuracy against the alternative hypothesis that one model has superior accuracy over the other.
-Table 3 reports the results obtained when the rolling forecast errors were formally compared against each other over different time horizons.
+Table 3 reports the results obtained when the rolling forecast errors were formally compared against each other over different time horizons using either a linear or quadratic loss function.
+Both loss functions are appropriate for forecast errors that have an apparent symmetrical distribution (see figure 5) and were both used to gain insight into any discrepancies that might arise.
 
+**Table 3: Results of the Diebold-Mariano hypothesis test for different loss functions**
 ```{r Table For Diebold Mariano Evaluation results, results = 'asis'}
@@
 dimnames(dm.mat) <- list(c(" "," "," "," "," "," "),   #row names
-                          c("Forecast Horizon, h","Loss Function","P Value"))  #column names
+                          c("Forecast Horizon, h","Loss Function , g(*)","p value"))  #column names
 dm.data = dm.mat %>% tbl_df()
 
-dm.table <- xtable(dm.data, caption = "Results of the Diebold-Mariano Hypothesis Test for different loss functions")
+dm.table <- xtable(dm.data)
   print.xtable(dm.table, 
@@
 
-The p-value across all the time horizons for either linear and quadratic loss functions suggest that we would fail to reject the null hypothesis.
+The p-value across all the time horizons for either a linear or quadratic loss function suggest that we would fail to reject the null hypothesis.
 This implies that the predictive accuracy of the models could be equal.
-These results formally validate the observations made obtained in section 5.2.1.
+These results formally validate the observations made in section 5.2.1.
 
-#Discussion of Results
+#Discussion of results
 The results of the study show that there is no significant evidence to conclude that Prophet generates superior forecasts for returns on the FTSE/JSE Top40 Index when compared against those of an ARIMA model when the Diebold-Mariano Test is used as a method of evaluation.
-These results are consistent for daily, weekly and monthly forecasting horizons.
+It is worth noting that the p-value for the quadratic loss function is, on average, lower than that of the linear loss function.
+The sensitivity of the results to the nature of the loss function is worth exploring in the future.
+The results are consistent for daily, weekly and monthly forecasting horizons irregardless of the choice of the loss function.
+
 Nevertheless, the study has flagged the possibility that, on average, the ARIMA model tends to overestimate forecasted returns on the index for the forecast horizons under consideration.
-This is contrary to Prophet, that, on average, underestimates forecasted returns over daily and weekly time horizons.
-This information could be useful for an analyst engaged in short-term strategies since the choice of the forecasting model can reflect a particular sentiment or risk appetite.
+This is contrary to Prophet, that, on average, underestimateD forecasted returns over daily and weekly time horizons.
+This information could be useful for an analyst engaged in **short-term** strategies since the choice of the forecasting model can reflect a particular sentiment or risk appetite.
 For example, if the analyst feels like prevailing returns are underpriced by the market, they might choose to use an ARIMA model, whereas an opposite sentiment could warrant the use of Prophet.
-Furthermore, it is worth noting that the merits of Prophet may have not been fully exposed in this study due to the use of transformed data and the choice of a forecasting horizon that excluded periods of high volatility.
-Nonetheless, it is reassuring that the model has performed as well as the ARIMA model.
-This observation could create room for the application of Prophet in circumstances where flexibility in the model formulation is desireable.
+A necessary caveat is the volatility of Prophet forecasting errors about the mean error and the ability to release exploitable gains in reality due to transaction costs.
+Furthermore, it is worth noting that the merits of Prophet may have not been fully explored in this study due to the use of transformed data and the choice of a forecasting horizon that excluded periods of high volatility.
+These areas create room for further study.
+Nonetheless, it is reassuring (and somewhat expected  because the ARIMA methodology is incorporated in the forecast evaluation of Prophet as a baseline method), that Prophet has performed as well as the ARIMA model.
+This observation could create room for the application of Prophet in circumstances where flexibility in the model formulation is desireable, or if the analyst has a limited statistical background.
+
+This study was confined to the analysis of predictive performance during general market conditions.
+Further research could explore the predictive ability of the models during different market conditions.
+For example, performance can be compared during bull and bear market conditions or during periods of high volatility.
+Additionally, further research can assess whether prophet would perform better on raw data that has not been transformed, since the model is built with growth and seasonality components.
+The use of which could lead to Prophet models that generate superior forecasts when compared against alternatives.
+
+Lastly, it is worth noting that the efficiency of the South African stock market is inconclusive.
+This means that forecasting the FTSE/JSE Top40 index might be a futile exercise if markets are efficient.
+However, there is room for traders with a lmited statistical background to exploit existing inefficiencies in some industries and market categories.
+This
 
@@
 The implications of this are that analysts can then choose which model to use based on their needs without significantly compromising forecast  accuracy.
-This study was confined to the analysis of predictive performance during general market conditions.
-Further research could explore the predictive ability of the models during different market conditions.
-For example, performance can be compared during bull and bear market conditions or during periods of high volatility.
-Additionally, further research can assess whether prophet would perform better on raw data that has not been transformed, since the model is built with growth and seasonality components.
-This use of which could lead to Prophet models that generate superior forecasts when compared against alternatives.
+The results also highlighted the possibility that ARIMA models tend to overestimate returns and thus using Prophet would imply a more conservative approach to predicting future returns, this could have widespread applications for investment strategies, one of which was highlighted above.
 

### 2017-11-06 template/Draft_Paper.tex
Final draft with all edits

--- 2017-10-14 template/Draft_Paper.tex
@@
 \begin{center}
-{\Huge \bf Comparing Facebook's Prophet Forecasting to an ARIMA Model: A
-Diebold-Mariano Evaluation of the Top40 Index}\\
+{\Huge \bf Comparing rolling period forecasts of Facebooks' Prophet model to an
+ARIMA model: A Diebold-Mariano evaluation of the FTSE/JSE Top40 Index}\\
 \today\\
 Marvelous Mubenesha ( MBNMAR005)\\
-{\tt MBNMAR005@uct.co.za}
+{\tt MBNMAR005@myuct.co.za}
 \end{center}
@@
 through MAP parameter estimation, and in a quasi-Bayesian form through
-an automated forecast evaluation that enables analyst to improve the
-model if it falls short when compared to other models. This paper assess
-the predictive accuracy of Prophets' forecasts to those of an ARIMA
-model that was identified using the Box-Jenkins methodology. This is
-achieved through a Diebold-Mariano evaluation of the rolling period
-forecast errors of returns on the FTSE/JSE Top40 Index across daily,
-weekly and monthly time horizons. The results of the study suggest that
-there is insufficient evidence to conclude that the forecasting models
-have unequal predictive ability within the scope of the JSE/Top40 Index.
+an automated forecast evaluation that enables analyst to improve a
+forecasting model if it underperforms when compared to other traditional
+models. This paper assesses the predictive accuracy of Prophets'
+forecasts to those of an ARIMA model that was identified using the
+Box-Jenkins methodology. This is achieved through a Diebold-Mariano
+evaluation of the rolling period forecast errors of returns on the
+FTSE/JSE Top40 Index across daily, weekly and monthly time horizons. The
+results of the study suggest that there is insufficient evidence to
+conclude that the forecasting models have unequal predictive ability
+within the scope of the FTSE/JSE Top40 Index.
 \noindent
@@
 Letham 2017). In February of 2017, Facebook's Core Datascience team open
-sourced ,Prophet, their time series forecasting tool whose main selling
-point is that it enables ``forecasting-at-scale''. A term coined by the
-data science team which is defined as, ``an approach that allows a large
-number of analysts to forecast a large number and variety of business
-time series'' (Taylor and Letham 2017). It has the added advantage of
-being less `expensive' than other alternatives. In addition, it is
-specifically prefarable over ARIMA models because of its non-linearity,
-flexibility and ability to accomodate varying time intervals (Taylor and
-Letham 2017).
+sourced Prophet, a time series forecasting tool with unique features
+that enable ``forecasting-at-scale''. Forecasting-at-scale is a term
+coined by the data science team which they define as, ``an approach that
+allows a large number of analysts to forecast a large number and variety
+of business time series'' (Taylor and Letham 2017). It has the added
+advantage of being less `expensive' than other alternatives. In
+addition, it is specifically preferable over ARIMA models because of its
+non-linearity, flexibility and ability to accomodate varying time
+intervals (Taylor and Letham 2017).
 
@@
 generated by Prophet (which uses automated, intensive Bayesian Modelling
-to `forecast-at-scale'), to those obtained by an ARIMA model
+to `forecast-at-scale'), to those obtained from an ARIMA model
 (constructed through the Box-Jenkins methodology), through a
-Diebold-Mariano evaluation of the JSE Top40 Index. The study is
-particularly useful because analysts from a wide background can
-automatically forecast stock price returns, and incorporate their prior
-knowledge in the model , without a wholesome understanding of the
-statistical intricacies involved.
+Diebold-Mariano evaluation of the FTSE/JSE Top40 Index. The study is
+particularly useful because analysts from a wide background can reliably
+forecast stock price returns automatically, and incorporate their prior
+knowledge in the forecasting model, without a wholesome understanding of
+the statistical intricacies involved. It is hoped that this could create
+opportunities for automated invesment mechanisms.
 
@@
 stock price forecasting with ARIMA models and other alternative
-approaches is introduced to give the reader background information on
-research relating to the subject matter. Thereafter, theory relating to
-the mathematical formulation and estimation of both the ARIMA and
-Prophet models is included in sections 2.2 and 2.3 respectively, to
-build the readers understanding of what the models look like and how
-they operate. Information relating to the data that was used in this
-study is contained in section 3, after which, the methodology used to
-answer the research question is described in section 4. Section 5,
+approaches is introduced in section 2.1 to give the reader background
+information on existing research relating to the subject matter.
+Thereafter, theory relating to the mathematical formulation and
+estimation of an ARIMA and Prophet model is included in sections 2.2 and
+2.3 respectively. Subsequently, in section 2.4, the theoretical
+motivation and mathematical formulation of the Diebold-Mariano test is
+introduced as a means of comparing the predictive accuracy of two
+prediction models. Information relating to the data that was used in
+this study is contained in section 3, after which, the methodology used
+to answer the research question is described in section 4. Section 5,
 contains the results obtained when the methodology was executed, whilst
-section 6 includes a discussion of the results and how they can be
-interpreted to answer the research question. Lastly, the conclusion in
-section 7 highlights the main findings of the study and recommends
-further areas of research.
+section 6 includes a discussion of the results. The discussion includes
+a response to the research question as well as some useful insights from
+the results. Furthermore, it highlights some recommendations for further
+areas of research based on the findings of the paper. Lastly, the
+conclusion in section 7 summarizes the results of the study relative to
+the main objective of the paper.
 
@@
 comparison of different approaches to forecasting stock returns. This
-will include both traditional approaches as well new methods such as
-automatic forecasting, machine learning techniques and quasi-Bayesian
-forms which incorporate analysts' prior information. The review of
-literature is concluded with a brief introduction to the Diebold-Mariano
-evaluation and an outline of its suitability for the purpose of this
-study. The theoretical and mathematical formulation of an ARIMA
-forecasting model is then briefly introduced. Thereafter, the
-theoretical framework and mathematical formulation of Facebooks' Prophet
-forecasting tool is discussed. Specific attention is drawn to Prophets
-distinguising features to give the reader an understanding of the value
-that this forecasting technique can potentially add to stock price
-prediction. Lastly, the Diebold-Mariano Evaluation technique is
-introduced as a method of formally testing superior predictive accuracy
-between two forecasting models.
+will include both traditional approaches as well as relatively new
+methods such as automatic forecasting, machine learning techniques and
+quasi-Bayesian forms which incorporate analysts' prior information. The
+theoretical and mathematical formulation of an ARIMA forecasting model
+is then briefly introduced. Thereafter, the theoretical framework and
+mathematical formulation of Facebooks' Prophet forecasting tool is
+discussed. Specific attention is drawn to Prophets distinguising
+features to give the reader an understanding of the value that this
+forecasting technique can potentially add to stock price prediction. The
+background information is concluded with an introduction to the
+Diebold-Mariano Evaluation as a suitable framework to formally compare
+the predictive accuracy of two forecasting models.
 
@@
 ARIMA models have been widely used as a standard model to forecast
-financial data and have been shown to yield acceptable forecasting
-errors (Adebiyi, Adewumi, and Ayo 2014). Adebiyi, Adewumi, and Ayo
-(2014) used the Box-Jenkins methodology to build short-term stock price
-prediction models for stocks on the New York Stock Exchange(NYSE) and
-the Nigerian Stock Exchange(NSE). Their results showed that the
-prediction errors of the model were within acceptable bounds.
+financial data and have been shown to yield forecasting errors that lie
+within acceptable bounds (Adebiyi, Adewumi, and Ayo 2014). Adebiyi,
+Adewumi, and Ayo (2014) used the Box-Jenkins methodology to build
+short-term stock price prediction models for stocks on the New York
+Stock Exchange(NYSE) and the Nigerian Stock Exchange(NSE). Their results
+showed that the predictive power of the ARIMA model was satisfactory
+based on forecast error diagnostics.
 
 ARIMA models rely on the assumption that residuals are
-heteroskedastic(they have a constant variance) and normally distributed,
-however, some financial data displays heteroskedasticity making GARCH
-models more applicable. This theory is backed by a study conducted by
-Mutendadzamera and Mutasa (2014) who compared the ability of ARIMA and
-GARCH models to forecast stock prices on the Zimbabwean Stock Exchange
-(ZSE). They found that the GARCH model outperforms the ARIMA model which
-suggests that the residuals are heteroskedastic. This is a case in point
-for emerging market stocks. The researchers noted that poor liquidity in
-the market could be a cause of the results they observed (Mutendadzamera
-and Mutasa 2014).
+homoskedastic(they have a constant variance) and normally distributed.
+However, the residual errors of some financial data display
+heteroskedasticity making GARCH models more applicable. This theory is
+backed by a study conducted by Mutendadzamera and Mutasa (2014) who
+compared the ability of ARIMA and GARCH models to forecast stock prices
+on the Zimbabwean Stock Exchange (ZSE). They found that the GARCH model
+outperforms the ARIMA model which suggests that incorporating
+heteroskedasticity of the residual errors improves the forecasting
+ability of the model. This is a case in point for emerging market
+stocks. The researchers noted that poor liquidity in the market could be
+a cause of the results they observed (Mutendadzamera and Mutasa 2014).
 
@@
 
-Artificial Neural Networks (ANNs) are a multi-layered perceptron as
-illustrated in figure 1.\\
+An Artificial Neural Networks (ANNs) is a multi-layered perceptron with
+nodes that simulate the action of neurons as illustrated in figure 1.\\
 \includegraphics{ANN.png}\\
@@
 and Despotovic 2012)\\
-An ANN consists of a sorted triple \((K, A, \omega)\). \(K\) is a set
-representing the multiple layers/levels of the network i.e.~input,
-hidden and output levels. A is a set of pairs that represents index for
-the estimate in the \(i^{th}\) node of the \(k^{th}\) layer of the
-network. In the enlarged section of Figure 1, \(a_{i,k} \in A\). Lastly,
-\(\omega\) represents a function that defines the weights
-(\(\omega_{i,r}\)) of connections between neurons/nodes (Kriesel 2007).
-Artificial Neural networks apply iterated optimization of model
+Formally, an ANN consists of a sorted triple \((K, A, \omega)\). \(K\)
+is a set representing the multiple levels of the network i.e.~input,
+hidden and output levels. A is a set of pairs that represents the index
+for the estimated parameter at the \(i^{th}\) node of the \(k^{th}\)
+level in the network. This is illustrated by the curved down arrow in
+Figure 1 which points to an enlarged segment of the \(i^{th}\) node in
+the \(k^{th}\) level i.e. \(a_{i,k} \in A\). Lastly, \(\omega\)
+represents a function that defines the weights (\(\omega_{i,r}\)) of
+connections between nodes that interact across adjacent levels (Kriesel
+2007). Artificial Neural networks apply iterated optimization of model
 parameters across the network in the form of weights conditional on
@@
 neural network with initialisations that were guided by their
-observations of the ARIMA model. This approach yielded forecasts with an
-acceptable prediction accuracy based on residual analysis of the
-out-of-sample data. Wang and Leu therefore concluded that the
-ARIMA-based Neural Network outperformed a Neural Network trained using
-raw stock price data (Wang and Leu 1996). Zhang and Wu (2009) used a
-combination of the backpropagation algorithm from Neural Networks with
-Improved Bacterial Chemotaxis Optimization (IBCO) to build a model that
-forecasts the S\&P 500 stock index by minimizing the mean square error.
-Model forecasts were evaluated through simulation experiments and the
-results led to the conclusion that the hybrid model produced superior
-forecasts (Zhang and Wu 2009). This study further highlighted the
-potential that nonlinear approaches have in forecasting stock prices. An
-issue that arises with such methods is the complexity of the proposed
-models which limits the flexibility of unseasoned analysts to adjust
-model parameters as a way of improving forecasting accuracy. As we will
-see, Prophet forecasting elegantly deals with this issue in the form of
-a nonlinear, Generalized Additive Model (GAM).
+observations of the ARIMA model. Their hybrid model yielded forecasts
+that were comparably superior to the individual models used in isolation
+based on an analysis of the out-of-sample residual error diagnostics.
+Wang and Leu (1996) therefore concluded that the ARIMA-based Neural
+Network outperformed a Neural Network trained using raw stock price data
+(Wang and Leu 1996). Zhang and Wu (2009) used a combination of the
+backpropagation algorithm from Neural Networks with Improved Bacterial
+Chemotaxis Optimization (IBCO) to build a model that forecasts the S\&P
+500 stock index by minimizing the mean square error. Model forecasts
+were evaluated through simulation experiments and the results led to the
+conclusion that the hybrid model produced superior forecasts (Zhang and
+Wu 2009). This study further highlighted the potential that nonlinear
+approaches have in forecasting stock prices. An issue that arises with
+such methods is the complexity of the proposed models which limits the
+flexibility of unseasoned analysts to adjust model parameters as a way
+of improving forecasting accuracy. As we will see, Prophet elegantly
+deals with this issue in the form of a nonlinear, Generalized Additive
+Model (GAM).
 
-\subsubsection{Quasi-Bayesian methods incorporating Analysts'
-knowledge}\label{quasi-bayesian-methods-incorporating-analysts-knowledge}
+\subsubsection{Quasi-Bayesian forecasting methods that incorporate an
+analysts'
+knowledge}\label{quasi-bayesian-forecasting-methods-that-incorporate-an-analysts-knowledge}
 
@@
 models. However, studies that have compared pure technical approaches to
-those that incorporate the analysts knowledge in a quasi-Bayesian form
+those that incorporate an analysts knowledge in a quasi-Bayesian form
 have been shown to be superior at predicting market prices (Givoly and
@@
 forecasts from the S\&P Annual Earnings forecaster and annual earnings
-forecasts generated by an ARIMA model with a constant. The combined
+forecasts generated by an ARIMA model with a nonzero mean. The combined
 model that was estimated using ordinary least squares reduced the mean
@@
 1.27, respectively to 1.04. These results suggest that analysts can
-substantially reduce forecasting errors by combining both approaches
-(Guerard 1989). Zahedi and Rounaghi (2015) applied the same overarching
-principle when they used principle component analysis to determine an
-appropriate input variable, which they then used to train an ANN in an
-attempt to predict stock prices on the Tehran Stock Exchange. Their
-model yielded acceptable forecasting errors and performed better than
-the pure ANN which further reiterates the potential of incorporating an
-analysts knowledge to improve stock price forecasts, in both linear and
-nonlinear models.
+reduce forecasting errors by combining technical methods with expertise
+developed through market knowledge (Guerard 1989). Zahedi and Rounaghi
+(2015) applied the same overarching principle when they used principle
+component analysis to determine an appropriate input variable which they
+then used to train an ANN in an attempt to predict stock prices on the
+Tehran Stock Exchange. Their model yielded reduced forecasting errors
+and performed better than the pure ANN. These results further reiterate
+the potential of incorporating an analysts knowledge to improve stock
+price forecasts, in both linear and nonlinear models.
 
-\subsubsection{Diebold-Mariano
-Evaluation}\label{diebold-mariano-evaluation}
-
-As is evident from the previous sections, the modelling approach and
-assumptions of Prophet differ to those of an ARIMA model. Since the two
-methods differ, the Diebold-Mariano evaluation (DM Test) is preferred as
-a means of testing forecasting accuracy. This is because the DM Test is
-a model free test of forecasting accuracy. It is applicable to a wide
-range of situations from multi-period forecasts, non-Gaussian forecast
-residuals, non-quadratic loss functions and even serially correlated
-data(Mariano 2000).
-
-\subsection{ARIMA Model Formulation}\label{arima-model-formulation}
+\subsection{ARIMA model formulation}\label{arima-model-formulation}
 
@@
 \[  r_t = \phi_1 r_{t-1} + \phi_2 r_{t-2} + ...+ \phi_p r_{t-p} + \epsilon_t +
-        \psi_1 \epsilon_{t-1} + \psi_2 \epsilon_{t-2} + ... +\psi_q \epsilon_{t-q} \]\\
-where:\\
-\(\bullet \phi_i\) and \(\psi_j\) are the parameters to be estimated;\\
-\(\bullet r_{t-k}\) is the \(k^{th}\) lagged return;\\
-\(\bullet \epsilon_t\) is the error term at time \(t\) which is assumed
-to be white noise and\\
-\(\bullet\epsilon_{t-k}\) is the \(k^{th}\) lagged error.\\
-\(\bullet\)p is the order of the autoregressive component (dependenceon
-history of \(r_t\)) \(\bullet\)q is the order of the moving average
-component (dependenceon history of \(\epsilon_t\))
+        \psi_1 \epsilon_{t-1} + \psi_2 \epsilon_{t-2} + ... +\psi_q \epsilon_{t-q} \]where:\\
+\(\bullet\) \(\phi_i\) and \(\psi_j\) are the parameters to be
+estimated\\
+\(\bullet\) \(r_{t-k}\) is the \(k^{th}\) lagged return\\
+\(\bullet\) \(\epsilon_t\) is the error term at time \(t\) which is
+assumed to be white noise\\
+\(\bullet\) \(\epsilon_{t-k}\) is the \(k^{th}\) lagged error\\
+\(\bullet\) \(p\) is the order of the autoregressive component
+(dependence on history of \(r_t\)) and\\
+\(\bullet\) \(q\) is the order of the moving average component
+(dependence on history of \(\epsilon_t\)).
 
@@
 
-\subsection{Prophet Model Formulation and
-Estimation}\label{prophet-model-formulation-and-estimation}
+\subsection{Prophet model formulation and
+estimation}\label{prophet-model-formulation-and-estimation}
 
 The challenge with many forecasting methods such as the Box-Jenkins
-methodology and machine-learning approaches are that they usually
-require a sufficient statistical background. Furthermore, these
+methodology and machine learning approaches discussed in the literature
+review are that they usually require a sufficient statistical background
+to construct and modify if necessary. Furthermore, some of these
 approaches tend to be inflexible with respect to incorporating prior
-information. Though this can be partially achieved by initializing
-parameter estimates for some methods (such as ANNs) ,an analyst with no
-statistical training can have difficulty interpreting and thus modifying
-parameter estimates(Taylor and Letham 2017).\\
+information. In some instances, such as with ANN, incorporating prior
+information can be partially achieved through initializing parameter
+estimates. However, an analyst with limited statistical training can
+have difficulty interpreting and thus modifying initial parameter
+estimates (Taylor and Letham 2017).\\
 Prophet is an \textbf{automatic forecasting} tool that uses a Bayesian
@@
 dependence that exists in ARIMA models. Prophet combines a configurable
-model that includes performance analysis evaluation with the interaction
-of an analyst in the loop (Taylor and Letham 2017). The model is
-configurable which allows an analyst to incorporate their knowledge of
-the behaviour of the series into the model building process through
-easily interpretable initial parameters that can be modified and
+model that includes an automated evaluation of forecasts with the
+interaction of an analyst in the loop (Taylor and Letham 2017). The
+model is configurable which allows an analyst to incorporate their
+knowledge of the behaviour of the series into the model building process
+through easily interpretable initial parameters that can be modified and
 interactive feedback when forecasts under-perform (Taylor and Letham
@@
 The first step of the forecasting procedure begins with the analyst
-simply specifying a general model. Prophet then goes into the automated
-process of estimating model parameters and doing forecast evaluation. If
-forecasts are considered problematic based on evaluations of the
-Simulated Historical Forecast (SHF) errors, then the package will
-surface these problems. The process then transitions to the top left of
-the cycle in Figure 2 labelled analyst-in-the-loop. Here, the analyst
-can then use visually presented aspects of the results that raised the
-issues to modify model parameters before a better model is then
-estimated. The loop is repeated until forecast evaluation surfaces no
-problems.
+simply specifying a general model (i.e.~modelling). Prophet then goes
+into the automated process of estimating model parameters and performing
+forecast evaluation. Automated forecast evaluation is undertaken by
+constructing baseline forecasts using methods such as the sample mean,
+ARIMA, exponential smoothing and naive estimates (e.g.~last value or
+seasonal last value). Thereafter, simulated historical forecasting (SHF)
+errors are evaluated by comparing forecasts of the baseline methods and
+prophet to the observed historical values. SHF errors are forecast
+errors generated by random points in the history of a time series model.
+Prophets SHF errors that are relatively large compared to the baseline
+methods are then surfaced with visual illustrations of the results. The
+process then transitions from surface problems to the top left of the
+cycle in Figure 2 labelled analyst-in-the-loop. Here, the analyst can
+then visually inspect forecasts to identify issues and modify features
+of the Prophet model before it is re-estimated in the modelling stage.
+The loop is repeated until forecast evaluation surfaces no problems
+(Taylor and Letham 2017).
 
-\includegraphics[width=1.00000\textwidth]{analyst in the loop.png}\\
+\includegraphics[width=0.80000\textwidth]{analyst in the loop.png}\\
 \emph{Figure 2: Illustration of Prophets' automated forecasting
@@
 The interpretability of Prophets' model parameters stems from the
-mathematical formulation of the forecasting model. The formulation
-posits that the series is generated by an additive, parametric function
-of time \(r(t)\), defined by,
+mathematical formulation of the generalised additive model. The
+formulation posits that the series is generated by an additive,
+parametric function of time \(r(t)\), defined by,
 \[ r(t) = g(t) + s(t) +  h(t) + \epsilon_t.\] The additive components
@@
 
-The growth component, \(g(t)\), is modelled using a generalized form of
-the logistic population growth model and is of the form;\\
+Firstly, the growth component, \(g(t)\), is modelled using a generalized
+form of the logistic population growth model and is of the form;\\
 \[g(t) = \frac{C(t)}{1+exp(-(k+\boldsymbol{a}(t)^T \delta(t-(b+\boldsymbol{a}(t)^T \gamma)))}\]\\
-Where:\\
-\(\bullet C(t)\) represents the carrying capacity of the growth
+where:\\
+\(\bullet\) \(C(t)\) represents the carrying capacity of the growth
 component and can be modelled using a polynomial function of time, the
 simplest being a constant or linear model.\\
-\(\bullet (k+\boldsymbol{a}(t)^T\delta)\) is the growth rate factor with
-the \(\boldsymbol{a}(t)^T\delta\) term enabling the forecaster to choose
-where the growth rate changes( i.e change points) and\\
-\(\bullet (b+\boldsymbol{a}(t)^T \gamma)\) is the adjusted offset
-parameter.
-
+\(\bullet\) \((k+\boldsymbol{a}(t)^T\delta)\) is the growth rate factor
+with the \(\boldsymbol{a}(t)^T\delta\) term enabling the forecaster to
+choose where the growth rate changes( i.e change points) and\\
+\(\bullet\) \((b+\boldsymbol{a}(t)^T \gamma)\) is the adjusted offset
+parameter.\\
 The growth component of the model has two useful features that allow the
@@
 analysts are able to incorporate market movements that lead to the
-growth rate either decreasing, increasing or becoming constant.\\
-The seasonality component, \(s(t)\) with periodicity \(P\) is given by,
-\[s(t) = \sum_{n = -N}^{N}c_n \exp(j\frac{2\pi nt}{P})\]
+growth rate either decreasing, increasing or becoming constant.
+
+Secondly, the seasonality component, \(s(t)\) with periodicity \(P\) is
+given by, \[s(t) = \sum_{n = -N}^{N}c_n \exp(j\frac{2\pi nt}{P})\]
 
@@
 series. Analysts can thus use prior knowledge to account for the effects
-of periodicities they have observed seasonally which is similar to the
-periodicity of Seasonal ARIMA models.
+of periodicities they have observed seasonally.
 
@@
 optimized to return point estimates of the parameters (Taylor and Letham
-2017). Thereafter, the package automatically evaluates the forecasts by
-comparing the Simulated Historical Forecast (SHF) errors of baseline
-forecasting methods (including ARIMA forecasts) to those of Prophet. SHF
-errors are forecast errors generated from random points in the history
-of the time series model. In Prophet, SHF errors are compared visually
-to allow the analysts to easily adjust the interpretable parameters
-before automatically estimating the model and evaluating forecasts.
+2017).
+
+\subsection{Diebold-Mariano
+evaluation}\label{diebold-mariano-evaluation}
+
+As is evident from sections 2.2 and 2.3, the modelling approach and
+assumptions of Prophet differ to those of an ARIMA model. Since the two
+methods differ, the Diebold-Mariano evaluation is preferred as a means
+of formally comparing the forecasting accuracy of the two models
+statistically. The Diebold-Mariano evaluation is a \textbf{model free}
+hypothesis test for the equal predictive accuracy of two forecasting
+methods, referred to as model 1 and model 2. It is applicable to a wide
+range of situations from forecast residuals that are non-Gaussian,
+asymmetric loss functions, serially correlated errors and multi-period
+forecasts (Mariano 2000). The Diebold-Mariano test (DM test) operates by
+evaluating the forecast errors for both model 1 and model 2 (i.e
+\(e_{t_1}\) and \(e_{t_2}\)) at a loss function \(g(\bullet)\). The loss
+function can be choosen flexibly as long as it is zero when the error is
+zero, is strictly positive for all values on the real line and, it
+increases with an increase in the absolute value of the error size. The
+loss differential between two forecasts at time \(t\) is then defined
+by, \[ d_t =  g(e_{t1}) - g(e_{t2})\] for \(t = (1, 2, ..., M)\) where
+\(M\) is the total number of forecasts that have been generated from
+each model. We then test whether, \[ \mu = E[d_t] = 0 \] by conducting
+the following hypothesis test;\\
+Null hypothesis: \[ E[d_t] = 0  \,\,\,\, \forall t\]\\
+Alternative hypothesis: \[ E[d_t] \neq 0 \]\\
+Test statistic:
+\[DM_{stat} = \frac{\sqrt M (\bar{d} - \mu)}{\sqrt{\hat{\gamma_d(0)}}} \]
+
+\(\bar{d}\) is the sample mean of the \(M\) loss differentials and
+\(\frac{\hat{\gamma_d(0)}}{M}\) is the asymptotic variance of
+(\(\bar{d} - \mu\)) with \(\hat{\gamma_d(k)}\) representing the
+autocovariance of the loss differential series at lag \(k=0\).
+Asymptotically, the \(DM_{stat}\) has a standard normal distribution if
+the loss differential series is short memory and covariance stationary
+(Mariano 2000).\\
+There are two caveats to using the DM Test for the purposes of this
+paper. Firstly, the choice of the loss function depends on whether the
+forecast errors can be assumed to be symmetric or asymmetric. When the
+forecast errors are symmetric, a linear or quadratic loss function can
+be used to evaluate the loss differential. For forecast errors that are
+asymmetric, a loss fuction that satisfies the conditions outlined in the
+first part of this section can be used. Secondly, literature does not
+clearly stipulate the sample size range that is sufficient for the
+asymptotic distribution assumption to hold. In light of this, relatively
+large sample sizes were used in this study and the DM test results are
+applied conservatively.
 
@@
 before formally testing for outliers. The data was then split into
-in-sample and out-of-sample periods. The in-sample period comprises of
-1400 trading days from 30 July 2012 to the 1 May 2016. The out-of-sample
-period comprises of the remaining 366 trading days from the dataset.
+in-sample and out-of-sample periods. As was mentioned in section 2.4,
+literature does not clearly stipulate what sample size is sufficient for
+the asymptotic distribution assumption to hold. In response to this,
+20\% of the available data was used as the out-of-sample period. The
+in-sample period comprises of 1044 trading days between 30 July 2012 to
+the 31 July 2016 whereas the out-of-sample period comprises of the
+trading days between 1 August 2016 and 31 July 2017.
 
@@
 order to determine whether the performance of the models vary when
-estimating weekly or monthly stock returns.
+estimating daily,weekly or monthly stock returns.
 
-The forecasts generated above where then used to calculate
-\textbf{h-step rolling forecast errors} for both the ARIMA and Prophet
-models. For each time point \(t\) in the out-of-sample period, the
-rolling forecast error is given by,
+The forecasts generated above were then used to calculate \textbf{h-step
+rolling forecast errors} for both the ARIMA and Prophet model. For each
+time point \(t\) in the out-of-sample period, the rolling forecast error
+is given by,
 
 \[ \hat{e_t} = r_t - \hat{r_t},\] where \(r_t\) is the actual return and
-\(\hat{r_t}\) is the rolling forecast at time \(t\).\\
+\(\hat{r_t}\) is the rolling forecast at time \(t\).
+
 Summary statistics of the errors where used to compare the performance
-of the forecasting models before a formal statistical test was done
-using a Diebold-Mariano evaluation.
+of the forecasting models before a formal statistical test was performed
+using the DM test with a linear and quadratic loss function. Using
+either a linear or quadratic loss function implicitly assumes that the
+forecast errors are symmetric (Mariano 2000). This assumption is
+reasonable given the apparent symmetry in the distribution of forecast
+errors as is evident from the boxplots in Figure 5, section 5.2.
 
@@
 
-\subsection{Box-Jenkins Methodology: Arima Model Identification and
-Selection}\label{box-jenkins-methodology-arima-model-identification-and-selection}
+\subsection{Box-Jenkins methodology: ARIMA model identification and
+selection}\label{box-jenkins-methodology-arima-model-identification-and-selection}
 
-To identify what ARIMA model to use, the methodology as outlined in Box
-and Jenkins (1970) was applied to the data. Firstly, stationarity
-conditions were verified formally by performing an Augmented
-Dickey-Fuller test on the in-sample returns data. The results of the
-test suggest the p\_value is less that 1\%, which implies that the
-in-sample returns series is stationary. The results in the
-autocorrelation(ACF) and partial autocorrelation(PACF) function
-illustrated in Figure 4 suggest that the second lag of both functions
-might be significantly different from zero.\\
-\includegraphics[width=1.00000\textwidth]{acf \& pacf plot.png}\\
+The methodology as outlined in Box and Jenkins (1970) was applied to the
+data through the following procedure. The stationarity of the time
+series was tested formally by performing an Augmented Dickey-Fuller test
+on the in-sample log returns data. The test yielded a p-value less than
+1\% which implies that we reject the null hypothesis which states that
+the in-sample returns series non-stationary.
+
+The autocorrelation(ACF) and partial autocorrelation(PACF) functions in
+Figure 4 were used to identify candidate \emph{ARIMA(p,d,q)} models to
+fit to the in-sample data. The results suggest that the second lag of
+both functions might be significantly different from zero.
+
+\begin{center}\includegraphics{Draft_Paper_files/figure-latex/acf_and_pacf_plot-1} \end{center}
+
 \emph{Figure 4 : ACF and PACF plots for in-sample data}
 
-Observations from the above plots informed the choice of candidate
-\emph{ARIMA(p,d,q)} models to fit to the data. The candidate models that
-were fit to the in-sample data are shown in Table 1 with their
-respective AIC values. The candidate model with the lowest AIC is the
-\(ARIMA(2,0,2)\). This model was thus selected as the ARIMA forecasting
-model.
+The candidate models that were fit to the in-sample data are shown in
+Table 1 with their respective AIC values. The \(ARIMA(2,0,2)\) model had
+the lowest AIC and was thus selected as the ARIMA forecasting model.
+
+\emph{Table 1: AIC for candidate ARIMA models}
 
@@
 \end{tabular}
-\caption{AIC for Candidate ARIMA models} 
 \end{table}
 
-\subsection{Forecast Error Evaluation}\label{forecast-error-evaluation}
+\subsection{Forecast error evaluation}\label{forecast-error-evaluation}
 
@@
 different time horizons appears to be constant.
-\includegraphics[width=1.05000\textwidth]{Forecast error boxplot.png}\\
+
+\begin{figure}
+
+\begin{center}\includegraphics{Draft_Paper_files/figure-latex/box_plot-1} \end{center}
+\end{figure}
+
 \emph{Figure 5: Boxplot illustrating distribution of rolling forecast
@@
 \centering
+\caption{Forecast error diagnostic statistics over different forecasting horizons} 
 \begin{tabular}{lllll}
@@
 \end{tabular}
-\caption{Forecast error diagnostic statistics over different forecasting horizons} 
 \end{table}
@@
 \centering
+\caption{Results of the Diebold-Mariano Hypothesis Test for different loss functions} 
 \begin{tabular}{lll}
@@
 \end{tabular}
-\caption{Results of the Diebold-Mariano Hypothesis Test for different loss functions} 
 \end{table}
@@
 
-\section{Discussion of Results}\label{discussion-of-results}
+\section{Discussion of results}\label{discussion-of-results}
 

### 2017-11-06 DM Bootstrap Functions.R

@@ -1,0 +1,191 @@
+#Dummy data
+#set.seed(1)
+#e1 <- rnorm(10, mean = 0, sd = 1)
+#set.seed(2)
+#e2 <- rnorm(10, mean = 0, sd = 1) 
+
+# FUNCTION 1 ; Do a dm test on bootstrapped sample
+
+dm.bootstrap <- function(e1,e2, h, seed){
+  # WHAT THE FUNCTION DOES
+  
+  #     This function takes two error vectors which are time indexed, then constructs a bootstrap sample 
+  #     (indentical rows, with replacement) from both error vectors . The time index of error vectors is 
+  #     matched by setting the same seed for each round of sampling using set.seed(seed). [See code below]
+  
+  #ARGUMENTS
+  #     - e1 and e2 are error vectors from which we will be resampling from
+  #     - h is a parameter that represents the forecasting time horizon for the error vectors
+  #     - seed is a random integer that will be used as an arg. for set.seed so sampling from error vectors 
+  #       is matched.
+  
+  #VALUE
+  # Function returns the p.val when Diebold-Mariano test is conducted for bootstrapped sample errors
+  
+  #Step1: Select random integer to be the seed for sampling
+  #seed <- round(runif(1, min = 1, max = 10000))  #This isn't changing when I re-run the code. I'll set the see outside the function
+  
+  #Step2: Sampling from error vectors(matching of time series indices are preserved using the random seed above)
+  #Only continue with test if error vectors have the same length
+  
+  l1 <- length(e1)
+  l2 <- length(e2)
+  
+  if(l1==l2){
+    
+    set.seed(seed)
+    boot_e1 <- sample(e1, l1, replace = TRUE)
+    
+    set.seed(seed)
+    boot_e2 <- sample(e2, l2, replace = TRUE)
+    
+  } else {
+    
+    print("Error vectors do not have the same length")
+    
+  }
+  
+  test <- dm.test(boot_e1, boot_e2, h = h,  alternative = c("two.sided"))  #Need to specify the forecast horizon(h), default is 1
+  return(c(test$p.value))
+  
+}
+
+# FUNCTION 2: Repeated the dm test on a bootstrap sample N times (uses function 1)
+#Question: Think of a more efficient way to do this using dplyr
+
+repeat.dm.bootstrap <- function(e1,e2,N,h){
+  
[…]

### 2017-11-06 rolling period code.R

@@ -41,8 +41,8 @@ adf.test(in.samp.log.returns) # p < 0.01 therefore we reject the null. Data is s
 
 #    Step 2: ACF and PACF plots for model identification
 par(mfrow = c(1,2))
-acf(in.samp.log.returns, main = "Autocorrelation Function for Log Returns")  #ma(2)
-pacf(in.samp.log.returns, main = "Partial Autocorrelation Function for Log Returns")  #ar(2), ar(30)
+plot1 <- acf(in.samp.log.returns, main = "Autocorrelation Function for Log Returns")  #ma(2)
+plot2 <- pacf(in.samp.log.returns, main = "Partial Autocorrelation Function for Log Returns")  #ar(2), ar(30)
 par(mfrow = c(1,1))
 #ma and ar(30) could be overfitting and since model is for forecasting, I want it to be parsimonious ->
 #looks like there could be seasonality but its below the bandwith 
